### 2020-08-27 writing/ch1.tex
starting revisions

\hypertarget{the-construction-of-space-in-ecological-models}{%
\section{The Construction of Space in Ecological
Models}\label{the-construction-of-space-in-ecological-models}}

\begin{quote}
The electron is a theory we use; it is so useful in understanding the
way nature works that we can almost call it real.

\begin{flushright}Richard P. Feynman\end{flushright}
\end{quote}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Life on Earth takes on an endless diversity of forms. Yet, the
characteristics that define `life' impose parameters on the properties
that biological entities must exhibit, and constraints on how these
entities can change over time. These are the fundamental principles of
evolution and ecology---entities that reproduce themselves more than
average become more frequent, there is a limit on resources, and so on.
Over the billions of years life has been on this planet, these forces
have produced an astonishing diversity of forms and functions. The
forces that drive evolutionary and ecological processes do not occur on
a single scale---they emerge out of the interactions that occur across
all levels of spatial and temporal biological organization
(@levin\_problem\_1992).

Understanding how ecological processes change across space, within and
across scales, is central to the questions of ecology and evolution. The
spatial distribution of individuals is central to behavioral ecology,
the spatial distribution of species to biogeography, the spatial
distribution of genes to speciation, and so on. Further, the influence
of space on ecological processes has significant applied importance as
human activity has drastically altered the face of our planet in an
extremely short period of time (@bigelow\_major\_2017,
@vitousek\_human\_1997). Land-use change has shifted the amount and
configuration of Earth's habitats, and climate change is shifting the
spatial distribution of environmental factors and species' distributions
(@pereira\_scenarios\_2010, @thomas\_climate\_2010,
@loarie\_velocity\_2009). This drastic change in the structure of our
planet's terrain has had overwhelmingly negative effects on the
biodiversity to which it is home (@haddad\_habitat\_2015,
@rands\_biodiversity\_2010). This has, in turn, degraded ecosystem
functioning, and the services provided by ecosystem functioning, around
the globe (@tilman\_biodiversity\_2014, @cardinale\_biodiversity\_2012).

Historically, studies of the effects of landscape structure on
biodiversity have fallen under the broad umbrella of ``habitat
fragmentation'' (@collinge\_ecology\_2009). However, there has been
significant debate regarding what is meant by this term. The primary
distinction here is the difference between the word \emph{fragmentation}
as it is commonly used as a catchall to describe a combination of
habitat loss and ensuing subdivision, and \emph{fragmentation per se},
which is purely the spatial separation or reconfiguration of the same
amount of habitat (@fahrig\_effects\_2003). The effect of spatial
subdivision independent of habitat loss (fragmentation \emph{per se}) on
biodiversity is still hotly contested (@fahrig\_is\_2019,
@fletcher\_is\_2018, @fahrig\_ecological\_2017,
@ewers\_confounding\_2006). However, it is abundantly clear that the
combined effects of habitat loss and subdivision have had negative
effects on the planet's biodiversity (@haddad\_habitat\_2015). As a
result, a major focus on conservation efforts has become understanding
and predicting the simultaneous effects of habitat loss and subdivision
on ecosystem structure and stability (@fletcher\_spatial\_2019).

In order to understand and predict the consequences of landscape change,
we first need models to bridge the gap between the data we can collect
and the inferences and predictions we wish to make about the world.
Before we can construct quantitative models to make predictions with
real data, we first need to conceptualize the process we are modeling in
space. Over the years, ecologists have considered many different ways of
representing the structure of space and its influence on ecological
processes, using a variety of frameworks and tools to represent
different processes across different scales. Here, we examine the role
that model construction plays in scientific epistemology to better
understand how models can be used and compared, and review the various
constructions of space ecological models have used in the past.

\hypertarget{what-is-a-model}{%
\subsection{What is a Model?}\label{what-is-a-model}}

Science is fundamentally a theory of epistemology---a way of knowing.
Within the constraints of scientific epistemology, for a scientific
theory to be considered valid, it must be falsifiable through
observation, and for a theory to be disproven through observation, it
must make predictions that may or may not agree with observed reality
(@popper\_logic\_1934). In this view, a scientific model can be seen as
a function---some ontological object, \(f\), that maps measurable input
conditions, \(\vec{x}\), to a predicted outcome, \(y=f(\vec{x})\).

Constructing models is inherently a creative act. Before we can
introduce any quantitative representation of a process, we need to
conceptualize what it is we should even be quantifying. In \emph{What is
Philosophy?}, Deleuze \& Guattari (1991) answer the eponymous question
by suggesting philosophy is the act of concept creation. They write:

\begin{quote}
In fact, the sciences, arts, and philosophies are all equally
creators\ldots{} Concepts do not wait for us ready-made, like celestial
bodies. There is no heaven for concepts. They must be invented,
fabricated, or rather created, and would be nothing without the
signature of those who create them. (Deleuze \& Guattari 1991)
\end{quote}

In this view, scienctific modeling, too, is an act of creation, not
discovery. The concepts we use to frame natural systems are not inherent
properties of nature. Especially in ecology, it often seems the atomic
unit is the individual, yet this ignores differences in that
individual's cells, driven by biochemical processes that are the product
of interactions molecules within those cells, and so on---there is not
natural or correct scale on which to model natural processes
(@levin\_problem\_1992), and the conceptual frameworks we use to model
ecosystems are tools created by the modeler. Just as Newtonian Gravity,
General Relativity, and the Principle of Least-Action all describe and
predict the same phenomena using entirely different conceptual
frameworks (@arkani-hamed\_future\_2012, @feynman\_feynman\_1965), valid
ecological models can be framed in a variety of different ways.

A model's ``correctness'' can be measured as the difference between its
prediction, \(y\), and observed reality, \(\hat{y}\)
(@jaynes\_probability\_2003). However, all models have information baked
into them based on how they define \(f\), and how much information is
already contained in \(\vec{x}\) . Further, as we add more parameters to
a model it becomes harder to disprove that our predictions \(y\) don't
agree with \(\hat{y}\). To deal with this general problem of
overfitting, we turn to the theory of model selection. Model selection
is rooted in information theory---many common methods for model
selection---Minimum Description Length, AIC, BIC---revolve around the
huerstic that better models maximize that ratio of how much information
a model provides in predictive power to how much is built information
into the model structure (@jaynes\_probability\_2003,
@stine\_model\_2004). Other methods, like crossvalidation, aim to test
the predictive accuracy of a model by testing it on data it hasn't
``seen'' yet. (@konishi\_information\_2008). Either way, this provides
criteria by which to assess how models compare to one another in their
capacity to explain natural phenomena. However, what is actually meant
by `model' in practice varies widely. We use models for a variety of
things---inference, hypothesis-testing, prediction---and for these
different purposes we use a different methods to construct \(f\).

Breiman (2001) partitions models into two types: data models and
algorithmic models. Both map input conditions to output conditions, but
the difference lies is in the approach we take to create \(f\).
Algorithmic models aim to construct \(f\) as an algorithm which
accurately predicts output conditions based on input conditions. The
algorithm tends to have little relation to the mechanisms producing the
data being modeled, but rather is designed with the aim of accurately
finding patterns in data---methods of this type tend to receive the
somewhat ambiguous label ``machine learning''. These models can be
exceptionally useful for prediction, but this comes at the cost that
they do not reliably provide information about the mechanisms producing
the data.

In contrast, data models attempts to model the way in which the data is
generated. We can further partition data models into statistical models
and process models. The vast majority of scientific studies use
statistical models (@breiman\_statistical\_2001), which conceptualize
\(f\) by assuming the output of \(f\) is some combination of its input
conditions, \(x_i\), each of which is drawn from an assumed distribution
for each observation. Models of this form have seen extensive use
because they are easily generalizable and can be applied to a variety of
problems. Further, by assuming that measured quantities follow
distributions that we can describe analytically, we are able to compose
likelihood functions which can then be used to fit models to data using
a variety of methods, both frequentist and Bayesian. In contrast,
process models aim to construct \(f\) to represent the hypothesized
mechanism that produces the data. This can be done in a variety of
methods---for example, \(f\) could be written as a stochastic
differential equation, as we will do in the next chapter, or \(f\) could
be described as a set of properties between agents in an
individual-based model. Process models can vary in the degree to which
they represent the fidelity of the actual process based on the scale and
detail of representation considered. However, this comes with the cost
that, in general, process models are much more difficult to fit data to
than statistical models. However, recent advances in Approximate
Bayesian Computation (@beaumont\_approximate\_2010) make it feasible to
fit data to complex simulation models which cannot be described by an
analytic likelihood function. This can give us a way to test between
mechanistic process models, given that we have a reasonably small number
of potential mechanisms we want to test.

Now, we turn to the history of spatial models in ecology to survey the
the ways in which ecological process models have conceptualized space.

\hypertarget{a-history-of-the-construction-of-space-in-ecology}{%
\subsection{A History of the Construction of Space in
Ecology}\label{a-history-of-the-construction-of-space-in-ecology}}

\hypertarget{the-theory-of-island-biogeography}{%
\subsubsection{The Theory of Island
Biogeography}\label{the-theory-of-island-biogeography}}

Nearly all of landscape ecology has its roots in the Theory of Island
Biogeography (TIBG, @macarthur\_theory\_1967). The TIBG modeled a system
of oceanic islands, each with a different size, corresponding to
resource levels. The primary motivation of the theory was to provide a
mechanistic explanation of the species area relationship, and so the
core of the TIBG relates the spatial structure of the islandscape---the
size of each island and the distances between them---to observed
patterns of species richness.

Although the TIBG was conceptualized for terrestrial communities on
oceanic islands, it quickly was applied elsewhere under the assumption
that many human-altered landscapes are well-approximated by an island
structure---isolated regions of homogeneous landscape separated by
inhabitable matrix (@haila\_conceptual\_2002, @macarthur\_theory\_1967,
@preston\_canonical\_1962). The analogy between islands and subdivided,
patchy landscapes was first made by Preston (1962), and is briefly
invoked in the TIBG. However, this analogy went on to strongly influence
how landscape ecologists modeled space in terrestrial landscapes in the
future. Traces of the central ideas of the TIBG---area as a proxy for
resource availability, richness as a product of an equilibrium between
colonization and extinction, that patches are internally
homogenous---can be seen all throughout the models of landscape ecology
that followed.

\hypertarget{metapopulation-theory}{%
\subsubsection{Metapopulation Theory}\label{metapopulation-theory}}

\hypertarget{classical-metapopulation-theory}{%
\paragraph{Classical Metapopulation
Theory}\label{classical-metapopulation-theory}}

The metapopulation framework was introduced by Levins (1969)---who
originally defined a metapopulation as `a population of populations'.
Metapopulation theory, like the TIBG, considers the dynamics of
occupancy due to colonization or extinction. Levins' model considers a
system of infinite populations, each with a fixed probability of
colonization, \(c\), or local extinction, \(e\) , each generation. In
this formulation, the proportion of populations occupied at a given
time, \(p\), is modeled as as

\[\frac{dp}{dt} = cp(1-p)-ep\]

From this, we can derive that the system will persist, i.e.~\(p \to 1\),
if and only if \(\frac{c}{e} >1\), and \(p\to0\) otherwise
(@levins\_demographic\_1969). The aim of this model was not to be used
to predict the consequences of subdivision in an empirical
system---rather, this is what Okubo (1978) calls a ``toy model'': a way
of examining the consequences imposed by an oversimplified version of
the dynamics in question. In reality, stochasticity in
colonization/extinction events cause real systems to diverge from the
deterministic predictions made by Levins' model, and the spatially
implicit structure of this model made it difficult to assess the
influence of landscape structure or use for prediction in a real
landscape.

\hypertarget{modern-metapopulation-theory}{%
\paragraph{Modern Metapopulation
Theory}\label{modern-metapopulation-theory}}

Metapopulation theory was more formally applied to fragmented landscapes
by Hanski and Ovaskainen (@hanski\_practical\_1994,
@hanski\_metapopulation\_2000), who adapted Levins' framework of a
network of occupied/unoccupied patches, and restricted it to a finite
number of populations with spatially-explicit locations, each with a
unique probability of colonization arising from the metapopulation's
spatial structure. This introduces more parameters---each population
\(i\) has a spatial coordinate in \(\mathbb{R}^2\), \((x_i, y_i)\), and
an area associated with it, \(A_i\). Here, like in the TIBG, area is
still a proxy for the resource availability in that patch, as the
probability of a patch going extinct in a generation is is inversely
proportional to its area, and directly proportional to its ability to
serve as a source for colonizers. Similarly, in a departure from Levins'
model, the probability of an unoccupied patch getting colonized during
any timestep is inversely proportional to how far away it is from
occupied patches. This relationship of isolation-by-distance (IBD),
called the dispersal kernel, was initially modeled as an exponential
(@hanski\_practical\_1994), \(f(d_{ij})= \ e^{-\alpha d_{ij}}\), which
has seen extensive use since.

What made Hanski and Ovaskainen's theory capable of interfacing with
real data is their use of incidence-function models (IFMs,
@hanski\_practical\_1994). The value of the incidence function \(J_i\),
for any population \(i\), is the long run proportion of time that
population \(i\) is occupied. From this, one can compute the probability
of changing from any state of occupancy to any other state, which
provides us with a likelihood function to estimate parameters from
occupancy data in real systems. Much like the TIBG, this is based on the
assumption that when the data we collect is indicative of a system that
has approached its long-term equilibrium state. This, however, has
enabled the extensive use of IFMs in conservation
(@risk\_robust-design\_2011, @macpherson\_metapopulation\_2011,
@vos\_incidence\_2000, @hanski\_quantitative\_1996). More recent
metapopulation studies use so-called Dynamical Models
(@fletcher\_spatial\_2019), which model occupancy in space as a
Bernoulli distribution, with weighted parameters to account for
isolation-by-distance/resistance, environmental variance, or other
covariates.

\hypertarget{sources-and-sinks}{%
\paragraph{Sources and Sinks}\label{sources-and-sinks}}

Another theoretical output of metapopulation theory is the conceptual
framework of sources and sinks. Source-sink metapopulations emerged from
the theory of birth-immigration-death-emigration (BIDE) demographic
models (@pulliam\_sources\_1988), the idea being that different
locations in space have differing levels of resource availability, and
that as a result, have different expected values of each of these
parameters every generation. Populations with more births than deaths
and more emigrants than immigrants are `sources', populations with
higher deaths than births are `sinks'. This conceptual framework has
since spread outside population ecology, and gone on to be influential
across metacommunity (@mouquet\_community\_2003), population genetics
(@gaggiotti\_population\_1996), and coevolutionary theory
(@thompson\_coevolution\_2002).

Within both the metapopulation framework and the TIBG, it is assumed
that there isn't heterogeneity within patches, and that the region
between habitats is entirely inhospitable. However, these assumptions
would later be challenged by the increasing availability of data that
allowed ecologists to think about space continuously.

\hypertarget{continuous-space-lattice-models-and-the-matrix}{%
\subsubsection{Continuous Space, Lattice Models, and the
Matrix}\label{continuous-space-lattice-models-and-the-matrix}}

One natural limit of the above models is that ecological processes do
not occur in discrete space. Although continuous-space models have long
been used in theoretical studies (@haldane\_theory\_1948,
@hastings\_global\_1978, Okubo 1978), the difficulty inherent to both
collecting `continuous' spatial data, plus the increased mathematical
complexity induced by considering processes in continuous space, long
made such models impractical to apply to an empirical system.

This changed due to two major developments: the increased availability
of remotely sensed land cover data and the onset of computational power
to feasibly run simulation models that well-approximate continuous
space. The advent of remote sensing imagery enabled construction of
global land-cover maps, which function as approximations of continuous
space on large scales. In response, landscape research focused on
studying various aspects of patch geometry in raster data
(@mcgarigal\_fragstats\_1995).

This, plus the continued growth of computational power, paved the way
for models using rasters to approximate continuous space, meaning we
were no longer limited in our ability to model continuous space with the
tools of analytical math. Lattice-based simulation models have been
applied to many questions, including the theory of percolation in
habitat loss (@bascompte\_habitat\_1996), evolutionary rescue models
(@boeye\_more\_2013), epidemiology models (@jeltsch\_pattern\_1997),
coevolutionary models (@sisterson\_coevolution\_2004), adaptive
radiation (@gavrilets\_dynamic\_2005), predator-prey models
(@matsuda\_statistical\_1992), and so on.

These developments led us to more complex conceptual models of landscape
structure. The relationship of different land cover types in relation to
one another marks a significant shift away from the habitat is either
suitable-or-unsuitable binary that has dominated landscape ecology for
many years, and to other models including the so-called mosaic model
which considers all elements of the matrix as important to understanding
landscape structure (Fortin \& Fletcher 2019, Wiens 1995).

\hypertarget{resistance-connectivity-and-spatial-graphs}{%
\subsubsection{Resistance, Connectivity, and Spatial
Graphs}\label{resistance-connectivity-and-spatial-graphs}}

The field of landscape genetics, introduced by Manel et al.~(2003), has
sought to synthesize an understanding of landscape structure with
population genetic models using the recent availability of
next-generation sequencing technology. This field has contributed many
conceptual frameworks to the study of space, including idea of
isolation-by-resistance (IBR, @mcrae\_isolation\_2006,
@mcrae\_using\_2008). IBR argues that different landscape features,
typically represented as different land cover types, impose different
strengths of ``resistance'' to dispersal or movement. Each cell in a
raster landscape, called a resistance surface, is modeled as a resistor,
and the minimum resistance path between any two points in the raster can
be estimated using circuit theory (@spear\_use\_2010). More recently
models of gene flow have been used to validate resistance surfaces
(@mateo-sanchez\_estimating\_2015), and construct them from from a
combination of genetic and landscape data
(@peterman\_resistancega\_2018).

The concept of landscape connectivity has seen much recent interest in
conservation ecology (@robertson\_isolating\_2018,
@mitchell\_linking\_2013, @carroll\_use\_2012,
@krosby\_ecological\_2010). The effect of connectivity on ecological
processes has been clear since the concept was first introduced
(@turner\_landscape\_1989, @taylor\_connectivity\_1993), however, the
theory of resistance surfaces has proven an invaluable tool in
quantifying connectivity (@kool\_population\_2013). The use of spatial
graphs in connectivity models has also seen recent interest, as they
provide a generalizable framework to study pattern and process in
landscapes across different scales (@chubaty\_r\_2020,
@dale\_graphs\_2010, @minor\_graph-theory\_2008,
@urban\_landscape\_2001).

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

Here, we have focused on the various ways that spatial models have been
constructed in ecology---however, as mentioned earlier, scientists use
models to serve a number of different functions, and different ways of
modeling space have different utility in different circumstances. In the
next chapter, we restrict our focus to simulation models. We suggest
that when planning efforts to restore landscape connectivity, there is
immense value in using simulation models to predict the effects of
potential corridors on functional connectivity
(@epperson\_utility\_2010).

\pagebreak

### 2020-08-27 writing/ch2.tex
starting revisions

\hypertarget{phase-transitions-in-metapopulation-synchrony}{%
\section{Phase Transitions in Metapopulation
Synchrony}\label{phase-transitions-in-metapopulation-synchrony}}

\begin{quote}
\begin{flushright}Nature is a mutable cloud, which is always and never the same.

Ralph Waldo Emerson \end{flushright}
\end{quote}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Ecosystems are quintessential complex systems. Ecological processes are
inherently the product of interactions across all scales of biological
organization, from the interactions between electrons that drive
biochemical processes, to interactions between individual cells that
constitute multicellular life, to the interactions between separate
organisms in population ecology, to interactions between species in
community ecology, to interactions between biogeographical patterns and
biosphere level forces like climate (@levin\_problem\_1992). This
process of \emph{emergence}, by which parts come together to form a
whole with properties that don't exist among the individual parts, has
been studied across a wide variety of disciplines
(@manrubia\_emergence\_2004), and is a ubiquitous phenomenon in complex
systems.

One potential cause of emergent behavior in complex systems is
\emph{synchrony} among individual parts. When many independent parts
come together to act as a whole, their dynamics become
\emph{synchronized}. This behavior is ubiquitous in biological systems
across all scale of organization. From collections of cells acting
together: the heart beating in rhythm (@womelsdorf\_modulation\_2007),
neurons firing in unison (@strogatz\_sync\_2003)---to behavior among
organisms: the flash of fireflies (@otte\_theories\_1980) or the
migration of birds (@spottiswoode\_extrapair\_2004)---to interactions
between organisms: synchrony between abundances of predators and prey,
and of phenology (@van\_asch\_phenology\_2007, @burkle\_future\_2011).
Synchrony, by definition, involves different entities changing over time
in the same way. Within ecology, there has long been a focus on spatial
synchrony, that is---how does spatial distribution of ecological
entities affect whether they change together or separately?
(@jarillo\_spatial\_2020, @kendall\_dispersal\_2000,
@hanski\_spatial\_1993). This is, in large part, due to the applied
importance of understanding the effect of habitat loss on natural
populations. Many theoretical studies have shown the two primary factors
that develop spatial synchrony across space are dispersal and
environmental covariance (@ripa\_analysing\_2000, @abbott\_does\_2007).
Within this theory, one maxim that has developed is \emph{Moran's rule},
which states that spatial synchrony is proportional to the covariance in
the environmental conditions across space (@ranta\_synchrony\_1995,
@bjornstad\_spatial\_1999).

A major goal of conservation has been developing corridors to promote
landscape connectivity. Most measures of landscape connectivity
represent \emph{structural} connectivity, meaning quantifying the
structure of the landscape, rather than \emph{functional} connectivity,
which measures the connectivity of a given process
(@kool\_population\_2013, @calabrese\_comparison-shoppers\_2004). Here,
in order to better understand functional connectivity, we use a
simulation model to measure how synchrony across space changes as a
function of landscape structure. We do this by developing a model of
metapopulation dynamics on spatial graphs, which have long been used to
model landscape connectivity (@martensen\_spatio-temporal\_2017,
@albert\_applying\_2017, @urban\_landscape\_2001), and analyze how
synchrony changes across space using the language of critical
transitions.

Further, we show that increasing population synchrony reduces the
variance in the generation-to-generation change in abundance, which is
central in reducing the probability of metapopulation extinction
(@lande\_risks\_1993, @lande\_extinction\_1998). This relationship
suggests that promoting functional landscape connectivity can help
mediate the probability of extinction for species facing significant
habitat loss. We suggest using simulation models, such as those
presented here, to aid in decision making regarding corridor placement.

\hypertarget{what-are-phase-transitions}{%
\subsection{What are phase
transitions?}\label{what-are-phase-transitions}}

When does a system change from one state to a different state? Due to
the rapid changes induced on the planet by human activity, there has
been recent focus on answering this question in ecology, especially the
potential for changes in spatial structure to drive transitions between
alternative stable states. Much of this theory has been aimed at the
practical problem of being able to predicting the onset of transitions
from time-series data (@scheffer\_anticipating\_2012,
@scheffer\_early-warning\_2009). The bulk of the quantitative theory
used to understand transitions between states is derived from
statistical mechanics, where it was originally used to study
phase-transitions in matter. In order to study phase transitions between
regimes, we must first be clear on what these regimes are. As this
theory was originally used to describe physical states of matter, the
original regimes were solid, liquid, gas. However, as our understanding
of condensed-matter has changed, so has the demarcation of what
constitutes different states. In reality, the way in which particles
come together to constitute matter is far more variable than these three
categories. In such cases the `state' of a collection of particles
cannot be represented by a single categorical label, and so the theory
of phase transitions as been adapted to model \emph{continuous} phase
transitions, where there is no clear demarcation point between different
states (@sethna\_statistical\_2006). This is useful for us in ecology,
where the line between alternative ecosystem `states' is even fuzzier.

We can formalize our understanding of phase transitions using the
language of statistical mechanics. We call the \emph{order parameter}
some measure of the system's state in space and time. The \emph{control
parameter}, then, is what causes the change in order parameter. When
dealing with dynamics that are inherently stochastic, one tool often
used in statistical mechanics is correlation functions which measure how
well the order parameter is correlated in both space and time at a
particular value of the control parameter (@sethna\_statistical\_2006).
For example, if we consider the population of a species inhabiting a
landscape, where along the gradient of landscape connectivity, our
control parameter, does that system go from consisting of one large,
single, population, to many small, independent populations? We measure
this qualitative shift from one system to many using \emph{synchrony},
the correlation in the dynamics of abundance across space.

\pagebreak

\hypertarget{the-model}{%
\subsection{The Model}\label{the-model}}

Here we present a spatial graph model of landscape connectivity based on
metapopulation theory (@hanski\_practical\_1994,
@grilli\_metapopulation\_2015) . We model connectivity as a function of
a few empirically estimable parameters, and then describe a diffusion
model of metapopulation dynamics on these spatial graphs. We then
simulate dynamics across parameters representative of landscape
connectivity to determine where transitions in the synchrony of
population dynamics across space occur.

\hypertarget{modeling-landscape-connectivity-with-a-spatial-graph}{%
\subsubsection{Modeling Landscape Connectivity with a Spatial
Graph}\label{modeling-landscape-connectivity-with-a-spatial-graph}}

Spatial graphs have long been used to model a system of habitat patches
(@dale\_graphs\_2010, @minor\_graph-theory\_2008,
@urban\_landscape\_2001). In figure \ref{concept}, we see a conceptual
example of how these graphs are constructed from land cover data.

\begin{figure}[h]

\includegraphics[width=15cm]{/Users/michael/Drive/thesis/figures/concept.png}

\caption{Spatial graph representation of a heterogenous landscape. Panel (a): A raster map of landscape classification, with each land cover type denoted in a different color. Panel (b): The habitat type of interest (green), repreented as a set of contiguous patches. Panel (c): Each patch represented as a node at the patch's centroid. Panel (d): A spatial graph derived from the landscape. }

\label{concept}

\end{figure}

Here we model a system of populations, represented as a vector of
vertices, \(V\), in a spatial graph \(G=(V,E)\), with edges, \(E\),
representing dispersal between populations. When considering the process
of metapopulation dynamics, we choose to model landscape connectivity as
a combination of two different factors: the probability than any
individual migrates during its lifetime, \(p_m\), and the conditional
distribution over spatial nodes of where an individual goes, given that
it migrates, \(P(V_j|V_i)\), which we call the dispersal potential.

We can model the dispersal potential using a few methods. In empirical
systems, this can be estimated with resistance surfaces, which provide
relative weights of the difficulty of migration between points on a
raster of land-cover type (@spear\_use\_2010). Theoretically, we model
the dispersal potential using isolation-by-distance (IBD). The relative
probability of dispersal between \(V_i\) to \(V_j\) is inversely
proportional to the distance between them, \(d_{ij}\), and the strength
of the isolation-by-distance relationship, \(\alpha\). We call the
functional form of this relationship, \(f(d_{ij}, \alpha)\), the
dispersal kernel. Here we consider two different types (partially after
Grilli et al.~(2015)), the
exponential, \(f(d_{ij}, \alpha)=e^{-\alpha d_{ij}}\), and
Gaussian, \(f(d_{ij}, \alpha)=e^{-\alpha^2 d_{ij}^2}\), kernels. These
functional forms have long been considered as models of dispersal
kernels in both theory and empirical work (@hanski\_practical\_1994,
@grilli\_metapopulation\_2015).

To convert this to a probability distribution \(P(V_j|V_i)\), we
normalize:

\[P(V_j|V_i)=\frac{f(d_{ij}, \alpha)}{\sum_k f(d_{ik},\alpha)}\]

Note that if \(\alpha=0\), then the value of both exponential and
Gaussian kernels is the same for all pairs of populations, and therefore
the dispersal potential is a uniform distribution. In figure \ref{mp},
we can see spatial graphs plotted representing the same set of
populations across differing values.

\begin{figure}[h]

\includegraphics[width=15cm]{/Users/michael/Drive/thesis/figures/mp.png}

\caption{The same set of populations plotted in the x-y plane, represented as blue points, across various value of isolation-by-distance strength, represented by (columns), and dispersal kernel type (rows). Edge width and opacity are proportional to the dispersal kernel between those two populations, $P(V_j | V_i)$.}

\label{mp}

\end{figure}

\hypertarget{local-dynamics}{%
\subsubsection{Local Dynamics}\label{local-dynamics}}

We model population dynamics within each local population \(V_i\) using
the stochastic logistic model. The dynamics of the number of individuals
in population \(V_i\) are described by the stochastic differential
equation (SDE)

\[dN_i=\lambda_i N_i \big(1 - \frac{N_i}{K_i}\big)dt+\sigma_pK_idW\]

Here, \(N_i\) is the abundance at population \(i\), \(\lambda_i\) is the
strength of density dependence in that population, and \(K_i\) is the
carrying capacity of \(V_i\). \(\sigma_p\) represents that standard
deviation in abundance due to local stochasticity as a proportion of
\(K_i\). Here \(\sigma_p\) represents an amalgamation of all factors
contributing to local stochasticity in population dynamics, although it
should be noted that the relative contribution of different factors to
local stochasticity can drive significant variation in dynamics
(@melbourne\_extinction\_2008). For the sake of reducing parameter
space, here we consider all populations as having the same \(\lambda_i\)
and \(K_i\), however, future work could include exploring the
source-sink dynamics in this system by varying intrinsic growth rates
and carrying capacities across populations.

We use an SDE representation because they have been used to study phase
transitions in stochastic systems before (@brock\_early\_2012,
@kuehn\_mathematical\_2011), and they have many nice properties. SDEs
have been used to study extinction dynamics before, as it is relatively
straightforward to compute the mean time until extinction (MTE) using
the Kolmogorov Backward Equation (@lande\_stochastic\_2003).

\hypertarget{diffusion-on-spatial-graphs}{%
\subsubsection{Diffusion on Spatial
Graphs}\label{diffusion-on-spatial-graphs}}

When we model a landscape with a spatial graph, we have to decide how
different nodes affect one another. The processes that connect
landscapes are inherently stochastic. The probability that an individual
migrates within its lifetime, \(p_m\), and where it goes, \(P(V_j|V_i)\)
, are both stochastic processes. Under some conditions, we can
effectively model stochastic dynamics across space using diffusion
models. In essence, a diffusion model assumes that at each timestep the
system will change according to the expected value of stochastic
dispersal. Diffusion models have seen widespread use in ecology and
other fields (@ovaskainen\_empirical\_2008,
@holmes\_partial-differential\_1994, @okubo\_diffusion\_2011).

Given our dispersal potential \(P(V_j|V_i)\), we can define the
diffusion matrix \(\Phi\) as

\[\Phi_{ij} = \begin{cases} 1 - p_m \quad\quad &\text{if}\ i = j \\ P(V_j|V_i)p_m & \text{if}\ i \neq j\end{cases}\]

where \(\Phi_{ij}\) represents the probability that any individual born
in \(V_i\) reproduces in \(V_j\).

Now we move to considering the dynamics of the abundance of the
population at \(V_i\), denoted \(N_i\). We can now represent the
dynamics due to diffusion of this system as

\[\dot{N_i}=(1-p_m)N_i+ \sum_j p_mP(V_j|V_i)N_j\]

In matrix notation, we can represent this diffusion model as

\[\frac{d\vec{N}}{dt}=\Phi^T\vec{N}\]

We can then combine this with local dynamics as a reaction-diffusion
model,

\[\frac{d\vec{N}}{dt} = g(\Phi^T \vec{N})\]

where \(g(x)\) is a function that represents the hypothesized mechanism
of how the ecological measurement evolves locally.

In principle, \(g(x)\) can represent any ecological process of
interest---for example if the state space of \(x\) is allelic
frequencies, \(g(x)\) could describe genetic drift, or if \(x\)
represents community compositions across space, \(g(x)\) could describe
competition between species as a function of environmental conditions,
coevolutionary states across space, etc. Here, we consider \(g(x)\) to
be the stochastic logistic model (see previous section). Combining this
with the diffusion model yields the SDE

\begin{equation} \label{diffusion_model}
  d \vec{N} = {I} (\Phi^T \vec{N})(\vec{k}-\vec{N}) dt + \sigma_p \vec{k} \ d\vec{W}
\end{equation}

which will be the primary object of study in this paper.

\hypertarget{measuring-synchrony}{%
\subsubsection{Measuring Synchrony}\label{measuring-synchrony}}

As described above, measures of correlation in space and time are often
used in the study of phase transitions. When the order parameter gains
or loses correlation in space and time is used as an indicator of when a
system changes qualitative phases. Within the context of ecology, the
crosscorrelation function, \(CC\), has long been used as a measure of
synchronous dynamics (@liebhold\_spatial\_2004). Here, with a subdivided
population, we consider the mean crosscorrelation compared across all
populations, \({PCC}=\frac{1}{n^2}\sum_{i,j} CC(V_i,V_j)\) . As an
example, in \ref{async_and_sync}, panel (a) show an example of low PCC,
and panel (b) shows an example of high PCC.

\begin{figure}[h]

\includegraphics[width=15cm]{/Users/michael/Drive/thesis/figures/async_and_sync.png}

\caption{The abundance of 5 populations (each in a different color) across time. Panel (a):  These populations show low synchrony, $PCC=0.1$. Panel (b): These populations show high sychrony, $PCC=0.89$ }

\label{async_and_sync}

\end{figure}

\hypertarget{simulating-a-phase-transition-across-a-migration-gradient}{%
\subsubsection{Simulating a Phase Transition across a Migration
Gradient}\label{simulating-a-phase-transition-across-a-migration-gradient}}

Now we consider how synchrony changes as a function of \(p_m\). Each run
consists of the following parameters,
\(\theta = \{N_p, \lambda, \sigma_p, \vec{K},p_m, \alpha \}\). For each
unique set of parameters, we run \(50\) replicates across all values of
\(p_m = \{0.01,0.02,\dots,1.0 \}\). For each replicate, we independently
draw the location of \(N_p\) populations uniformly in \([0,1]^2\). We
draw the intial value of abundance for all populations from a uniform
distribution, \(N_i \sim U(0, K_i)\). We then integrate equation
\ref{diffusion_model} forward \(500\) timesteps using the
Euler--Maruyama method with \(\Delta t=0.1\). After integrating forward,
we compute the crosscorrelation coefficient, \(CC\) for each pair of
populations \(i \neq j\). Then, we compute the mean pairwise
crosscorrelation for that replicate, \(PCC\), and start the next
replicate.

\pagebreak

\hypertarget{phase-transitions-in-synchrony-across-a-migration-gradient}{%
\subsection{Phase Transitions in Synchrony across a Migration
Gradient}\label{phase-transitions-in-synchrony-across-a-migration-gradient}}

\hypertarget{synchrony-phase-diagrams}{%
\subsubsection{Synchrony Phase
Diagrams}\label{synchrony-phase-diagrams}}

Here, we introduce synchrony transition diagrams. When does the
equilibrium state of our order parameter, \(\text{PCC}\), change as we
change our control parameter, \(p_m\)? First, we consider the case of
only two populations, \(N_p=2\). Here, the dispersal potential,
\(P(V_j|V_i)\), becomes irrelevant, as the fact that an individual
migrates means that it migrates to the other
population---i.e.~\(P(L_2|L_1)=1\) and \(P(L_1|L_2)=1\). Here, then,
\(p_m\) is the only parameter driving connectivity in the system. How
does the expected amount of synchrony between the two populations change
as we change the probability an individual migrates, \(p_m\)? In Figure
\ref{2pops_across_lambda}, we see \(\text{PCC}\) plotted as a function
of \(p_m\). Each point represents the mean value of \(\text{PCC}\)
across 50 replicates. Each panel of figure \ref{2pops_across_lambda} is
for different strengths of density-dependent regulation, \(\lambda\).

\begin{figure}[h]

\includegraphics[width=15cm]{/Users/michael/Drive/thesis/figures/2pops_across_lambda.png}

\caption{Each panel is a synchrony transition diagram for differing values of density-dependent regulation, $\lambda$. Each panel considers how mean $PCC$ changes as a function of the probability of migration, $p_m$. Each point represents the mean value of $PCC$ across 50 replicates at that parameter set.}

\label{2pops_across_lambda}

\end{figure}

Each plot in figure \ref{2pops_across_lambda} is a phase transition
diagram, showing how synchrony changes as a function of migration
probability. Each show a similar functional form, as we see correlation
between the two populations maximized near \(p_m=0.5\). Intuitively,
this maximum makes sense: exchanging 50\% of the individuals between
populations in each time step leads to a well-mixed metapopulation.
Furthermore, since there is no local adaptation in this model, a
migration rate of \(p\) is functionally equivalent to a migration rate
of \(1 - p\).

We see that this is a continuous, or type-II phase transition---we don't
reach a point of \(p_m\) where the system suddenly jumps from
\(\text{PCC}= 0\) to \(\text{PCC}=1\). Instead, we see that the expected
level of synchrony drops off as we move away from \(p_m=0.5\). Further,
as we increase \(\lambda\), we see this peak become `narrower', in that
values near \(p_m = 0.5\) have lower expected \(PCC\) as \(\lambda\)
gets larger. Why does increasing the strength of density dependent
regulation, \(\lambda\), narrow the curve? As \(\lambda\) increases,
values of \(p_m\) that enable in-between levels of synchrony for low
\(\lambda\) have their moderate synchrony broken down by local
density-dependent regulation, unless the populations are at a value of
\(p_m\) such that they synchronized tightly enough that they are both
experiencing the same force of density-dependent regulation, meaning
moving either up or down towards \(K_i\), at the same time.

In figure \ref{2pops_across_sigma}, we see transition diagrams for
\(N_p=2\) across across demographic stochasticity, \(\sigma_p\) .

\begin{figure}[h]

\includegraphics[width=15cm]{/Users/michael/Drive/thesis/figures/2pops_across_sigma.png}

\caption{Each panel is a synchrony transition diagram for differing values of density-dependent regulation, $\sigma$. Each panel considers how mean $PCC$ changes as a function of the probability of migration, $p_m$. Each point represents the mean value of $PCC$ across 50 replicates at that parameter set.}

\label{2pops_across_sigma}

\end{figure}

Here, we see that amount of demographic stochasticity doesn't
dramatically alter the shape of the synchrony transition. This leaves us
with a few questions---why would this phase transition by centered
around \(p_m = 0.5\)? Here we suggest the following hypothesis: at
\(p_m=0.5\), the composition of each population at any time is equally
composed of descendants from each population---this enables even-mixing
of demographic stochasticity within each population, which maximizes
synchrony. We make this case further in \emph{Synchrony due to
Even-Mixing of Local Stochasticity}. Why does the strength of
density-dependence, \(\lambda\), narrow the region of \(p_m\) where
synchrony is maximized, but \(\sigma_p\) doesn't? We suggest that this
is due to the small amount of subdivision present, which we'll examine
in the next section.

\hypertarget{effects-of-demography-and-landscape-connectivity}{%
\subsubsection{Effects of Demography and Landscape
Connectivity}\label{effects-of-demography-and-landscape-connectivity}}

Now, we vary demographic parameters, \(N_p\), \(K\), \(\sigma_p\) , and
\(\lambda\) , to determine how shifts in demography change the
functional form of the transition across \(p_m\). In figure
\ref{subdiv_across_lambda}, we vary the strength of density dependence,
\(\lambda\), across different numbers of populations, \(N_p\).

\begin{figure}[h]

\includegraphics[width=15cm]{/Users/michael/Drive/thesis/figures/subdiv_across_lambda.png}

\caption{Each panel is a transition diagram showing $PCC$ across $p_m$ at different combinations of $\lambda$ and $N_p$.}

\label{subdiv_across_lambda}

\end{figure}

Here we see that both \(\lambda\) and \(N_p\) have effects on the
structure of the curve. Like in the two-population model, increasing
\(\lambda\) narrows the peak around which \(p_m\) values produce
synchrony. Let \(m_{max}\) be the value of migration that maximizes
synchrony. We see that as subdivision, \(N_p\), increases, we see the
transition curve narrows around increasingly high values of \(m_{max}\).
This supports with our hypothesis of even-mixing, which we'll explore
more later. In figure \ref{subdiv_across_lambda}, each color represents
a different value of isolation-by-distance strength, \(\alpha\). We see
that changing \(\alpha\) doesn't change the shape of the curve, but
instead just flattens it---decreasing the maximum value of synchrony
across the \(p_m\) gradient. This is intuitive---as connectivity
decreases, it becomes less likely to produce global synchrony at any
\(p_m\) .

Next we varied the total carrying capacity, \(K\), across different
numbers of populations, \(n_p\) (Figure \ref{subdiv_across_k}).

\begin{figure}[h]

\includegraphics[width=15cm]{/Users/michael/Drive/thesis/figures/subdiv_across_k.png}

\caption{Each panel is a transition diagram showing $PCC$ across $p_m$ at different combinations of $K$ and $N_p$.}\label{subdiv_across_k}

\end{figure}

In Figure \ref{subdiv_across_k}, the number of populations has the same
effect as above, but carrying capacity doesn't alter the shape of this
curve, which is unsurprising given that our model considers the variance
of local stochasticity to be a proportion of \(K\).

Finally, we consider shifting \(\sigma_p\). (Figure
\ref{subdiv_across_sigma}). Here, we see that the magnitude of
\(\sigma_p\) narrows the curve, as larger amounts of local stochasticity
drown-out synchrony at values near, but not at, \(m_{max}\). Here we
observe the same effect of subdivision, and similar narrowing around
\(m_{max}\) . Unlike in the two-population model, more variability
locally does mean less synchrony at same migration levels. We suggest
that this is because increasing subdivision increases the number of
stochastic demographic events that happen every timestep, which
inherently increases the variance in the change in abundances during any
interval.

\begin{figure}[h]

\includegraphics[width=15cm]{/Users/michael/Drive/thesis/figures/subdiv_across_sigma.png}

\caption{Each panel is a transition diagram showing $PCC$ across $p_m$ at different combinations of $\sigma_p$ and $N_p$.}\label{subdiv_across_sigma}

\end{figure}

Clearly, we see landscape subdivision, represented by \(N_p\), as having
the most prominent effect on the shape of the transition curve. The more
landscape subdivision there is, the higher migration has to be to enable
synchrony. Further, subdivision induces a qualitative shift in the shape
of this curve, from being only convex, (e.g., figure
\ref{subdiv_across_sigma}A), to having both concave and convex regions
(e.g., figure \ref{subdiv_across_sigma}D). As subdivision increases, the
transition curves exhibit more `critical' behavior, meaning they reach a
tipping point value of \(p_m\) where synchrony starts to increase very
suddenly (figure \ref{subdiv_across_sigma}(A-D)).

We next considered how \(\text{PCC}\) changes across both \(p_m\) and
\(\alpha\) with different dispersal kernels. Figure \ref{kernels} shows
that the dispersal kernel has significant effects on the region where
global synchrony is established. This means that different structures of
resistance features lead to different outcomes in the ability to build
up global synchrony.

\begin{figure}[h]

\includegraphics[width=15cm]{/Users/michael/Drive/thesis/figures/kernels.png}

\caption{Phase digram showing $PCC$ across $\alpha$  (x-axis) and $p_m$ (y-axis) for the exponential dispersal kernel (left) and gaussian dispersal kernel (right). }

\label{kernels}

\end{figure}

\hypertarget{synchrony-due-to-even-mixing-of-local-stochasticity}{%
\subsection{Synchrony due to Even-Mixing of Local
Stochasticity}\label{synchrony-due-to-even-mixing-of-local-stochasticity}}

The Moran effect claims that the synchrony of populations across space
in proportional to the covariance of their environmental conditions
(@ranta\_synchrony\_1995, @bjornstad\_spatial\_1999). Here we make the
case that what drives synchrony across space is the `even-mixing' of
local stochasticity. As is evidenced by figures \ref{2pops_across_sigma}
and \ref{2pops_across_lambda}, in a two population scenario, synchrony
is maximized near \(p_m =0.5\). We suggest that this is because the
population at any given time step is an even mix of descendants from
each population, and thereby averages out the demographic stochasticity
across all populations, leading to synchronized dynamics, as the change
in each population at every time is the averaged value of local change
across all populations. If this were the case, we would expect the value
of migration which maximizes synchrony,
\(m_{max}=arg max_{p_{m}} PCC(p_m)\) to be the value such that the
probability of an individual reproduces in population \(j\) given they
were born in population \(i\), \(\Phi_{ij}\), to be equal for all \(j\).
In the case that \(P(V_j | V_i)\) is uniform, i.e.~\(\alpha = 0\), this
would mean setting \(p_m = m_{max}=1 − \frac{1}{N_p}\) would result in a
\(\Phi_{ij}\) that evenly mixes indivudals each timestep. In figure
\ref{even_mixing}, we can see this function is plotted in black against
simulated data points (purple).

\begin{figure}[h]

\includegraphics[width=15cm]{/Users/michael/Drive/thesis/figures/even_mixing.png}

\caption{Each panel represents number of populations, $N_p$ on the x-axis, and $p_m$  on the y-axis. The value of $p_m$ which maximizes syncrhony across a $p_m$ gradient, with $50$ replicates at each value of $p_m$, is represented in purple. The line in black represents even mixing,  $m_{max}=1 - \frac{1}{N_p}$. Each panel represents a different strength of isolation-by-distance, $\alpha$.}

\label{even_mixing}

\end{figure}

This indicates that synchrony is enabled when local stochasticity is
averaged out by dispersal. The Moran effect suggests that population
synchrony across space is proportional to covariance in environment.
However, dispersal can substitute for covariance in the environment when
it results in the admixture of individuals from populations in
proportion to the environmental covariance. Further, we see in
\ref{even_mixing} that increasing the strength of isolation-by-distance
this relationship weaker, as different subsets of the total set of
populations form synchronized dynamics (see the final section,
\emph{Mesopopulation}).

\hypertarget{synchrony-and-variance}{%
\subsubsection{Synchrony and Variance}\label{synchrony-and-variance}}

Consider the change in size of population \(i\) size across all
timesteps, \(\Delta_i = [\Delta_i(t) - \Delta_i(t-1)] \ \forall \ t\).
How does the global synchrony of the metapopulation change the variance
of this distribution?

In figure \ref{variance_over_m}, we see how the variance of \(\Delta\),
\(V(\Delta)\), changes with \(p_m\). We see that the overall variance in
the change in abundance over time is reduced as \(p_m\) increases, in a
pattern strikingly similar to the way \(PCC\) changes over \(p_m\). In
panel (b), we see that \(V(\Delta)\) is negatively correlated with
\(PCC\), and that this relationship becomes stronger as \(\sigma_p\)
increases. This indicates that synchrony reduces the values of
\(V(\Delta)\). Now consider the value of \(p_m\) which minimizes
\(V(\Delta)\). How correlated is this with the value of \(p_m\) that
maximizes synchrony, \(m_{crit}\) ? Consider figure
\ref{delta_pcc_corr}.

\begin{figure}[h]

\includegraphics[width=15cm]{/Users/michael/Drive/thesis/figures/variance_over_m.png}

\caption{Panel (a): The variance in the change in abundance at each timestep, $V\Delta$, plotted across $p_m$. Each point represents $50$ replicates at that parameter set. Different colors indicate different levels of demographic stochasticity, $\sigma_p$. Panel (b): The value of $V(\Delta)$ (y-axis) and $PCC$ (x-axis) for each set of $50$ replicates. Different colors indicate different levels of demographic stochasticity, $\sigma_p$.  }

\label{variance_over_m}

\end{figure}

\begin{figure}[h]

\includegraphics[width=15cm]{/Users/michael/Drive/thesis/figures/delta_pcc_corr.png}

\caption{The value of $p_m$ which minimizes the variance of $\Delta$, $V(\Delta)$ (x-axis) and the value of $p_m$ which maximizes $PCC$, (y-axis). Each point represents the value of $p_m$ which maximizes synchrony, $m_{max}$,  across $p_m = \{0.01, 0.02, \dots, 1\}$ , at different $N_p=2,3,\dots,25$. The black line is $y=x$.}

\label{delta_pcc_corr}

\end{figure}

We know from theory that reducing that variance of \(N_t - N_{t-1}\)
reduces extinction risk (@lande\_stochastic\_2003,
@melbourne\_extinction\_2008). It has been suggested in the past the
spatial synchrony can increase the risk of extinction
(@matter\_local\_2010, @heino\_synchronous\_1997). However, our evidence
here runs counter to that claim---increasing synchrony decreases the
variance of the change in abundance over time, \(V(\Delta)\), which is
key in maximizing the probability of a populations persistence
(@lande\_stochastic\_2003). We suggest that prior studies are flawed in
that they do not consider that the advent of synchrony reduces the
probability of catastrophes, or high variance in \(\Delta\).

\pagebreak

\hypertarget{the-balance-of-migration-and-demographic-stochasticity}{%
\subsection{The Balance of Migration and Demographic
Stochasticity}\label{the-balance-of-migration-and-demographic-stochasticity}}

The question remains---does this diffusion model accurately describe the
actual dynamics of dispersal in ecological systems? Dispersal is a
stochastic process, and diffusion assumes that each time step, the
dispersal between pops in the expected value of that process. Here we
present a method for using our representation of landscape connectivity
to simulate stochastic dispersal, and compare the stochastic dispersal
model to the diffusion model.

\hypertarget{stochastic-dispersal-model}{%
\subsubsection{Stochastic Dispersal
Model}\label{stochastic-dispersal-model}}

We use the same construction of landscape connectivity in the random
dispersal model as the diffusion model---that is that each individual
migrates from its original population with probability \(P(m)\), and the
conditional distribution of where migrations go, \(P(V_j|V_i)\).
However, rather than using fixed coefficients, \(\Phi_{ij}\), to
represent dispersal, each generation we draw the number of total
migrations from population \(i\), \(N_{mi}\), from a Binomial
distribution, \(N_{mi} \sim \text{Binomial}(N_i,p_m)\). Then, for each
migrant \(\{1,2,N_{mi}\}\), the population \(j\) the migrant goes is
drawn directly from the distribution \(P(V_j|V_i)\)

\hypertarget{diffusion-vs.-stochastic-model}{%
\subsubsection{Diffusion vs.~Stochastic
Model}\label{diffusion-vs.-stochastic-model}}

First we consider the difference in synchrony transitions across a
migration gradient between using stochastic migration (\emph{Stochastic
Dispersal Model} above) vs.~the diffusion model (equation
\ref{diffusion_model}). Figure \ref{stoch_vs_determ_grid} shows phase
transition curves for both the diffusion (dark blue) and stochastic
(light blue) models across different values of \(N_p\) and \(\sigma_p\).

\begin{figure}[h]

\includegraphics[width=15cm]{/Users/michael/Drive/thesis/figures/stoch_vs_determ_grid.png}

\caption{Each panel shows a phase transition curves across $p_m$. The top row has $N_p=2$ and the bottom row has $N_p=5$. Columns correspond to different values of demographic stochasticity, $\sigma_p$. }

\label{stoch_vs_determ_grid}

\end{figure}

We see that as of demographic stochasticity, \(\sigma_p\), increases,
the difference in PCC between the deterministic model and the stochastic
model decreases. This is evident of a transition in which stochastic
mechanism, demographic stochasticity or dispersal stochasticity, is
driving the variation between populations at any fixed time.

In cases where the amount of demographic stochasticity is low (Figure
\ref{stoch_vs_determ_grid}A), migration events are the primary source of
the variability in the size of each population over time. When we model
dispersal deterministically, there is no variability in migration each
timestep. So, there is more correlation across populations in the
deterministic than the stochastic model, where migration events occur
according to Monte Carlo sampling. However, as we increase the amount of
demographic stochasticity, \(\sigma_p\), we reach a point where the
variability in the abundance of each population over time is dominated
by demographic stochasticity within populations, as opposed to migration
events. When demographic stochasticity is the primary source of
variance, migration between populations, even if it is not distributed
exactly according to the dispersal potential at each time point, spreads
this stochasticity out across all populations. When migration
stochasticity is the primary source of variance, the abundances at each
population \(i\) are stable, except for the variance induced by
immigration and emigration events.

Figure \ref{stoch_vs_determ_surface} show that beyond some threshold
level of demographic stochasticity, both the deterministic and
stochastic dispersal model can be used to understand the effects of
dispersal and landscape structure on correlation between dynamics as the
temporal variance that would `break down' correlations between
abundances is dominated by demography rather than the stochastic nature
of migration. We can better understand this transitions using the phase
diagram in Figure \ref{stoch_vs_determ_surface}b.

\begin{figure}[h]

\includegraphics[width=15cm]{/Users/michael/Drive/thesis/figures/stoch_vs_determ_surface.png}

\caption{(a): Phase transition diagram across $\sigma_p$ (x-axis) and $p_m$ (y-axis). Each cell represents the difference in mean PCC across $200$ replicates of both the stochastic dispersal and diffusion model. Labeled contour lines are in white. (b): The same phase transition diagram as (a), with regimes labeled. Note the logarithmic scale on the x-axes on both panels.}

\label{stoch_vs_determ_surface}

\end{figure}

\pagebreak

Figure \ref{stoch_vs_determ_surface} shows different two regimes---one
where the stochastic and diffusion model make the same predictions, and
one where they don't, highlighted in figure
\ref{stoch_vs_determ_surface}b. This indicates that as long as
demogaphic stochasticity is larger than \(\sigma_p = 0.05\), the
diffusion model and stochastic dispersal model make the same predictions
about metapopulation synchrony. In the migration dominated regime,
building up high synchrony isn't possible at any \(p_m\). However, in
the `demographic stochasticity' region, the that stochastic migration
and diffusion models make the same prediction, and develop the synchrony
transitions we see in the previous section.

\hypertarget{the-mesopopulation}{%
\subsection{The Mesopopulation}\label{the-mesopopulation}}

What does it mean biologically that this is a continuous phase
transition? When we increase the strength of isolation-by-distance, the
system does not collapse from one unifed whole to \(N_p\)
subdivisions---instead, synchronous dynamics form within subsets of the
total set of populations. In figure \ref{mesopop} we see the same
metapopulation plotted across varying values of isolation-by-distance,
\(\alpha\).

\begin{figure}[H]

\includegraphics[width=15cm]{/Users/michael/Drive/thesis/figures/mesopop.png}

\caption{Each panel (top, middle, bottom) represents the same set of $N_p=5$ populations. On the left, the population is plotted in the $x-y$ plane, with edge opacity and width corresponding the value of the dispersal potential $P(V_j | V_i)$. On the right is a set of pairwise correlation coefficient plots. Here, each point represents the mean value of crosscorrelation between those specific populations across $50$ replicates. Each panel represents different values of isolation-by-distance, $\alpha$.  }

\label{mesopop}

\end{figure}

At \(\alpha=0\), where dispersal is uniform, each pair of populations
shows a synchrony transition diagram that is identical, and we could
qualitatively say that this system functions as a whole.

As we increase \(\alpha\), (figure \ref{mesopop}b) we see that
populations \(4\) and \(5\) form a system between themselves, where
synchrony between those two populations is maximized near \(p_m=0.5\),
while the synchrony values between populations \(4\) and \(5\) with
populations \(1\), \(2\), and \(3\) moves toward \(PCC =0\). Meanwhile,
populations \(1\), \(2\), and \(3\) form their own system with
synchronous dynamics maximized at \(p_m = 1-\frac{1}{3}= \frac{2}{3}\),
as we would expect in a \(3\)-population system.

At \(\alpha = 15\), figure \ref{mesopop}b, these dynamics solidify, and
populations \(4\) and \(5\) form a single, dual-synchronized system, and
\(1\), \(2\) and \(3\) form a single \(3\)-population system.
Understanding how shifts in landscape connectivity cause whole system of
populations to break down into several mesopopulations, systems of that
form a unity through synchronized dynamics within the whole, has
important applied context. In some conservation scenarios, developing
corridors to connect these clusters could have significant benefits to
biodiversity and ecosystem functioning.

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

Here, we present a way of measuring of \emph{functional connectivity}
using simulation models. We construct a model of diffusion that relates
to measurable values of resistance in real landscapes, and show that
this diffusion model can make accurate predictions about stochastic
dispersal at sufficiently high values of demographic stochasticity,
\(\sigma_p\).

We emphasize the importance of these types of simulation models in
measuring \emph{functional} connectivity. Given the constraints that
exist on conservation efforts, it is clear that we must make use of
limited resources in the most effective way possible if we want to
preserve biodiversity as it currently exists on Earth. Developing
corridors between existing habitat patches has seen extensive use as a
strategy in conservation (@beier\_conceptualizing\_2012). Within this
framework, the question still remains---what corridors are the most
effective to add? The resources available to conservation are often
tightly limited and as a result being able to make the most effective
decisions to preserve biodiversity given the potential resources
available is important. We propose that simulation models, like we
describe here, can be effective in assessing functional landscape
connectivity, and planning future landscape development such that humans
can live side-by-side with the rest of Earth's inhabitants.

\pagebreak

### 2020-08-27 writing/manuscript.tex
starting revisions

\documentclass[]{article}
\usepackage{caption}
\usepackage{tocloft}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript

\usepackage{longtable}
\usepackage{graphicx}
\usepackage{tikz}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

\usepackage[left=1.5in,right=1.5in]{geometry}
\usepackage{float}
\floatplacement{figure}{h}

\usepackage{sectsty}
\usepackage[normalem]{ulem}
\sectionfont{\rmfamily\bfseries\large}
\subsectionfont{\rmfamily\bfseries\centering\upshape\normalsize}
\subsubsectionfont{\rmfamily\itshape\centering\normalsize}

\fontsize 12
\usepackage{hyperref}
\usepackage{lineno}
\linenumbers
\usepackage[font={footnotesize,sf}]{caption}
\usepackage{siunitx}

\usepackage{setspace}
\linespread{1.5}

\raggedbottom

\usepackage{titling}
\pretitle{\begin{flushleft}\Large\bfseries}
\posttitle{\end{flushleft}}
\preauthor{\begin{flushleft}\large}
\postauthor{\end{flushleft}}
\predate{\begin{flushleft}\large}
\postdate{\end{flushleft}}

\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{  \nouppercase  \leftmark}
\rhead{\thepage}
\cfoot{}

\hypersetup{breaklinks=true,
            bookmarks=true,
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\usepackage{authblk}

\title{Phase Transitions in Landscape Connectivity}
\author[1]{M.D. Catchen}
\author[2]{S.M. Flaxman}
\affil[1]{\small{Department of Ecology and Evolutionary Biology, University of Colorado at Boulder}}
\affil[2]{\small{Department of Biology, McGill University}}
\date{\today}

\usepackage[square, numbers]{natbib}
\bibliographystyle{unsrtnat}
\begin{document}
\maketitle
\begin{abstract}

dang ol absrtact here

\end{abstract}
\clearpage

\section{Introduction}

Life on Earth takes on an endless diversity of forms.
Yet, the characteristics that define ‘life’ impose parameters on the properties that biological entities must exhibit, and constraints on how these entities can change over time.
These are the fundamental principles of evolution and ecology—entities that reproduce themselves more than average become more frequent, there is a limit on resources, and so on.
Over the billions of years life has been on this planet, these forces have produced an astonishing diversity of forms and functions.
The forces that drive evolutionary and ecological processes do not occur on a single scale---they emerge out of the interactions that occur across all levels of spatial and temporal biological organization \cite{levin_problem_1992}.

Ecosystems are quintessential complex systems.
Ecological processes are inherently the product of interactions across all scales of biological organization, from the interactions between electrons that drive biochemical processes, to interactions between individual cells that constitute multicellular life, to the interactions between separate organisms in population ecology, to interactions between species in community ecology, to interactions between biogeographical patterns and biosphere level forces like climate \cite{levin_problem_1992}.
This process of \textit{emergence}, by which parts come together to form a whole with properties that don't exist among the individual parts, has been studied across a wide variety of disciplines \cite{manrubia_emergence_2004)} and is a ubiquitous phenomenon in complex systems.

One potential cause of emergent behavior in complex systems is \textit{synchrony} among individual parts. When many independent parts come together to act as a whole, their dynamics become synchronized.
This behavior is ubiquitous in biological systems across all scale of organization.
From collections of cells acting together: the heart beating in rhythm \cite{womelsdorf_modulation_2007}, neurons firing in unison \cite{strogatz_sync_2003}---to behavior among organisms: the flash of fireflies \cite{otte_theories_1980} or the migration of birds \cite{spottiswoode_extrapair_2004}---to interactions between organisms: synchrony between abundances of predators and prey, and of phenology (\cite{van_asch_phenology_2007}, \cite{burkle_future_2011}).
Synchrony, by definition, involves different entities changing over time in the same way. Within ecology, there has long been a focus on spatial synchrony, that is—how does spatial distribution of ecological entities affect whether they change together or separately? \cite{jarillo_spatial_2020, kendall_dispersal_2000, hanski_spatial_1993}.
This is, in large part, due to the applied importance of understanding the effect of habitat loss on natural populations. Many theoretical studies have shown the two primary factors that develop spatial synchrony across space are dispersal and environmental covariance \cite{ripa_analysing_2000, abbott_does_2007}.
Within this theory, one maxim that has developed is \textit{Moran's rule}, which states that spatial synchrony is proportional to the covariance in the environmental conditions across space \cite{ranta_synchrony_1995, bjornstad_spatial_1999}.

\begin{figure}
		%Do not try to scale figure in .tex or you loose font size consistency
	    \centering
		%The code to input the plot is extremely simple
		\input{plot_test}
		%Captions and Labels can be used since this is a figure environment
		\caption{Sample output from tikzDevice}
		\label{plot:test}
	\end{figure}

\clearpage
{
\footnotesize
\bibliography{references}
}

\end{document}

### 2020-08-27 output/vis.R

@@ -1,0 +1,33 @@
+library(tidyverse)
+library(ggthemr)
+library(latex2exp)
+ggthemr('fresh')
+thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
+            text=element_text(family="Iosevka Semibold",size=16), axis.title.y = element_text(angle = 0, hjust=1))
+data = read.csv('metadata.csv') %>% full_join(read.csv('treatment_set.csv'), by="treatment")
+
+data %>% 
+  group_by(treatment) %>%
+  mutate(mean_pcc = mean(summary_stat)) %>%
+  ggplot(aes(migration_rate_distribution, mean_pcc, color=factor(alpha) )) + 
+  geom_point(size=2,alpha=0.5) +
+  theme_minimal() +
+  scale_x_continuous(limits = c(0.01,1), expand = c(0, 0), breaks = c(0.01, 0.25 ,0.5, 0.75, 1.0)) +
+  labs(title='', y=TeX("$PCC$"), x=TeX("$m$"), fill=TeX("$\alpha$")) + 
+  thm + 
+  coord_cartesian(xlim = c(0.01,1.0), ylim=c(0,1)) 
+
+
+data %>% 
+  group_by(treatment) %>%
+  mutate(mean_pcc = mean(summary_stat)) %>%
+  mutate(lower_pcc = quantile(summary_stat, probs=c(0.025))) %>%
+  mutate(upper_pcc = quantile(summary_stat, probs=c(0.975))) %>%
+  ggplot(aes(migration_rate_distribution, summary_stat, group=factor(alpha))) + 
+  geom_ribbon(aes(ymin=lower_pcc, ymax=upper_pcc, fill=factor(alpha)), size=1,alpha=0.4) +
+  theme_minimal() +
+  scale_x_continuous(limits = c(0.01,1), expand = c(0, 0), breaks = c(0.01, 0.25 ,0.5, 0.75, 1.0)) +
+  labs(title='', y=TeX("$PCC$"), x=TeX("$m$"), fill=TeX("$\alpha$")) + 
+  thm + 
+  coord_cartesian(xlim = c(0.01,1.0)) 
+

### 2020-08-27 src/DispersalKernel.jl

@@ -1,15 +1,0 @@
-abstract type DispersalKernel end
-
-struct ExpKernel <: DispersalKernel
-    ExpKernel(alpha::Number, distance::Number) = exp(-1*alpha*distance)
-end
-
-struct GaussKernel <: DispersalKernel
-    GaussKernel(alpha::Number, distnace::Number) = exp( (-1* distance^2 * alpha^2) )
-end
-
-struct UniformKernel <: DispersalKernel
-    UniformKernel(::Any) = 1.0
-end
-
-

### 2020-08-27 src/DispersalPotential.jl

@@ -1,28 +1,22 @@
 
-struct DispersalPotential
-    DispersalPotential(mp::Metapopulation) = (get_dispersal_potential(mp))
-end
-
-function DispersalPotential(mp::Metapopulation, kernel::T, alpha::Number) where {T <: DispersalKernel}
-    n_pops::Int64 = mp.num_populations
-    dispersal_potential::Array{Float64} = zeros(n_pops, n_pops)
-
+function get_dispersal_potential(coordinates, kernel, alpha::Number)
+    n_pops::Int64 = length(coordinates[:,1])
+    matrix::Array{Float64,2} = zeros(n_pops, n_pops)
     for p1 = 1:n_pops
         row_sum::Float64 = 0.0
         for p2 = 1:n_pops
             if (p1 != p2)
-                d_ij::Float64 = get_distance_between_pops(mp.populations[p1], mp.populations[p2])
-                dispersal_distribution[p1,p2] = kernel(alpha, d_ij)
-                row_sum += dispersal_distribution[p1,p2]
+                d_ij::Float64 = get_distance_between_pops(coordinates[p1], coordinates[p2])
+                matrix[p1,p2] = kernel(alpha, d_ij)
+                row_sum += matrix[p1,p2]
             end
         end
 
         for p2 = 1:n_pops
             if (p1 != p2)
-                dispersal_distribution[p1,p2] /= row_sum
+                matrix[p1,p2] /= row_sum
             end
         end
     end
-    return dispersal_distribution
+    return DispersalPotential(matrix, kernel, alpha)
 end
-

### 2020-08-27 src/Dynamics.jl

@@ -1,108 +1,25 @@
-abstract type Dynamics  end
+function run_dynamics(treatment_instance::TreatmentInstance)
+    num_timesteps::Int64 = treatment_instance.simulation_parameters.number_of_timesteps
+    num_populations::Int64 = treatment_instance.metapopulation.num_populations
+    dx_dt = treatment_instance.dx_dt
+    dt::Float64 = treatment_instance.simulation_parameters.timestep_width
+    logging_frequency = treatment_instance.simulation_parameters.log_frequency
 
-struct StochasticLogisticWDiffusion <: Dynamics
-    function StochasticLogisticWDiffusion(model::DynamicsModel)
-        current_state::Array{Float64} = dynamics_model.state
-        num_populations = dynamics_model.metapopulation.num_populations
+    abundance_matrix_time_index = 1
 
-        sigma::Array{Float64} = dynamics_model.parameters.sigma_vector
-        lambda::Array{Float64} = dynamics_model.parameters.lambda_vector
-        carrying_capacity::Array{Float64} = [dynamics_model.metapopulation.populations[p].carrying_capacity for p in 1:num_populations]
-
-
-        dt::Float64 = dynamics_model.parameters.timestep_width_deterministic
-        dW::Float64 = dynamics_model.parameters.timestep_width_stochastic
-
-        new_state::Array{Float64} = zeros(num_populations)
-
-        # stochastic logistic model locally
-        for p = 1:num_populations
-            new_state[p] = stochastic_logistic_single_population(current_state[p], sigma[p], lambda[p], dt, dW, carrying_capacity[p])
-        end
+    abundance_matrix = zeros(Int64(ceil(treatment_instance.simulation_parameters.number_of_timesteps/treatment_instance.simulation_parameters.log_frequency)), treatment_instance.metapopulation.num_populations)
 
 
-        # diffusion between populations
-        dispersal_potential = dynamics_model.dispersal_potential
-        diffusion_matrix = get_dispersal_matrix(dispersal_potential, dynamics_model.parameters.migration_rate)
-        new_state = diffusion_matrix * new_state
 
-        dynamics_model.state = new_state
-
-    end
-end
-
-
-struct DynamicsModelParameters
-    number_of_timesteps::Int64
-    timestep_width_deterministic::Float64
-    timestep_width_stochastic::Float64
-    dispersal_decay_strength::Float64
-    migration_rate::Float64
-    lambda_vector::Array{Float64}
[…]

### 2020-08-27 src/IBM/IBM.jl

@@ -1,0 +1,45 @@
+# -------------------------------------------------------------
+#   IBM()
+# -------------------------------------------------------------
+#
+#
+#
+# -------------------------------------------------------------
+
+IBM(; parameters=IBMParameters(), derivative=PoissonWRandomDispersal)::IBM = IBM(parameters, derivative, zeros(parameters.max_number_individuals), zeros(parameters.max_number_individuals), zeros(parameters.metapopulation.num_populations),zeros(parameters.metapopulation.num_populations))
+
+# -------------------------------------------------------------
+#   set_initial_conditions((::IBM))
+# -------------------------------------------------------------
+#
+#
+#
+# -------------------------------------------------------------
+function set_initial_conditions(ibm::IBM)
+    num_populations = parameters.mp.num_populations
+    num_indivs_initial::Int64 = ibm.parameters.max_number_individuals / 2
+
+    trait_labels = rand(Uniform, num_indivs_initial)
+    population_labels = rand(DiscreteUniform(1, num_populations))
+
+    environmental_vector::Vector{Float64} = rand(Uniform(), num_populations)
+    abundance_vector::Vector{Int64} = get_abundance_from_population_labels(population_labels)
+end
+
+# -------------------------------------------------------------
+#   run_IBM(::IBM)
+# -------------------------------------------------------------
+#
+#
+#
+# -------------------------------------------------------------
+
+function run_IBM(ibm::IBM)
+    num_timesteps::Int64 = ibm.parameters.number_of_timesteps
+
+    for t = 1:num_timesteps
+        dxdt::Function = ibm.derivative
+        dxdt(ibm)
+    end
+
+end

### 2020-08-27 src/IBM/IBMParameters.jl

@@ -1,0 +1,28 @@
+# -------------------------------------------------------------
+#   IBMParameters()
+# -------------------------------------------------------------
+#
+#
+#
+# -------------------------------------------------------------
+
+IBMParameters(; metapopulation::Metapopulation = get_random_metapopulation(),
+                heritable_trait_variability::Number = 0.1,
+                environmental_variability::Number = 0.1,
+                selection_strength::Number = 1.0,
+                base_lambda::Number  = 1.2,
+                initial_trait_variation::Number = 0.3,
+                max_number_individuals::Number = 1000,
+                number_of_timesteps::Number = 200,
+                log_frequency::Number = 10
+             ) = IBMParameters(metapopulation,
+                               heritable_trait_variability,
+                               environmental_variability,
+                               selection_strength,
+                               base_lambda,
+                               initial_trait_variation,
+                               max_number_individuals,
+                               number_of_timesteps,
+                               log_frequency
+                              )
+

### 2020-08-27 src/IBM/PoissonWRandomDispersal.jl

@@ -1,0 +1,52 @@
+
+
+# -------------------------------------------------------------
+#   RandomDispersal()
+# -------------------------------------------------------------
+#
+#
+#
+# -------------------------------------------------------------
+function RandomDispersal(ibm::IBM)
+    (population_labels, migration_rates, dispersal_potential) = ibm.population_labels, [ibm.metapopulation], ibm.metapopulation.dispersal_potential
+
+    pop_i::Int64 = 0
+    for i = 1:max_n_indivs
+        pop_i = population_labels[i]
+        if pop_i > 0
+
+        end
+    end
+
+
+end
+
+
+# -------------------------------------------------------------
+#   PoissonWSelection()
+# -------------------------------------------------------------
+#
+#
+#
+# -------------------------------------------------------------
+function PoissonWSelection(ibm::IBM)
+
+end
+
+
+
+# -------------------------------------------------------------
+#   PoissonWRandomDispersal()
+# -------------------------------------------------------------
+#
+#
+#
+# -------------------------------------------------------------
+function PoissonWRandomDispersal(ibm::IBM)
+    num_populations::Int64 = ibm.parameters.num_populations
+    max_n_indivs::Int64 = ibm.parameters.max_number_individuals
+
+    RandomDispersal(ibm)
+    PoissonWSelection(ibm)
+
+  end

### 2020-08-27 src/IBM/derivatives.jl

@@ -1,0 +1,3 @@
+function PoissonWRandomDispersal()
+
+end

### 2020-08-27 src/Metapopulation.jl

@@ -1,30 +1,23 @@
-
-struct Population
-    carrying_capacity::Number
-    coordinate::Vector{Float64}
+function get_distance_between_pops(coord1::Array{Float64,1}, coord2::Array{Float64,1}; norm=Euclidean())
+    distance::Float64 = evaluate(norm, coord1, coord2)
+    return distance
 end
-
-struct Metapopulation
-    num_populations::Int64
-    populations::Vector{Population}
-end
-
 function get_distance_between_pops(pop1::Population, pop2::Population; norm=Euclidean())
     distance::Float64 = evaluate(norm, pop1.coordinate, pop2.coordinate)
     return(distance)
 end
 
-function get_random_metapopulation(num_populations::Int64, total_carrying_capacity::Float64; dimensions::Int64 = 2, alpha::Float64 = 3.0, carrying_capacity_distribution=nothing)
+function get_random_metapopulation(;num_populations::Int64=10, dimensions::Int64 = 2, alpha::Number = 3.0, kernel=ExpKernel)
 
-    k_subdivided::Float64 = total_carrying_capacity / num_populations
-    k::Array{Float64} = [k_subdivided for i = 1:num_populations]
     populations::Vector{Population} = []
-    #
+    coordinates::Vector{Array{Float64}} = []
+
     # random locations in the unit square
     for p = 1:num_populations
         coordinate = rand(Uniform(), dimensions)
-        push!(populations, Population(k_subdivided, coordinate))
+        push!(populations, Population(coordinate))
+        push!(coordinates, coordinate)
     end
 
-    return(Metapopulation(num_populations, populations))
+    return(Metapopulation(num_populations, populations, get_dispersal_potential(coordinates, kernel, alpha)))
 end

### 2020-08-27 src/Parameters.jl

@@ -1,0 +1,21 @@
+
+function get_stochastic_logistic_parameter_bundle(dimensionality::Int64 ;alpha_distr=Exponential(3.0), lambda_distr = Uniform(0.5,1.5), mig_distr=Uniform(), sigma_distr=(Exponential(1.0)), carrying_distr=Uniform(0.5,1.5))
+    alpha = Parameter(alpha_distr, 1) # always has dim = 1
+    m = Parameter(mig_distr, dimensionality)
+    lambda = Parameter(lambda_distr, dimensionality)
+    sigma = Parameter(sigma_distr, dimensionality)
+    k = Parameter(carrying_distr, dimensionality)
+
+    return StochasticLogisticParameterBundle(alpha, m, lambda, sigma, k)
+end
+
+function draw_from_parameter_bundle(param_bundle::StochasticLogisticParameterBundle)
+    param_values::StochasticLogisticParameterValues = StochasticLogisticParameterValues(
+        param_bundle.num_populations, 
+        param_bundle.alpha,
+        rand(param_bundle.lambda.distribution, param_bundle.lambda.dimensionality),
+        rand(param_bundle.migration_rate.distribution, param_bundle.migration_rate.dimensionality),
+        rand(param_bundle.carrying_capacity.distribution, param_bundle.carrying_capacity.dimensionality),
+        rand(param_bundle.sigma.distribution, param_bundle.sigma.dimensionality))
+    return param_values
+end

### 2020-08-27 src/StochasticLogisticWDiffusion.jl

@@ -1,53 +1,58 @@
-
 function stochastic_logistic_single_population(abundance::Float64, sigma::Float64, lambda::Float64, dt::Float64, dW::Float64, carrying_capacity::Float64)::Float64
-    delta_abundance_deterministic::Float64 =  abundance*lambda*(1.0-(abundance/carrying_capacity))
-    delta_abundance_stochastic::Float64 = rand(Normal(0, sigma*abundance))
+    if (abundance > 0)
+        delta_abundance_deterministic::Float64 =  abundance*lambda*(1.0-(abundance/carrying_capacity))
+        delta_abundance_stochastic::Float64 = rand(Normal(0, abs(sigma*abundance)))
 
-    new_abundance::Float64 = abundance + delta_abundance_deterministic*dt + delta_abundance_stochastic*dW
-    return new_abundance
-end
-
-
-function stochastic_logistic_w_diffusion(dynamics_model::DynamicsModel)::Array{Float64}
-    current_state::Array{Float64} = dynamics_model.state
-    num_populations = dynamics_model.metapopulation.num_populations
-
-    sigma::Array{Float64} = dynamics_model.parameters.sigma_vector
-    lambda::Array{Float64} = dynamics_model.parameters.lambda_vector
-    carrying_capacity::Array{Float64} = [dynamics_model.metapopulation.populations[p].carrying_capacity for p in 1:num_populations]
-
-
-    dt::Float64 = dynamics_model.parameters.timestep_width_deterministic
-    dW::Float64 = dynamics_model.parameters.timestep_width_stochastic
-
-    new_state::Array{Float64} = zeros(num_populations)
-
-    # stochastic logistic model locally
-    for p = 1:num_populations
-        new_state[p] = stochastic_logistic_single_population(current_state[p], sigma[p], lambda[p], dt, dW, carrying_capacity[p])
+        new_abundance::Float64 = abundance + delta_abundance_deterministic*dt + delta_abundance_stochastic*dW
+        return new_abundance
+    else
+        return 0.0
     end
+end
 
+struct StochasticLogisticWDiffusion <: Dynamics
+    StochasticLogisticWDiffusion() =
+    function(treatment_instance::TreatmentInstance)
+        simulation_parameters = treatment_instance.simulation_parameters
+        parameter_values::StochasticLogisticParameterValues = treatment_instance.parameter_values
+        num_populations = treatment_instance.metapopulation.num_populations
 
-    # diffusion between populations
[…]

### 2020-08-27 src/SummaryStats.jl

@@ -1,20 +1,20 @@
-abstract type SummaryStat <: Function end
 
+function PCC(abundance_matrix::Array{Float64, 2})
+        n_pops = length(abundance_matrix[1,:])
+        mean_cc::Float64 = 0.0
+        s::Float64 =0.0
+        ct::Int64 = 0
+        for p1 = 1:n_pops
+            for p2 = (p1+1):n_pops
+                v1 = abundance_matrix[:,p1]
+                v2 = abundance_matrix[:,p2]
+                cc = crosscor(v1,v2,[0])
+                s += cc[1]
+                ct += 1
+            end
+        end
 
-
-# Pairwise Crosscorrelation
-struct PCC <: SummaryStat
-
-    function PCC(::DynamicsModel, ::Population, ::Population)
-        return 1
-    end
-
-    function PCC(::DynamicsModel)
-        return 0
-    end
+        mean_cc = s/ct
+        return mean_cc
 end
 
-# Mean Pairwise Crosscorrelation, where all
-struct MeanPCC <: SummaryStat
-
-end

### 2020-08-27 src/TreatmentFactory.jl

@@ -1,31 +1,104 @@
 
+function check_for_nondefaults(param_dict::Dict)
+
+    # Flag to not generate a random metapopulation every replicate
+        # use-cases:
+        #   - user defined metapopulation provided as CSV
+        #   - only generate a new metapopulation at the Treatment or higher level
+
+    # Flag to use a defined random seed
+    #
+    #
+
+    # Flag to use stochastic dispersal, where dispersal events
+    # occur according random draws from a Binomial(abundance, migration_probability).
+    #
+
+    # Flag to use
+
+end
+
+function create_metadata_df(param_dict::Dict)
+	names::Array{String} = []
+	vals = []
+	metadata = DataFrame(names...)
+	metadata.treatment = []
+
+	for (param_name, param_values) in param_dict
+		push!(names, param_name)
+		push!(vals, param_values)
+		metadata[!,  Symbol(param_name)] = []
+	end
+	ct = 1
+	for p in Iterators.product(vals...)
+		for (i, val) in enumerate(p)
+			param = names[i]
+			push!(metadata[!, Symbol(param)], val)
+		end
+		push!(metadata.treatment, ct)
+		ct += 1
+	end
+	return(metadata)
+end
+
+
 # =================================================
 #    treatment maker
 #
 #
 # =================================================
-function create_treatments(param_dict::Dict)
-	names::Array{String} = []
-	vals = []
-	treatment_df = DataFrame(names...)
-	treatment_df.treatment = []
-
-	for (param_name, param_values) in param_dict
-		push!(names, param_name)
-		push!(vals, param_values)
-		treatment_df[!,  Symbol(param_name)] = []
+function get_dist(item)
+	if typeof(item) <: Distribution
+		return item
+	else
+		return Normal(item, 0.0)
 	end
+end
 
-	ct = 1
-	for p in Iterators.product(vals...)
-		for (i, val) in enumerate(p)
-			param = names[i]
-			push!(treatment_df[!, Symbol(param)], val)
+
+function create_treatments(param_dict::Dict; replicates_per_treatment::Int64 = 50)
+	metadata::DataFrame = create_metadata_df(param_dict)
+	n_treatments::Int64 = nrow(metadata)
+	
+	treatments::Vector{Treatment} = []
+
[…]

### 2020-08-27 src/Vis.jl

@@ -1,8 +1,3 @@
-function plot_timeseries_of_abundances(dynamics_model::DynamicsModel)
-    @show dynamics_model.log
-    num_timesteps = length(dynamics_model.log[:,1])
-    plot(dynamics_model.time_vector, dynamics_model.log[:,:],
-         linewidth=2,
-         ylims=(0.7,1.2),
-         size=(700, 400))
+function plot_timeseries_of_abundances(abundance_matrix)
+    plot(collect(1:length(abundance_matrix[:,1])), abundance_matrix, linewidth=2,      size=(700, 400))
 end

### 2020-08-27 src/deps.jl

@@ -3,6 +3,8 @@ using StatsBase, Distributions, DataFrames, CSV, ProgressMeter, LinearAlgebra, D
 wd = pwd()
 cd("/home/michael/phase_transitions_in_metapopulation_synchrony/src")
 
+include("./types.jl")
+include("./Parameters.jl")
 include("./Metapopulation.jl")
 include("./DispersalKernel.jl")
 include("./DispersalPotential.jl")
@@ -13,5 +15,11 @@ include("./StochasticLogisticWDiffusion.jl")
 include("./SummaryStats.jl")
 include("./Vis.jl")
 
+
+# Source IBM stuff, maybe should be packaged as a module in the future
+include("./IBM/types.jl")
+include("./IBM/IBMParameters.jl")
+include("./IBM/IBM.jl")
+include("./IBM/PoissonWRandomDispersal.jl")
 
 cd(wd)

### 2020-08-27 src/types.jl

@@ -1,0 +1,159 @@
+abstract type Dynamics end
+abstract type DispersalKernel end
+
+
+struct GaussKernel <: DispersalKernel
+    GaussKernel(alpha::Number, distance::Number) = exp(-1*alpha^2*distance^2)
+end
+
+
+struct ExpKernel <: DispersalKernel
+    ExpKernel(alpha::Number, distance::Number) = exp(-1*alpha*distance)
+end
+
+# -------------------------------------------------------------
+#   ParameterBundle()
+# -------------------------------------------------------------
+#
+#
+#
+# -------------------------------------------------------------
+
+struct Parameter
+    distribution::Distribution
+    hyperparameters::Vector{Parameter}
+    dimensionality::Int64
+    Parameter(distribution::Distribution, dimensionality::Int64; hyperparameters=[])  = new(distribution, hyperparameters, dimensionality)
+    Parameter(distribution::Distribution; dimensionality=1) = new(distribution, [], dimensionality)
+end
+
+Base.show(io::IO, p::Parameter) = print(io, "Parameter ~ ", p.distribution, "\n")
+
+
+abstract type ParameterBundle end
+abstract type ParameterValues end
+
+struct StochasticLogisticParameterBundle <: ParameterBundle
+    num_populations::Int64
+    alpha::Float64
+    migration_rate::Parameter
+    lambda::Parameter
+    sigma::Parameter
+    carrying_capacity::Parameter
+end
+
+
+
+struct StochasticLogisticParameterValues <: ParameterValues
+    num_populations::Int64
+    alpha::Float64  # dispersal kernel alpha
+    lambda::Vector{Float64} # lambda across pops
+    migration_rate::Vector{Float64} # mig across pops
+    carrying_capacity::Vector{Float64} # K across pops
+    sigma::Vector{Float64} # sigma across pops
+end
+
+Base.show(io::IO, params::StochasticLogisticParameterValues) = print(io, 
+                                                                    "Stochastic Logistic Model with Parameters: \n",
+                                                                    "\t\t\talpha:", params.alpha, "\n",
+                                                                    "\t\t\tlambda:", params.lambda, "\n",
+                                                                    "\t\t\tmigration:", params.migration_rate, "\n",
+                                                                    "\t\t\tcarrying capacity:", params.carrying_capacity, "\n",
+                                                                    "\t\t\tsigma:", params.sigma, "\n"
+
+                                                                )
+
+
[…]

### 2020-08-27 test.jl

@@ -1,45 +1,50 @@
 include("./src/deps.jl")
 
 
-mp = get_random_metapopulation(10, 10.)
 
-model = DynamicsModel(mp, StochasticLogisticWDiffusion, parameters=DynamicsModelParameters(mp, migration_rate = 0.9))
+#model = get_treatment()
+#@show model
+#run_dynamics(model)
+#plot_timeseries_of_abundances(model)
 
-@show model
 
-run_dynamics(model)
-plt = plot_timeseries_of_abundances(model)
 treatment_dictionary =
 
                     Dict(
                         # =======================================
                         # dynamics model parameters
                         # =======================================
-                        "migration_rate" => vcat(collect(0.01:0.01:1.)),
-                        "base_lambda" => [1.0],
-                        "base_sigma" => [0.4],
+                        "migration_rate_distribution" => [i for i = 0.01:0.01:1.0],
+                        "lambda_distribution" => [Uniform(0.5,1.5)],
+                        "sigma_distribution" => [1.0],
+                        "carrying_capacity_distribution" => [Uniform(0.5,1.5)],
+
+                        "parameters_fixed_across_populations" => [false],
 
                         # ========================================
                         # parameters for metapopulation generation
                         # ========================================
-                        "metapopulation_generator" => [get_random_metapopulation]
-                        "fixed_metapopulation" => [false],
+                        "metapopulation_generator" => [get_random_metapopulation],
                         "num_populations" => [20],
-                        "base_carrying_capacity" => [20.],
+                        "alpha" => [0, 10, 20],
+                        "fixed_metapopulation_per_treatment" => [false],
 
                         # ===============================================================
                         # parameters that should not change unless you have a good reason
                         # ===============================================================
-                        "number_of_timesteps" => [1000],
-                        "variable_sigma_across_populations" => [false],
-                        "variable_lambda_across_populations" => [false],
+                        "number_of_timesteps" => [300]
                     );
 
 
-create_treatments(treatment_dictionary)
+
+treatment_set = create_treatments(treatment_dictionary, replicates_per_treatment=1)
+df = run_treatments(treatment_set)
+
+inst = treatment_set.treatments[200].instances[10]
+inst
+plot(collect(1:length(inst.abundance_matrix[:,1])), inst.abundance_matrix)
 
 
-
-
-
-plt
+cd("/home/michael/phase_transitions_in_metapopulation_synchrony/")
+CSV.write("output/metadata.csv",treatment_set.metadata)
+CSV.write("output/treatment_set.csv", df)

### 2020-08-28 writing/manuscript.tex
updates

@@
 \sectionfont{\rmfamily\bfseries\large}
-\subsectionfont{\rmfamily\bfseries\centering\upshape\normalsize}
-\subsubsectionfont{\rmfamily\itshape\centering\normalsize}
+\subsectionfont{\rmfamily\centering\upshape\normalsize}
+\subsubsectionfont{\centering\normalsize}
 
@@
 \title{Phase Transitions in Landscape Connectivity}
-\author[1]{M.D. Catchen}
-\author[2]{S.M. Flaxman}
+\author[1,2]{M.D. Catchen}
+\author[1]{S.M. Flaxman}
 \affil[1]{\small{Department of Ecology and Evolutionary Biology, University of Colorado at Boulder}}
@@
 
-Life on Earth takes on an endless diversity of forms.
-Yet, the characteristics that define ‘life’ impose parameters on the properties that biological entities must exhibit, and constraints on how these entities can change over time.
-These are the fundamental principles of evolution and ecology—entities that reproduce themselves more than average become more frequent, there is a limit on resources, and so on.
-Over the billions of years life has been on this planet, these forces have produced an astonishing diversity of forms and functions.
-The forces that drive evolutionary and ecological processes do not occur on a single scale---they emerge out of the interactions that occur across all levels of spatial and temporal biological organization \cite{levin_problem_1992}.
+Landscape connectivity intro
 
-Ecosystems are quintessential complex systems.
-Ecological processes are inherently the product of interactions across all scales of biological organization, from the interactions between electrons that drive biochemical processes, to interactions between individual cells that constitute multicellular life, to the interactions between separate organisms in population ecology, to interactions between species in community ecology, to interactions between biogeographical patterns and biosphere level forces like climate \cite{levin_problem_1992}.
-This process of \textit{emergence}, by which parts come together to form a whole with properties that don't exist among the individual parts, has been studied across a wide variety of disciplines \cite{manrubia_emergence_2004)} and is a ubiquitous phenomenon in complex systems.
-
-One potential cause of emergent behavior in complex systems is \textit{synchrony} among individual parts.
+Ecological processes are inherently the product of interactions across all scales of biological organization \cite{levin_problem_1992}.
+This process of \textit{emergence}, by which parts come together to form a whole with properties that don't exist among the individual parts, has been studied across a wide variety of disciplines \cite{manrubia_emergence_2004} and is a ubiquitous phenomenon in complex systems.
+One potential cause of emergent behavior is \textit{synchrony} among the individual parts of a complex system.
 When many independent parts come together to act as a whole, their dynamics become synchronized.
 This behavior is ubiquitous in biological systems across all scale of organization.
-From collections of cells acting together: the heart beating in rhythm \cite{womelsdorf_modulation_2007}, neurons firing in unison \cite{strogatz_sync_2003}---to behavior among organisms: the flash of fireflies \cite{otte_theories_1980} or the migration of birds \cite{spottiswoode_extrapair_2004}---to interactions between organisms: synchrony between abundances of predators and prey, and of phenology (\cite{van_asch_phenology_2007}, \cite{burkle_future_2011}).
+From collections of cells acting together: the heart beating in rhythm \cite{womelsdorf_modulation_2007}, neurons firing in unison \cite{strogatz_sync_2003}---to behavior among organisms: the flash of fireflies \cite{otte_theories_1980} or the migration of birds \cite{spottiswoode_extrapair_2004}---to interactions between organisms: synchrony between abundances of predators and prey \cite{lv}, and of phenology \cite{van_asch_phenology_2007,burkle_future_2011}.
 Synchrony, by definition, involves different entities changing over time in the same way.
@@
 
+A major goal of conservation is developing corridors to conserve
+landscape connectivity. Most measures of landscape connectivity used in the literature represent \emph{structural} connectivity, meaning quantifying the
+structure of the landscape \cite{}.
+Structural connectivity stands in constract with \emph{functional} connectivity, which measures the connectivity of a given process \cite{kool_population_2013, calabrese_comparison-shoppers_2004}.
+Here,
+in order to better understand functional connectivity, we use a
+simulation model to measure how synchrony across space changes as a
+function of landscape structure. We do this by developing a model of
+metapopulation dynamics on spatial graphs, which have long been used to
+model landscape connectivity \cite{martensen_spatio-temporal_2017,
+@albert_applying_2017, @urban_landscape_2001}, and analyze how
+synchrony changes across space using the language of critical
+transitions.
+
+Further, we show that increasing population synchrony reduces the
+variance in the generation-to-generation change in abundance, which is
+central in reducing the probability of metapopulation extinction
+(@lande\_risks\_1993, @lande\_extinction\_1998). This relationship
+suggests that promoting functional landscape connectivity can help
+mediate the probability of extinction for species facing significant
+habitat loss. We suggest using simulation models, such as those
+presented here, to aid in decision making regarding corridor placement.
+
+\section{What are phase transitions?}
+
+\hypertarget{what-are-phase-transitions}{%
+\subsection{What are phase
+transitions?}\label{what-are-phase-transitions}}
+
+When does a system change from one state to a different state? Due to
+the rapid changes induced on the planet by human activity, there has
+been recent focus on answering this question in ecology, especially the
+potential for changes in spatial structure to drive transitions between
+alternative stable states. Much of this theory has been aimed at the
+practical problem of being able to predicting the onset of transitions
+from time-series data \cite{scheffer_anticipating_2012, scheffer_early-warning_2009}. The bulk of the quantitative theory
+used to understand transitions between states is derived from
+statistical mechanics, where it was originally used to study
+phase-transitions in matter. In order to study phase transitions between
+regimes, we must first be clear on what these regimes are. As this
+theory was originally used to describe physical states of matter, the
+original regimes were solid, liquid, gas. However, as our understanding
+of condensed-matter has changed, so has the demarcation of what
+constitutes different states. In reality, the way in which particles
+come together to constitute matter is far more variable than these three
+categories. In such cases the `state' of a collection of particles
+cannot be represented by a single categorical label, and so the theory
+of phase transitions as been adapted to model \emph{continuous} phase
+transitions, where there is no clear demarcation point between different
+states \cite{sethna_statistical_2006}. This is useful for us in ecology,
+where the line between alternative ecosystem `states' is even fuzzier.
+
+We can formalize our understanding of phase transitions using the
+language of statistical mechanics. We call the \emph{order parameter}
+some measure of the system's state in space and time. The \emph{control
+parameter}, then, is what causes the change in order parameter. When
+dealing with dynamics that are inherently stochastic, one tool often
+used in statistical mechanics is correlation functions which measure how
+well the order parameter is correlated in both space and time at a
+particular value of the control parameter \cite{sethna_statistical_2006}.
+For example, if we consider the population of a species inhabiting a
+landscape, where along the gradient of landscape connectivity, our
+control parameter, does that system go from consisting of one large,
+single, population, to many small, independent populations? We measure
+this qualitative shift from one system to many using \emph{synchrony},
+the correlation in the dynamics of abundance across space.
+
+\pagebreak
+
+\hypertarget{the-model}{%
+\section{The Model}\label{the-model}}
+
+Here we present a spatial graph model of landscape connectivity based on
+metapopulation theory \cite{hanski_practical_1994, grilli_metapopulation_2015} . We model connectivity as a function of
+a few empirically estimable parameters, and then describe a stochastic
+model of metapopulation dynamics on these spatial graphs. We then
+simulate dynamics across a gradient of landscape
+connectivity parameters to measure transitions in the synchrony of
+population dynamics across space occur.
+
+\hypertarget{modeling-landscape-connectivity-with-a-spatial-graph}{%
+\subsection{Modeling Landscape Connectivity with a Spatial
+Graph}\label{modeling-landscape-connectivity-with-a-spatial-graph}}
+
+Spatial graphs have long been used to model a system of habitat patches
+\cite{dale_graphs_2010, minor_graph-theory_2008,
+urban_landscape_2001}. Here we model a system of populations, represented as a vector of
+vertices, $\vec{L}$ in a spatial graph \(G=(L,E)\).
+Here the edges \(E\)  represent dispersal between populations.
+To define $E$,  we choose to model landscape connectivity with respect to the process of metapopulation dynamics as a combination of two different factors: the probability than any individual migrates during its lifetime, $m$, and the conditional distribution over spatial nodes of where an individual goes, given that
+it migrates, \(P(L_j|L_i)\), which we call the dispersal potential, $\Phi_{ij}$. We can model the dispersal potential using a few methods. In empirical
+systems, this can be estimated with resistance surfaces, which provide
+relative weights of the difficulty of migration between points on a
+raster of land-cover type \cite{spear_use_2010}. Theoretically, we model
+the dispersal potential using isolation-by-distance (IBD). The relative
+probability of dispersal between \(L_i\) to \(L_j\) is inversely
+proportional to the distance between them, \(d_{ij}\), and the strength
+of this isolation-by-distance relationship, \(\alpha\). The
+functional form of this relationship, \(f(d_{ij}, \alpha)\), has long been called the dispersal kernel \cite{hanski_practical_1994}.
+Here we consider two different types of dispersal kernels: the
+exponential, \(f(d_{ij}, \alpha)=e^{-\alpha d_{ij}}\), and
+Gaussian, \(f(d_{ij}, \alpha)=e^{-\alpha^2 d_{ij}^2}\), which have both been considered as dispersal kernels in both theoretical and empirical work \cite{hanski_practical_1994,grilli_metapopulation_2015}.
+
+To construct a dispersal potential $\Phi_{ij}$ with a kernel $f(d_{ij}, \alpha)$,  we normalize:
+
+\[P(V_j|V_i)=\frac{f(d_{ij}, \alpha)}{\sum_k f(d_{ik},\alpha)}\]
+
+Note that if \(\alpha=0\), then the value of both exponential and
+Gaussian kernels is the same for all pairs of populations, and therefore
+the dispersal potential is a uniform distribution. In Figure \ref{fig:mp},
+we can see spatial graphs plotted representing the same set of
+populations across differing values.
+
 \begin{figure}
@@
 		%The code to input the plot is extremely simple
-		\input{plot_test}
+		\includegraphics[width=15cm]{./figs/figure1}
 		%Captions and Labels can be used since this is a figure environment
-		\caption{Sample output from tikzDevice}
-		\label{plot:test}
+		\caption{Sample output from tikzDevice}	\label{fig:mp}
+\end{figure}
+
+Now that we have a model of landscape connectivity, we turn to defining a model of metapopulation dynamics on these spatial graphs.
+
+\hypertarget{local-dynamics}{%
+\subsection{Local Dynamics}\label{local-dynamics}}
+
+We model population dynamics within each local population $L_i$ using
+the stochastic logistic model. The dynamics of the number of individuals $N_i$
+in population $L_i$  are described by the stochastic differential
+equation (SDE)
+
+\[dN_i=\lambda_i N_i \big(1 - \frac{N_i}{K_i}\big)dt+\sigma_pN_idW\]
+
+Here, \(N_i\) is the abundance of population \(L_i\), \(\lambda_i\) is the
+strength of density dependence in that population, and \(K_i\) is the
+carrying capacity of \(L_i\). \(\sigma_p\) represents that standard
+deviation in abundance due to local stochasticity as a proportion of
+\(N_i\). Here \(\sigma_p\) represents an amalgamation of all factors
+contributing to local stochasticity in population dynamics, although it
+should be noted that the relative contribution of different factors to
+local stochasticity can drive significant variation in dynamics
+\cite{melbourne_extinction_2008}. For the sake of reducing the size of parameter
+space, here we consider all populations as having the same \(\lambda_i\)
+and \(K_i\), however, future work could include exploring the
+source-sink dynamics in this system by varying intrinsic growth rates
+and carrying capacities across populations.
+
+We use an SDE representation because they have been used to study phase
+transitions in stochastic systems before \cite{brock_early_2012, kuehn_mathematical_2011}, and they have many nice properties for inference. SDEs
+have been used to study extinction dynamics before, as it is relatively
+straightforward to compute the mean time until extinction (MTE) using
+the Kolmogorov Backward Equation (@lande\_stochastic\_2003).
+
+\hypertarget{diffusion-on-spatial-graphs}{%
+\subsubsection{Diffusion on Spatial
+Graphs}\label{diffusion-on-spatial-graphs}}
+
+When we model a landscape with a spatial graph, we have to decide how
+different nodes affect one another. The processes that connect
+landscapes are inherently stochastic. The probability that an individual
+migrates within its lifetime, \(p_m\), and where it goes, \(P(V_j|V_i)\)
+, are both stochastic processes. Under some conditions, we can
+effectively model stochastic dynamics across space using diffusion
+models. In essence, a diffusion model assumes that at each timestep the
+system will change according to the expected value of stochastic
+dispersal. Diffusion models have seen widespread use in ecology and
+other fields (@ovaskainen\_empirical\_2008,
+@holmes\_partial-differential\_1994, @okubo\_diffusion\_2011).
+
+Given our dispersal potential \(P(V_j|V_i)\), we can define the
+diffusion matrix \(\Phi\) as
+
+\[\Phi_{ij} = \begin{cases} 1 - p_m \quad\quad &\text{if}\ i = j \\ P(V_j|V_i)p_m & \text{if}\ i \neq j\end{cases}\]
+
+where \(\Phi_{ij}\) represents the probability that any individual born
+in \(V_i\) reproduces in \(V_j\).
+
+Now we move to considering the dynamics of the abundance of the
+population at \(V_i\), denoted \(N_i\). We can now represent the
+dynamics due to diffusion of this system as
+
+\[\dot{N_i}=(1-p_m)N_i+ \sum_j p_mP(V_j|V_i)N_j\]
+
+In matrix notation, we can represent this diffusion model as
+
+\[\frac{d\vec{N}}{dt}=\Phi^T\vec{N}\]
+
+We can then combine this with local dynamics as a reaction-diffusion
+model,
+
+\[\frac{d\vec{N}}{dt} = g(\Phi^T \vec{N})\]
+
+where \(g(x)\) is a function that represents the hypothesized mechanism
+of how the ecological measurement evolves locally.
+
+In principle, \(g(x)\) can represent any ecological process of
+interest---for example if the state space of \(x\) is allelic
+frequencies, \(g(x)\) could describe genetic drift, or if \(x\)
+represents community compositions across space, \(g(x)\) could describe
+competition between species as a function of environmental conditions,
+coevolutionary states across space, etc. Here, we consider \(g(x)\) to
+be the stochastic logistic model (see previous section). Combining this
+with the diffusion model yields the SDE
+
+\begin{equation} \label{diffusion_model}
+  d \vec{N} = {I} (\Phi^T \vec{N})(\vec{k}-\vec{N}) dt + \sigma_p \vec{k} \ d\vec{W}
+\end{equation}
+
+which will be the primary object of study in this paper.
+
+\hypertarget{measuring-synchrony}{%
+\subsubsection{Measuring Synchrony}\label{measuring-synchrony}}
+
+As described above, measures of correlation in space and time are often
+used in the study of phase transitions. When the order parameter gains
+or loses correlation in space and time is used as an indicator of when a
+system changes qualitative phases. Within the context of ecology, the
+crosscorrelation function, \(CC\), has long been used as a measure of
+synchronous dynamics (@liebhold\_spatial\_2004). Here, with a subdivided
+population, we consider the mean crosscorrelation compared across all
+populations, \({PCC}=\frac{1}{n^2}\sum_{i,j} CC(V_i,V_j)\) . As an
+example, in \ref{async_and_sync}, panel (a) show an example of low PCC,
+and panel (b) shows an example of high PCC.
+
+\begin{figure}[h]
+\caption{The abundance of 5 populations (each in a different color) across time.
+Panel (a):  These populations show low synchrony, $PCC=0.1$.
+Panel (b): These populations show high sychrony, $PCC=0.89$ }
+
+\label{async_and_sync}
+
+\end{figure}
+
+\hypertarget{simulating-a-phase-transition-across-a-migration-gradient}{%
+\subsubsection{Simulating a Phase Transition across a Migration
+Gradient}\label{simulating-a-phase-transition-across-a-migration-gradient}}
+
+Now we consider how synchrony changes as a function of \(p_m\). Each run
+consists of the following parameters,
+\(\theta = \{N_p, \lambda, \sigma_p, \vec{K},p_m, \alpha \}\). For each
+unique set of parameters, we run \(50\) replicates across all values of
+\(p_m = \{0.01,0.02,\dots,1.0 \}\). For each replicate, we independently
+draw the location of \(N_p\) populations uniformly in \([0,1]^2\). We
+draw the intial value of abundance for all populations from a uniform
+distribution, \(N_i \sim U(0, K_i)\). We then integrate equation
+\ref{diffusion_model} forward \(500\) timesteps using the
+Euler--Maruyama method with \(\Delta t=0.1\). After integrating forward,
+we compute the crosscorrelation coefficient, \(CC\) for each pair of
+populations \(i \neq j\). Then, we compute the mean pairwise
+crosscorrelation for that replicate, \(PCC\), and start the next
+replicate.
+
+\pagebreak
+
+\begin{figure}
+    \includegraphics[width=15cm]{figs/figure4.pdf}
+    \caption{}
+    \label{}
 \end{figure}

### 2020-08-28 figs/figure1_metapopulations_as_a_network/figure1.R

@@ -6,6 +6,8 @@ library(tidyverse)
 library(extrafont)
 library(latex2exp)
 library(Cairo)
+library(extrafont)
+loadfonts()
 setwd("~/phase_transitions_in_metapopulation_synchrony/figs")
 loadfonts()
 

### 2020-08-28 figs/figure3_2population_phase_transition/figure3.jl

@@ -1,0 +1,51 @@
+
+# Project Folder
+proj_folder = "/home/michael/phase_transitions_in_metapopulation_synchrony/"
+cd(proj_folder)
+
+include("./src/deps.jl")
+output_dir_path = string(proj_folder , "figs/figure3_2population_phase_transition/output/")
+
+
+
+# _______________________________________________________________________________________
+#
+#
+#   phase transition diagram for two populations
+#
+#
+#
+#
+#
+# _______________________________________________________________________________________
+
+param_dictionary = Dict(
+                        "migration_rate"    => [i for i = 0.0:0.01:1.0],
+                        "lambda"            => [0.5,1.0,2],
+                        "sigma"             => [0.5, 0.1, 1.0 ],
+                        "carrying_capacity" => [1.0],
+                        "num_populations" => [2],
+                        "alpha"           => [0],  # alpha is meaningless with two pops
+
+                        # _________________________________________________
+                        #
+                        # simulation parameters
+                        #
+                        # _________________________________________________
+                        "number_of_timesteps" => [300],
+                        "metapopulation_generator" => [get_random_metapopulation]
+                    )
+
+
+
+treatment_set = create_treatments(param_dictionary, replicates_per_treatment = 50)
+
+df = run_treatments(treatment_set)
+
+mkdir(output_dir_path)
+
+CSV.write(string(output_dir_path, "treatment_set.csv"), df)
+CSV.write(string(output_dir_path, "metadata.csv"), treatment_set.metadata)
+
+
+

### 2020-08-28 figs/figure3_2population_phase_transition/vis.R

@@ -1,0 +1,60 @@
+setwd("~/phase_transitions_in_metapopulation_synchrony/figs/figure3_2population_phase_transition/output")
+library(tidyverse)
+library(ggthemr)
+library(latex2exp)
+library(extrafont)
+loadfonts()
+ggthemr('fresh', spacing=2)
+cmuserif = p$`CMU Serif`$family
+thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
+            panel.spacing=unit(3, "lines"),
+            text=element_text(family=cmuserif,size=20), axis.title.y = element_text(angle = 0, vjust=0.5),
+            axis.title = element_text(margin = margin(t = 20, r = 20, b = 0, l = 0)),
+            axis.text = element_text(size=14),
+            panel.grid = element_line(size=0.5, linetype="dotted"),
+            # The new stuff
+            strip.text = element_text(size = 22),
+            strip.text.y = element_text(size = 22, angle = 0, vjust=0.5))
+data = read.csv('metadata.csv') %>% full_join(read.csv('treatment_set.csv'), by="treatment")
+
+
+data %>% 
+  group_by(treatment) %>%
+  filter(lambda==1.0, sigma==0.5) %>%
+  ggplot(aes(migration_rate, summary_stat, group=factor(replicate) )) + 
+  geom_line(size=1,alpha=0.3) +
+  theme_minimal() +
+  scale_x_continuous(limits = c(0.0,1), expand = c(0, 0), breaks = c(0.0, 0.25 ,0.5, 0.75, 1.0)) +
+  labs(title='', y=TeX("$PCC$"), x=TeX("$m$"), color=TeX("$\\alpha$")) + 
+  thm + 
+  facet_grid(vars(sigma), vars(lambda)) +
+  coord_cartesian(xlim = c(0.0,1.0), ylim=c(0,1)) 
+
+
+convert_to_tex_label = function(string) {
+  return(TeX(string))
+}
+
+
+
+data %>% 
+  group_by(treatment) %>%
+  mutate(mean_pcc = mean(summary_stat)) %>%
+  mutate(lowest_pcc = quantile(summary_stat, probs=c(0.025))) %>%
+  mutate(lower_pcc = quantile(summary_stat, probs=c(0.25))) %>%
[…]

### 2020-08-28 figs/figure4_20population_phase_transition_varying_alpha/figure4.jl

@@ -1,0 +1,47 @@
+
+# Project Folder
+proj_folder = "/home/michael/phase_transitions_in_metapopulation_synchrony/"
+cd(proj_folder)
+
+include(string(proj_folder, "src/deps.jl"))
+
+output_dir_path = string(proj_folder , "figs/figure4_20population_phase_transition_varying_alpha/output/")
+mkdir(output_dir_path)
+# _______________________________________________________________________________________
+#
+#
+#   phase transition diagram for 20 populations, varying alpha
+#
+#
+#
+#
+#
+# _______________________________________________________________________________________
+
+param_dictionary = Dict(
+                        "migration_rate"    => [i for i = 0.0:0.01:1.0],
+                        "lambda"            => [1.0],
+                        "sigma"             => [0.5 ],
+                        "carrying_capacity" => [1.0],
+                        "num_populations" => [20],
+                        "alpha"           => [0, 5, 10, 15],
+                        # _________________________________________________
+                        #
+                        # simulation parameters
+                        #
+                        # _________________________________________________
+                        "number_of_timesteps" => [300],
+                        "metapopulation_generator" => [get_random_metapopulation]
+                    )
+
+
+
+treatment_set = create_treatments(param_dictionary, replicates_per_treatment = 50)
+
+df = run_treatments(treatment_set)
+
+CSV.write(string(output_dir_path, "treatment_set.csv"), df)
+CSV.write(string(output_dir_path, "metadata.csv"), treatment_set.metadata)
+
+
+

### 2020-08-28 figs/figure4_20population_phase_transition_varying_alpha/figure4_vis.R

@@ -1,0 +1,63 @@
+setwd("~/phase_transitions_in_metapopulation_synchrony/figs/figure4_20population_phase_transition_varying_alpha/output")
+library(tidyverse)
+library(ggthemr)
+library(latex2exp)
+library(extrafont)
+loadfonts()
+ggthemr('fresh', spacing=3)
+cmuserif = pdfFonts()$`CMU Serif`$family
+thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
+            panel.spacing=unit(3, "lines"),
+            text=element_text(family=cmuserif,size=20), axis.title.y = element_text(angle = 0, vjust=0.5),
+            axis.title = element_text(margin = margin(t = 20, r = 20, b = 0, l = 0)),
+            axis.text = element_text(size=14),
+            panel.grid = element_line(size=1, linetype="dotted"),
+            # The new stuff
+            strip.text = element_text(size = 22),
+            strip.text.y = element_text(size = 22, angle = 0, vjust=0.5))
+
+data = read.csv('metadata.csv') %>% full_join(read.csv('treatment_set.csv'), by="treatment")
+
+
+data %>% 
+  group_by(treatment) %>%
+  filter(lambda==1.0, sigma==0.5) %>%
+  ggplot(aes(migration_rate, summary_stat, group=factor(replicate) )) + 
+  geom_line(size=1,alpha=0.3) +
+  theme_minimal() +
+  scale_x_continuous(limits = c(0.0,1), expand = c(0, 0), breaks = c(0.0, 0.25 ,0.5, 0.75, 1.0)) +
+  labs(title='', y=TeX("$PCC$"), x=TeX("$m$"), color=TeX("$\\alpha$")) + 
+  thm + 
+  facet_grid(vars(sigma), vars(alpha)) +
+  coord_cartesian(xlim = c(0.0,1.0), ylim=c(0,1)) 
+
+convert_to_tex_label = function(string) {
+  return(TeX(string))
+}
+
+plt = data %>% 
+  group_by(treatment) %>%
+  mutate(mean_pcc = mean(summary_stat)) %>%
+  mutate(lowest_pcc = quantile(summary_stat, probs=c(0.025))) %>%
+  mutate(lower_pcc = quantile(summary_stat, probs=c(0.25))) %>%
[…]

### 2020-08-28 figs/figure5_20population_phase_transition_varying_sigma_and_lambda/figure5.jl

@@ -1,0 +1,47 @@
+
+# Project Folder
+proj_folder = "/home/michael/phase_transitions_in_metapopulation_synchrony/"
+cd(proj_folder)
+
+include(string(proj_folder, "./src/deps.jl"))
+
+output_dir_path = string(proj_folder , "figs/figure5_20population_phase_transition_varying_sigma_and_lambda/output/")
+mkdir(output_dir_path)
+# _______________________________________________________________________________________
+#
+#
+#   phase transition diagram for two populations
+#
+#
+#
+#
+#
+# _______________________________________________________________________________________
+
+param_dictionary = Dict(
+                        "migration_rate"    => [i for i = 0.0:0.01:1.0],
+                        "lambda"            => [0.5,1.0,2],
+                        "sigma"             => [0.5, 0.1, 1.0 ],
+                        "carrying_capacity" => [1.0],
+                        "num_populations" => [20],
+                        "alpha"           => [0],
+                        # _________________________________________________
+                        #
+                        # simulation parameters
+                        #
+                        # _________________________________________________
+                        "number_of_timesteps" => [300],
+                        "metapopulation_generator" => [get_random_metapopulation]
+                    )
+
+
+
+treatment_set = create_treatments(param_dictionary, replicates_per_treatment = 50)
+
+df = run_treatments(treatment_set)
+
+CSV.write(string(output_dir_path, "treatment_set.csv"), df)
+CSV.write(string(output_dir_path, "metadata.csv"), treatment_set.metadata)
+
+
+

### 2020-08-28 figs/figure5_20population_phase_transition_varying_sigma_and_lambda/figure5_vis.R

@@ -1,0 +1,65 @@
+setwd("~/phase_transitions_in_metapopulation_synchrony/figs/figure5_20population_phase_transition_varying_sigma_and_lambda/output")
+library(tidyverse)
+library(ggthemr)
+library(latex2exp)
+library(extrafont)
+loadfonts()
+ggthemr('fresh', spacing=3)
+cmuserif = p$`CMU Serif`$family
+thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
+            panel.spacing=unit(3, "lines"),
+            text=element_text(family=cmuserif,size=20), axis.title.y = element_text(angle = 0, vjust=0.5),
+            axis.title = element_text(margin = margin(t = 20, r = 20, b = 0, l = 0)),
+            axis.text = element_text(size=14),
+            panel.grid = element_line(size=0.5, linetype="dotted"),
+            # The new stuff
+            strip.text = element_text(size = 22),
+            strip.text.y = element_text(size = 22, angle = 0, vjust=0.5))
+data = read.csv('metadata.csv') %>% full_join(read.csv('treatment_set.csv'), by="treatment")
+
+
+data %>% 
+  group_by(treatment) %>%
+  filter(lambda==1.0, sigma==0.5) %>%
+  ggplot(aes(migration_rate, summary_stat, group=factor(replicate) )) + 
+  geom_line(size=1,alpha=0.3) +
+  theme_minimal() +
+  scale_x_continuous(limits = c(0.0,1), expand = c(0, 0), breaks = c(0.0, 0.25 ,0.5, 0.75, 1.0)) +
+  labs(title='', y=TeX("$PCC$"), x=TeX("$m$"), color=TeX("$\\alpha$")) + 
+  thm + 
+  facet_grid(vars(sigma), vars(lambda)) +
+  coord_cartesian(xlim = c(0.0,1.0), ylim=c(0,1)) 
+
+convert_to_tex_label = function(string) {
+  return(TeX(string))
+}
+
+plt=data %>% 
+  group_by(treatment) %>%
+  mutate(mean_pcc = mean(summary_stat)) %>%
+  mutate(lowest_pcc = quantile(summary_stat, probs=c(0.025))) %>%
+  mutate(lower_pcc = quantile(summary_stat, probs=c(0.25))) %>%
[…]

### 2020-08-28 figs/figure6_even_mixing/figure6.jl

@@ -1,0 +1,47 @@
+
+# Project Folder
+proj_folder = "/home/michael/phase_transitions_in_metapopulation_synchrony/"
+cd(proj_folder)
+
+include(string(proj_folder, "src/deps.jl"))
+
+output_dir_path = string(proj_folder , "figs/figure6_even_mixing/output/")
+mkdir(output_dir_path)
+# _______________________________________________________________________________________
+#
+#
+#   phase transition diagram for 20 populations, varying alpha
+#
+#
+#
+#
+#
+# _______________________________________________________________________________________
+
+param_dictionary = Dict(
+                        "migration_rate"    => [i for i = 0.0:0.01:1.0],
+                        "lambda"            => [1.0],
+                        "sigma"             => [1.0],
+                        "carrying_capacity" => [1.0],
+                        "num_populations" => collect(2:25),
+                        "alpha"           => [0.0],
+                        # _________________________________________________
+                        #
+                        # simulation parameters
+                        #
+                        # _________________________________________________
+                        "number_of_timesteps" => [300],
+                        "metapopulation_generator" => [get_random_metapopulation]
+                    )
+
+
+
+treatment_set = create_treatments(param_dictionary, replicates_per_treatment = 50)
+
+df = run_treatments(treatment_set)
+
+CSV.write(string(output_dir_path, "treatment_set.csv"), df)
+CSV.write(string(output_dir_path, "metadata.csv"), treatment_set.metadata)
+
+
+

### 2020-08-28 figs/figure6_even_mixing/figure6_vis.R

@@ -1,0 +1,26 @@
+setwd("~/phase_transitions_in_metapopulation_synchrony/figs/figure6_even_mixing/output")
+library(tidyverse)
+library(ggthemr)
+library(latex2exp)
+ggthemr('fresh', spacing=3)
+cmuserif = p$`CMU Serif`$family
+thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
+            text=element_text(family=cmuserif,size=16), axis.title.y = element_text(angle = 0, hjust=1))
+
+data = read.csv('metadata.csv') %>% full_join(read.csv('treatment_set.csv'), by="treatment")
+
+
+
+data %>% 
+  group_by(treatment) %>%
+  mutate(mean_pcc = mean(summary_stat)) %>%
+  ungroup %>%
+  group_by(num_populations) %>% 
+  filter(mean_pcc==max(mean_pcc)) %>%
+  ggplot(aes(num_populations,migration_rate))  + 
+  stat_function(fun = function(x) 1.0 - 1.0/x) + 
+  geom_point(size=2) 
+  
+#+ 
+ # coord_cartesian(xlim = c(0.0,1.0)) 
+

### 2020-08-28 figs/figure7_expected_num_pops/figure7.jl

@@ -1,0 +1,47 @@
+
+# Project Folder
+proj_folder = "/home/michael/phase_transitions_in_metapopulation_synchrony/"
+cd(proj_folder)
+
+include(string(proj_folder, "src/deps.jl"))
+
+output_dir_path = string(proj_folder , "figs/figure7_expected_num_pops/output/")
+mkdir(output_dir_path)
+# _______________________________________________________________________________________
+#
+#
+#   phase transition diagram for 20 populations, varying alpha
+#
+#
+#
+#
+#
+# _______________________________________________________________________________________
+
+param_dictionary = Dict(
+                        "migration_rate"    => [i for i = 0.0:0.01:1.0],
+                        "lambda"            => [1.0],
+                        "sigma"             => [1.0],
+                        "carrying_capacity" => [1.0],
+                        "num_populations" => [15],
+                        "alpha"           => [i for i = 0.0:2:20],
+                        # _________________________________________________
+                        #
+                        # simulation parameters
+                        #
+                        # _________________________________________________
+                        "number_of_timesteps" => [300],
+                        "metapopulation_generator" => [get_random_metapopulation]
+                    )
+
+
+
+treatment_set = create_treatments(param_dictionary, replicates_per_treatment = 200)
+
+df = run_treatments(treatment_set)
+
+CSV.write(string(output_dir_path, "treatment_set.csv"), df)
+CSV.write(string(output_dir_path, "metadata.csv"), treatment_set.metadata)
+
+
+

### 2020-08-28 figs/figure7_expected_num_pops/figure7_vis.R

@@ -1,0 +1,54 @@
+setwd("~/phase_transitions_in_metapopulation_synchrony/figs/figure7_expected_num_pops/output")
+library(tidyverse)
+library(ggthemr)
+library(latex2exp)
+ggthemr('fresh', spacing=3)
+cmuserif = p$`CMU Serif`$family
+thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
+            panel.spacing=unit(3, "lines"),
+            text=element_text(family=cmuserif,size=20), axis.title.y = element_text(angle = 0, vjust=0.5),
+            axis.title = element_text(margin = margin(t = 20, r = 20, b = 0, l = 0)),
+            axis.text = element_text(size=14),
+            # The new stuff
+            strip.text = element_text(size = 22),
+            strip.text.y = element_text(size = 22, angle = 0, vjust=0.5))
+
+data = read.csv('metadata.csv') %>% full_join(read.csv('treatment_set.csv'), by="treatment")
+
+compute_exp_pops = function(m_max){
+  np = (1.0)/(1.0 - m_max)
+  return(np)
+}
+
+data %>% 
+  group_by(treatment) %>%
+  filter(alpha==10) %>%
+  mutate(mean_pcc = mean(summary_stat)) %>%
+  ggplot(aes(migration_rate, mean_pcc)) + geom_point()
+
+data %>% 
+  group_by(treatment) %>%
+  mutate(mean_pcc = mean(summary_stat)) %>%
+  ungroup %>%
+  group_by(alpha) %>% 
+  filter(abs(mean_pcc - max(mean_pcc)) < 0.01) %>%
+  mutate(expected_num_pops = compute_exp_pops(migration_rate)) %>%
+  mutate(mean_critical_migration=mean(migration_rate)) %>%
+  mutate(expected_num_pops = compute_exp_pops(mean_critical_migration)) %>%
+  ggplot(aes(alpha, expected_num_pops))  + 
+  geom_point() +
+  geom_hline(aes(yintercept=15), linestyle='dashed') + scale_y_continuous(limits=c(0,18))
+
+exp_data = data %>% 
+  group_by(treatment) %>%
+  mutate(mean_pcc = mean(summary_stat)) %>%
+  ungroup %>%
+  group_by(alpha) %>% 
+  filter(abs(mean_pcc - max(mean_pcc)) < 0.001) %>%
[…]

### 2020-08-28 figs/network_vis.R

@@ -1,0 +1,131 @@
+elibrary(ggplot2)
+library(ggnet)
+library(gridExtra)
+library(tikzDevice)
+library(tidyverse)
+library(extrafont)
+library(latex2exp)
+library(Cairo)
+setwd("~/phase_transitions_in_metapopulation_synchrony/figs")
+loadfonts()
+
+data(mtcars)
+mtcars %>% ggplot(aes(x=mpg, y=disp)) + geom_col(aes(col=cyl)) + 
+    theme(text  = element_text(size=10, family=p$`CMU Serif`$family)) + ggtitle("for the boys")
+
+
+cmuserif = p$`CMU Serif`$family
+## ============================================ 
+## DISPERSAL KERNEL PLOTTING / FIG 1
+## ============================================
+exp_kern = function(alpha, d_ij){
+    return(exp(-1*alpha*d_ij))
+}
+
+gauss_kern = function(alpha, d_ij){
+    return(exp(-1*d_ij^2*alpha^2))
+}
+
+
+get_pop_points = function(x_vals, y_vals, alpha, mig_prop, kernel=exp_kern, kernel_name="Exp.", xaxis=F, yaxis=F){
+    # make df with all pairwise and group by indiv pairwise val
+    df = data.frame(matrix(ncol=5))
+    
+    line_ct = 1
+    grp_ct = 0 
+    
+    # make a dis potential first
+    
+    npops = length(x_vals)
+    
+    dispersal_potential = matrix(nrow=npops, ncol=npops)
+    for (pt_ct in seq(1, n_pops)){
+        s = 0
+        for (pt_ct2 in seq(1, n_pops)){
+            x1 = x_vals[pt_ct] 
+            y1 = y_vals[pt_ct] 
+            x2 = x_vals[pt_ct2] 
+            y2 = y_vals[pt_ct2] 
+            
+            dist = sqrt((x2-x1)^2 + (y2-y1)^2)
+            
+            dispersal_potential[pt_ct, pt_ct2] = kernel(alpha, dist)
+            
+            s = s + dispersal_potential[pt_ct, pt_ct2] 
+        }
+        for (pt_ct2 in seq(1, n_pops)){
+            dispersal_potential[pt_ct, pt_ct2]  = dispersal_potential[pt_ct, pt_ct2]  / s
+        }
+    }
+    
+    for (pt_ct in seq(1, n_pops)){
[…]

### 2020-08-28 src/TreatmentFactory.jl

@@ -59,24 +59,20 @@ end
 function create_treatments(param_dict::Dict; replicates_per_treatment::Int64 = 50)
 	metadata::DataFrame = create_metadata_df(param_dict)
 	n_treatments::Int64 = nrow(metadata)
-	
+
 	treatments::Vector{Treatment} = []
 
 	# Treatment(mp, dynamics_model, sim_params, param_bundle, param_values, PCC(), log, abundance_matrix, initial_condition)
 	for t = 1:n_treatments
 		num_pops = metadata.num_populations[t]
 		alpha = metadata.alpha[t]
-	
-		migration_rate_distribution::Distribution = get_dist(metadata.migration_rate_distribution[t])
-		lambda_distribution::Distribution = get_dist(metadata.lambda_distribution[t])
-		sigma_distribution::Distribution = get_dist(metadata.sigma_distribution[t])
-		carrying_capacity_distribution::Distribution = get_dist(metadata.carrying_capacity_distribution[t])
 
-		dimensionality = num_pops
+		migration_rate_distribution::Distribution = get_dist(metadata.migration_rate[t])
+		lambda_distribution::Distribution = get_dist(metadata.lambda[t])
+		sigma_distribution::Distribution = get_dist(metadata.sigma[t])
+		carrying_capacity_distribution::Distribution = get_dist(metadata.carrying_capacity[t])
 
-		if (metadata.parameters_fixed_across_populations[t] == 1)
-			dimensionality = 1
-		end
+        dimensionality::Int64 = num_pops
 
 		alpha_dist = Parameter(Normal(alpha, 0.01), 1)
 		m_dist = Parameter(migration_rate_distribution, dimensionality)
@@ -84,17 +80,12 @@ function create_treatments(param_dict::Dict; replicates_per_treatment::Int64 = 5
 		sigma_dist = Parameter(sigma_distribution, dimensionality)
 		carrying_cap_dist = Parameter(carrying_capacity_distribution, dimensionality)
 
-
 		param_bundle = StochasticLogisticParameterBundle(num_pops, alpha, m_dist, lambda_dist,sigma_dist, carrying_cap_dist)
-
-
-#		param_values = draw_from_parameter_bundle(param_bundle)
 
 		dynamics_model = StochasticLogisticWDiffusion()
 		sim_params = SimulationParameters(metadata.number_of_timesteps[t], 0.1, 10, false)
-		
[…]

### 2020-08-28 src/types.jl

@@ -53,7 +53,7 @@ struct StochasticLogisticParameterValues <: ParameterValues
     sigma::Vector{Float64} # sigma across pops
 end
 
-Base.show(io::IO, params::StochasticLogisticParameterValues) = print(io, 
+Base.show(io::IO, params::StochasticLogisticParameterValues) = print(io,
                                                                     "Stochastic Logistic Model with Parameters: \n",
                                                                     "\t\t\talpha:", params.alpha, "\n",
                                                                     "\t\t\tlambda:", params.lambda, "\n",
@@ -87,10 +87,10 @@ mutable struct Metapopulation
     populations::Vector{Population}
     dispersal_potential::DispersalPotential
 end
-Base.show(io::IO, metapopulation::Metapopulation) = print(io, "Metapopulation with ", 
-metapopulation.num_populations, 
-" populations and alpha = ", 
-metapopulation.dispersal_potential.alpha,  
+Base.show(io::IO, metapopulation::Metapopulation) = print(io, "Metapopulation with ",
+metapopulation.num_populations,
+" populations and alpha = ",
+metapopulation.dispersal_potential.alpha,
 "\n",
 " Coordinates: ", [metapopulation.populations[p].coordinate for p = 1:metapopulation.num_populations]
 )
@@ -139,10 +139,10 @@ mutable struct Treatment
     simulation_parameters::SimulationParameters
     theta::ParameterBundle
     summary_stat::Function
-    instances::Vector{TreatmentInstance} 
+    instances::Vector{TreatmentInstance}
 end
 
-Base.show(io::IO, treatment::Treatment) = print(io, "\n\n\nTreatment\n", 
+Base.show(io::IO, treatment::Treatment) = print(io, "\n\n\nTreatment\n",
                                                     "--------------------------------------------\n",
                                                     "\tdX/dt : ", (treatment.dx_dt), "\n",
                                                     )

### 2020-08-29 writing/manuscript.tex
critical pig

@@
 
-\[P(V_j|V_i)=\frac{f(d_{ij}, \alpha)}{\sum_k f(d_{ik},\alpha)}\]
+\[\Phi_{ij} = P(V_j|V_i)=\frac{f(d_{ij}, \alpha)}{\sum_k f(d_{ik},\alpha)}\]
 
@@
 
-\[dN_i=\lambda_i N_i \big(1 - \frac{N_i}{K_i}\big)dt+\sigma_pN_idW\]
+\[dN_i=\lambda_i N_i \bigg(1 - \frac{N_i}{K_i}\bigg)dt+\sigma N_idW\]
 
@@
 deviation in abundance due to local stochasticity as a proportion of
-\(N_i\). Here \(\sigma_p\) represents an amalgamation of all factors
+\(N_i\). Here \(\sigma\) represents an amalgamation of all factors
 contributing to local stochasticity in population dynamics, although it
@@
 local stochasticity can drive significant variation in dynamics
-\cite{melbourne_extinction_2008}. For the sake of reducing the size of parameter
+\cite{melbourne_extinction_2008}.
+
+We can write this system as the matrix equation
+\begin{equation}
+        d\boldsymbol{N} = \boldsymbol{I} \boldsymbol{\lambda} \bigg(\boldsymbol{I}(\boldsymbol{K} - \boldsymbol{N})\boldsymbol{K}^{-1} \bigg) d\boldsymbol{t} + \boldsymbol{I} \boldsymbol{\sigma} \boldsymbol{N} d\boldsymbol{W}
+\end{equation}
+
+For the sake of reducing the size of parameter
 space, here we consider all populations as having the same \(\lambda_i\)
@@
 transitions in stochastic systems before \cite{brock_early_2012, kuehn_mathematical_2011}, and they have many nice properties for inference. SDEs
-have been used to study extinction dynamics before, as it is relatively
-straightforward to compute the mean time until extinction (MTE) using
-the Kolmogorov Backward Equation (@lande\_stochastic\_2003).
+have also been used in ecology to study extinction dynamics before, as it is relatively straightforward to compute the mean time until extinction (MTE) using
+the Kolmogorov Backward Equation \cite{lande_stochastic_2003}.
 
 \hypertarget{diffusion-on-spatial-graphs}{%
-\subsubsection{Diffusion on Spatial
+\subsection{Diffusion on Spatial
 Graphs}\label{diffusion-on-spatial-graphs}}
@@
 landscapes are inherently stochastic. The probability that an individual
-migrates within its lifetime, \(p_m\), and where it goes, \(P(V_j|V_i)\)
-, are both stochastic processes. Under some conditions, we can
+migrates within its lifetime, $m$, and where it goes, $\Phi_{ij}$, are both stochastic . Under some conditions, we can
 effectively model stochastic dynamics across space using diffusion
@@
 dispersal. Diffusion models have seen widespread use in ecology and
-other fields (@ovaskainen\_empirical\_2008,
-@holmes\_partial-differential\_1994, @okubo\_diffusion\_2011).
+other fields \cite{ovaskainen_empirical_2008, holmes_partial-differential_1994, okubo_diffusion_2011}.
 
-Given our dispersal potential \(P(V_j|V_i)\), we can define the
-diffusion matrix \(\Phi\) as
+Given our dispersal potential $\Phi_{ij}$, we can define the
+diffusion matrix \( D \) as
 
-\[\Phi_{ij} = \begin{cases} 1 - p_m \quad\quad &\text{if}\ i = j \\ P(V_j|V_i)p_m & \text{if}\ i \neq j\end{cases}\]
+\[D_{ij} = \begin{cases} 1 - m \quad\quad &\text{if}\ i = j \\ \Phi_{ij}m & \text{if}\ i \neq j\end{cases}\]
 
-where \(\Phi_{ij}\) represents the probability that any individual born
-in \(V_i\) reproduces in \(V_j\).
-
-Now we move to considering the dynamics of the abundance of the
-population at \(V_i\), denoted \(N_i\). We can now represent the
+We can now represent the
 dynamics due to diffusion of this system as
 
-\[\dot{N_i}=(1-p_m)N_i+ \sum_j p_mP(V_j|V_i)N_j\]
+\[\dot{N_i}=(1-m)N_i+ \sum_j \Phi_{ij}mN_j\]
 
@@
 
-\[\frac{d\vec{N}}{dt}=\Phi^T\vec{N}\]
+\[\frac{d\vec{N}}{dt}=D^T\vec{N}\]
 
@@
 
-\[\frac{d\vec{N}}{dt} = g(\Phi^T \vec{N})\]
+\[\frac{d\vec{N}}{dt} = g(D^T \vec{N})\]
 
@@
 \begin{equation} \label{diffusion_model}
-  d \vec{N} = {I} (\Phi^T \vec{N})(\vec{k}-\vec{N}) dt + \sigma_p \vec{k} \ d\vec{W}
+  d \boldsymbol{N} = \boldsymbol{I} (D^T \boldsymbol{N})\bigg(\boldsymbol{I}\boldsymbol{K}^{-1}(\boldsymbol{K}-D^T \boldsymbol{N})\bigg) d\boldsymbol{t} + \boldsymbol{I}\boldsymbol{\sigma} \boldsymbol{N} \ d\boldsymbol{W}
 \end{equation}
@@
 crosscorrelation function, \(CC\), has long been used as a measure of
-synchronous dynamics (@liebhold\_spatial\_2004). Here, with a subdivided
+synchronous dynamics \cite{liebhold_spatial_2004}. Here, with a subdivided
 population, we consider the mean crosscorrelation compared across all
@@
 \begin{figure}[h]
-\caption{The abundance of 5 populations (each in a different color) across time.
-Panel (a):  These populations show low synchrony, $PCC=0.1$.
-Panel (b): These populations show high sychrony, $PCC=0.89$ }
-
+\includegraphics[width=15cm]{./figs/figure2}
+\caption{The abundance of 2 populations (top) and 8 populations (bottom) across time. Get the values of synchrony }
 \label{async_and_sync}
@@
 \pagebreak
+\begin{figure}
+    \includegraphics[width=15cm]{figs/figure3}
+    \caption{}
+    \label{}
+\end{figure}
 
 \begin{figure}
-    \includegraphics[width=15cm]{figs/figure4.pdf}
+    \includegraphics[width=15cm]{figs/figure4}
+    \caption{}
+    \label{}
+\end{figure}
+
+\begin{figure}
+    \includegraphics[width=15cm]{figs/figure5}
     \caption{}

### 2020-08-29 figs/figure1_metapopulations_as_a_network/figure1.R

@@ -1,4 +1,4 @@
-elibrary(ggplot2)
+library(ggplot2)
 library(ggnet)
 library(gridExtra)
 library(tikzDevice)
@@ -10,12 +10,6 @@ library(extrafont)
 loadfonts()
 setwd("~/phase_transitions_in_metapopulation_synchrony/figs")
 loadfonts()
-
-data(mtcars)
-mtcars %>% ggplot(aes(x=mpg, y=disp)) + geom_col(aes(col=cyl)) + 
-    theme(text  = element_text(size=10, family=p$`CMU Serif`$family)) + ggtitle("for the boys")
-
-
 cmuserif = p$`CMU Serif`$family
 ## ============================================ 
 ## DISPERSAL KERNEL PLOTTING / FIG 1
@@ -116,18 +110,23 @@ get_pop_points = function(x_vals, y_vals, alpha, mig_prop, kernel=exp_kern, kern
 
 n_pops = 20
 
+set.seed(5)
+
 x_vals = runif(n_pops, 0, 1)    
 y_vals = runif(n_pops, 0, 1)
 
 plt1 = get_pop_points(x_vals, y_vals, 2, 1.5, kernel=exp_kern,yaxis=T)
-plt2 = get_pop_points(x_vals, y_vals, 6, 1.5, kernel=exp_kern, yaxis=F, xaxis=F)
-plt3 = get_pop_points(x_vals, y_vals, 10, 1.5, kernel=exp_kern, yaxis=F, xaxis=F)
+plt2 = get_pop_points(x_vals, y_vals, 7, 1.5, kernel=exp_kern, yaxis=F, xaxis=F)
+plt3 = get_pop_points(x_vals, y_vals, 12, 1.5, kernel=exp_kern, yaxis=F, xaxis=F)
 plt4 = get_pop_points(x_vals, y_vals, 2, 1.5, kernel=gauss_kern, yaxis=T,xaxis=T,kernel_name="Gauss")
-plt5 = get_pop_points(x_vals, y_vals, 6, 1.5, kernel=gauss_kern, xaxis=T,kernel_name="Gauss")
-plt6 = get_pop_points(x_vals, y_vals, 10, 1.5, kernel=gauss_kern, xaxis=T, yaxis=F, kernel_name="Gauss")
+plt5 = get_pop_points(x_vals, y_vals, 7, 1.5, kernel=gauss_kern, xaxis=T,kernel_name="Gauss")
[…]

### 2020-08-29 figs/figure2_synchrony_example/figure2.jl

@@ -1,0 +1,49 @@
+
+# Project Folder
+proj_folder = "/home/michael/phase_transitions_in_metapopulation_synchrony/"
+cd(proj_folder)
+
+include(string(proj_folder, "./src/deps.jl"))
+output_dir_path = string(proj_folder , "figs/figure2_synchrony_example/output/")
+mkdir(output_dir_path)
+
+
+# _______________________________________________________________________________________
+#
+#
+#   phase transition diagram for two populations
+#
+#
+#
+#
+#
+# _______________________________________________________________________________________
+
+param_dictionary = Dict(
+                        "migration_rate"    => [0.1, 0.3, 0.7],
+                        "lambda"            => [1.0],
+                        "sigma"             => [0.5],
+                        "carrying_capacity" => [1.0],
+                        "num_populations" => [2, 8],
+                        "alpha"           => [0],  # alpha is meaningless with two pops
+
+                        # _________________________________________________
+                        #
+                        # simulation parameters
+                        #
+                        # _________________________________________________
+                        "number_of_timesteps" => [300],
+                        "metapopulation_generator" => [get_random_metapopulation],
+                        "fixed_metapopulation" => [false],
+                        "log_abundances" => [true]
+                    )
+
+
+
+treatment_set = create_treatments(param_dictionary, replicates_per_treatment = 50)
+
+df = run_treatments(treatment_set)
+
+
+CSV.write(string(output_dir_path, "treatment_set.csv"), df)
+CSV.write(string(output_dir_path, "metadata.csv"), treatment_set.metadata)

### 2020-08-29 figs/figure2_synchrony_example/figure2_vis.R

@@ -1,0 +1,40 @@
+setwd("~/phase_transitions_in_metapopulation_synchrony/figs/figure2_synchrony_example/output")
+library(tidyverse)
+library(ggthemr)
+library(latex2exp)
+library(extrafont)
+loadfonts()
+ggthemr('fresh', spacing=3)
+cmuserif = pdfFonts()$`CMU Serif`$family
+thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
+            panel.spacing=unit(3, "lines"),
+            text=element_text(family=cmuserif,size=20), axis.title.y = element_text(angle = 0, vjust=0.5),
+            axis.title = element_text(margin = margin(t = 20, r = 20, b = 0, l = 0)),
+            axis.text = element_text(size=14),
+            panel.grid = element_line(size=0.5, linetype="dotted"),
+            # The new stuff
+            strip.text = element_text(size = 22),
+            strip.text.y = element_text(size = 22, angle = 0, vjust=0.5))
+
+
+data = read.csv('metadata.csv') %>% full_join(read.csv('abundances.csv'), by="treatment")
+
+convert_to_tex_label = function(string) {
+  return(TeX(string))
+}
+
+plt = data %>% 
+  mutate(num_pops_facet=paste("$N_p = ", num_populations, "$")) %>%
+  mutate(migration_facet=paste("$m = ", migration_rate, "$")) %>%
+  ggplot(aes(timestep, abundance, group=interaction(replicate,population), color=factor(population))) + 
+    geom_line(alpha=(0.7), size=1.3) + 
+    facet_grid(vars(num_pops_facet), vars(migration_facet), labeller = as_labeller(convert_to_tex_label, label_parsed)) +
+    labs(title='', y=TeX("$Abundance$"), x=TeX("$m$"), color="Population") +
+    geom_hline(aes(yintercept=1.0) ,linetype="dashed", color='black') + 
+    thm
+
+
+
+output_path = "~/phase_transitions_in_metapopulation_synchrony/writing/figs/figure2.png"
+ggsave(output_path, plot=plt, dpi=320, width = 18, height = 10, units = "in", device=png())
+

### 2020-08-29 figs/figure3_2population_phase_transition/figure3.jl

@@ -3,7 +3,7 @@
 proj_folder = "/home/michael/phase_transitions_in_metapopulation_synchrony/"
 cd(proj_folder)
 
-include("./src/deps.jl")
+include(string(proj_folder, "./src/deps.jl"))
 output_dir_path = string(proj_folder , "figs/figure3_2population_phase_transition/output/")
 
 
@@ -33,7 +33,9 @@ param_dictionary = Dict(
                         #
                         # _________________________________________________
                         "number_of_timesteps" => [300],
-                        "metapopulation_generator" => [get_random_metapopulation]
+                        "metapopulation_generator" => [get_random_metapopulation],
+                        "fixed_metapopulation" => [false],
+                        "log_abundances" => [false]
                     )
 
 
@@ -46,6 +48,3 @@ mkdir(output_dir_path)
 
 CSV.write(string(output_dir_path, "treatment_set.csv"), df)
 CSV.write(string(output_dir_path, "metadata.csv"), treatment_set.metadata)
-
-
-

### 2020-08-29 figs/figure3_2population_phase_transition/vis.R

@@ -5,7 +5,7 @@ library(latex2exp)
 library(extrafont)
 loadfonts()
 ggthemr('fresh', spacing=2)
-cmuserif = p$`CMU Serif`$family
+cmuserif = pdfFonts()$`CMU Serif`$family
 thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
             panel.spacing=unit(3, "lines"),
             text=element_text(family=cmuserif,size=20), axis.title.y = element_text(angle = 0, vjust=0.5),
@@ -37,7 +37,7 @@ convert_to_tex_label = function(string) {
 
 
 
-data %>% 
+plt = data %>% 
   group_by(treatment) %>%
   mutate(mean_pcc = mean(summary_stat)) %>%
   mutate(lowest_pcc = quantile(summary_stat, probs=c(0.025))) %>%
@@ -56,5 +56,10 @@ data %>%
   labs(title='', y=TeX("$PCC$"), x=TeX("$m$"), color=TeX("$\\sigma$"), fill=TeX("$\\sigma$")) +
   facet_grid(vars(sigma_facet), vars(lambda_facet), labeller =as_labeller(convert_to_tex_label, label_parsed)) + thm
  
+
+output_path = "~/phase_transitions_in_metapopulation_synchrony/writing/figs/figure3.png"
+ggsave(output_path, plot=plt, dpi=320, width = 12, height = 12, units = "in", device=png())
+
+
  # coord_cartesian(xlim = c(0.0,1.0)) 
 

### 2020-08-29 figs/figure4_20population_phase_transition_varying_alpha/figure4.jl

@@ -24,7 +24,7 @@ param_dictionary = Dict(
                         "sigma"             => [0.5 ],
                         "carrying_capacity" => [1.0],
                         "num_populations" => [20],
-                        "alpha"           => [0, 5, 10, 15],
+                        "alpha"           => [0, 10, 20, 30],
                         # _________________________________________________
                         #
                         # simulation parameters

### 2020-08-29 figs/figure4_20population_phase_transition_varying_alpha/figure4_vis.R

@@ -57,7 +57,7 @@ plt = data %>%
   thm + 
   coord_cartesian(xlim = c(0.0,1.0), ylim=c(0.0,1.0)) 
 
-output_path = "~/phase_transitions_in_metapopulation_synchrony/writing/figs/figure4.pdf"
-ggsave(output_path, plot=plt, dpi=320, width = 12, height = 8, units = "in", device=cairo_pdf)
+output_path = "~/phase_transitions_in_metapopulation_synchrony/writing/figs/figure4.png"
+ggsave(output_path, plot=plt, dpi=320, width = 12, height = 8, units = "in", device=png())
 
 

### 2020-08-29 figs/figure5_20population_phase_transition_varying_sigma_and_lambda/figure5_vis.R

@@ -5,7 +5,7 @@ library(latex2exp)
 library(extrafont)
 loadfonts()
 ggthemr('fresh', spacing=3)
-cmuserif = p$`CMU Serif`$family
+cmuserif = pdfFonts()$`CMU Serif`$family
 thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
             panel.spacing=unit(3, "lines"),
             text=element_text(family=cmuserif,size=20), axis.title.y = element_text(angle = 0, vjust=0.5),
@@ -34,6 +34,7 @@ convert_to_tex_label = function(string) {
   return(TeX(string))
 }
 
+
 plt=data %>% 
   group_by(treatment) %>%
   mutate(mean_pcc = mean(summary_stat)) %>%
@@ -54,12 +55,36 @@ plt=data %>%
   facet_grid(vars(sigma_facet), vars(lambda_facet), labeller = as_labeller(convert_to_tex_label, label_parsed)) +
   thm
 
+output_path = "~/phase_transitions_in_metapopulation_synchrony/writing/figs/figure5.png"
+ggsave(output_path, plot=plt, dpi=320, width = 10, height = 10, units = "in", device=png)
+
+get_logit_intercept = function(data, pcc, mig){
+ m = glm(formula=pcc ~ mig, data=data, family="quasibinomial")
+ return(m$coefficients[1])
+}
+
+get_logit_slope = function(data, pcc, mig ){
+  m = glm(formula=pcc ~ mig, data=data, family="quasibinomial")
+  return(m$coefficients[2])
+}
+
+data %>%
+  group_by(treatment) %>%
+  mutate(mean_pcc = mean(summary_stat)) %>%
+  ungroup() %>% 
+  group_by(lambda, sigma) %>%
+  ggplot(aes(migration_rate, mean_pcc)) + facet_grid(vars(lambda), vars(sigma)) + geom_point()
+  
+data %>%
+  group_by(treatment) %>%
+  mutate(mean_pcc = mean(summary_stat)) %>%
+  ungroup() %>% 
+  group_by(lambda, sigma) %>%
+  mutate(intercept=get_logit_intercept(., mean_pcc, migration_rate)) %>%
+  mutate(slope=get_logit_slope(., mean_pcc, migration_rate)) %>% 
+  mutate(logit_fcn=(1.0)/(1.0+exp(-1*slope*migration_rate + intercept))) %>%
[…]

### 2020-08-29 figs/figure6_even_mixing/figure6.jl

@@ -24,7 +24,7 @@ param_dictionary = Dict(
                         "sigma"             => [1.0],
                         "carrying_capacity" => [1.0],
                         "num_populations" => collect(2:25),
-                        "alpha"           => [0.0],
+                        "alpha"           => [0.0, 5.0, 10.0, 20.0],
                         # _________________________________________________
                         #
                         # simulation parameters

### 2020-08-29 figs/figure6_even_mixing/figure6_vis.R

@@ -3,24 +3,40 @@ library(tidyverse)
 library(ggthemr)
 library(latex2exp)
 ggthemr('fresh', spacing=3)
-cmuserif = p$`CMU Serif`$family
+cmuserif = pdfFonts()$`CMU Serif`$family
 thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
-            text=element_text(family=cmuserif,size=16), axis.title.y = element_text(angle = 0, hjust=1))
-
+            panel.spacing=unit(3, "lines"),
+            text=element_text(family=cmuserif,size=20), axis.title.y = element_text(angle = 0, vjust=0.5),
+            axis.title = element_text(margin = margin(t = 20, r = 20, b = 0, l = 0)),
+            axis.text = element_text(size=14),
+            panel.grid = element_line(size=0.5, linetype="dotted"),
+            # The new stuff
+            strip.text = element_text(size = 22),
+            strip.text.y = element_text(size = 22, angle = 0, vjust=0.5))
 data = read.csv('metadata.csv') %>% full_join(read.csv('treatment_set.csv'), by="treatment")
 
 
 
 data %>% 
-  group_by(treatment) %>%
-  mutate(mean_pcc = mean(summary_stat)) %>%
-  ungroup %>%
-  group_by(num_populations) %>% 
-  filter(mean_pcc==max(mean_pcc)) %>%
-  ggplot(aes(num_populations,migration_rate))  + 
+  group_by(num_populations, replicate, alpha) %>% 
+  filter(alpha==0) %>%
+  filter(abs(max(summary_stat)-summary_stat) < 0.01) %>%
+  mutate(low = quantile(migration_rate,0.25), high=quantile(migration_rate, 0.75)) %>%
+  mutate(lowest = quantile(migration_rate,0.025), highest=quantile(migration_rate, 0.975)) %>%
+  ggplot(aes(num_populations,migration_rate, fill=factor(alpha)))  + 
+  geom_ribbon(aes(ymin=lowest, ymax=highest), alpha=0.3) +
+  geom_ribbon(aes(ymin=low, ymax=high), alpha=0.4) +
   stat_function(fun = function(x) 1.0 - 1.0/x) + 
-  geom_point(size=2) 
+  labs(x=TeX("$N_p$"), y=TeX("$m$"))  + 
[…]

### 2020-08-29 figs/figure7_expected_num_pops/figure7.jl

@@ -23,7 +23,7 @@ param_dictionary = Dict(
                         "lambda"            => [1.0],
                         "sigma"             => [1.0],
                         "carrying_capacity" => [1.0],
-                        "num_populations" => [15],
+                        "num_populations" => [5, 10, 15, 20, 25],
                         "alpha"           => [i for i = 0.0:2:20],
                         # _________________________________________________
                         #
@@ -36,12 +36,9 @@ param_dictionary = Dict(
 
 
 
-treatment_set = create_treatments(param_dictionary, replicates_per_treatment = 200)
+treatment_set = create_treatments(param_dictionary, replicates_per_treatment = 50)
 
 df = run_treatments(treatment_set)
 
 CSV.write(string(output_dir_path, "treatment_set.csv"), df)
 CSV.write(string(output_dir_path, "metadata.csv"), treatment_set.metadata)
-
-
-

### 2020-08-29 figs/figure7_expected_num_pops/figure7_vis.R

@@ -1,9 +1,10 @@
 setwd("~/phase_transitions_in_metapopulation_synchrony/figs/figure7_expected_num_pops/output")
 library(tidyverse)
+library(gridExtra)
 library(ggthemr)
 library(latex2exp)
 ggthemr('fresh', spacing=3)
-cmuserif = p$`CMU Serif`$family
+cmuserif = pdfFonts()$`CMU Serif`$family
 thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
             panel.spacing=unit(3, "lines"),
             text=element_text(family=cmuserif,size=20), axis.title.y = element_text(angle = 0, vjust=0.5),
@@ -20,33 +21,21 @@ compute_exp_pops = function(m_max){
   return(np)
 }
 
+
 data %>% 
-  group_by(treatment) %>%
-  filter(alpha==10) %>%
-  mutate(mean_pcc = mean(summary_stat)) %>%
-  ggplot(aes(migration_rate, mean_pcc)) + geom_point()
+  group_by(num_populations, alpha) %>% 
+  filter(max(summary_stat)==summary_stat) %>%
+  mutate(exp_num_pops = compute_exp_pops(migration_rate)/num_populations) %>% 
+  ggplot(aes(alpha,exp_num_pops, color=factor(num_populations)),) + geom_point() + geom_line() + ylim(0,2)
+  
+  
+
 
 data %>% 
   group_by(treatment) %>%
   mutate(mean_pcc = mean(summary_stat)) %>%
-  ungroup %>%
-  group_by(alpha) %>% 
-  filter(abs(mean_pcc - max(mean_pcc)) < 0.01) %>%
-  mutate(expected_num_pops = compute_exp_pops(migration_rate)) %>%
-  mutate(mean_critical_migration=mean(migration_rate)) %>%
-  mutate(expected_num_pops = compute_exp_pops(mean_critical_migration)) %>%
-  ggplot(aes(alpha, expected_num_pops))  + 
-  geom_point() +
-  geom_hline(aes(yintercept=15), linestyle='dashed') + scale_y_continuous(limits=c(0,18))
-
-exp_data = data %>% 
-  group_by(treatment) %>%
-  mutate(mean_pcc = mean(summary_stat)) %>%
-  ungroup %>%
-  group_by(alpha) %>% 
-  filter(abs(mean_pcc - max(mean_pcc)) < 0.001) %>%
[…]

### 2020-08-29 figs/figure8_sigma_gradient_single_panel/figure8.jl

@@ -1,0 +1,46 @@
+
+# Project Folder
+proj_folder = "/home/michael/phase_transitions_in_metapopulation_synchrony/"
+cd(proj_folder)
+
+include(string(proj_folder, "./src/deps.jl"))
+
+output_dir_path = string(proj_folder , "figs/figure8_sigma_gradient_single_panel/output/")
+mkdir(output_dir_path)
+# _______________________________________________________________________________________
+#
+#
+#   phase transition diagram for two populations
+#
+#
+#
+#
+#
+# _______________________________________________________________________________________
+
+param_dictionary = Dict(
+                        "migration_rate"    => [i for i = 0.0:0.01:1.0],
+                        "lambda"            => [1.0],
+                        "sigma"             => [0.1, 0.3, 0.5, 0.7, 0.9, 1.2],
+                        "carrying_capacity" => [2.0],
+                        "num_populations" => [20],
+                        "alpha"           => [0],
+                        # _________________________________________________
+                        #
+                        # simulation parameters
+                        #
+                        # _________________________________________________
+                        "number_of_timesteps" => [300],
+                        "metapopulation_generator" => [get_random_metapopulation],
+                        "fixed_metapopulation" => [false],
+                        "log_abundances" => [true]
+                    )
+
+
+
+treatment_set = create_treatments(param_dictionary, replicates_per_treatment = 50)
+
+df = run_treatments(treatment_set)
+
+CSV.write(string(output_dir_path, "treatment_set.csv"), df)
+CSV.write(string(output_dir_path, "metadata.csv"), treatment_set.metadata)

### 2020-08-29 figs/figure8_sigma_gradient_single_panel/figure8_vis.R

@@ -1,0 +1,93 @@
+setwd("~/phase_transitions_in_metapopulation_synchrony/figs/figure8_sigma_gradient_single_panel/output")
+library(tidyverse)
+library(ggthemr)
+library(latex2exp)
+library(extrafont)
+loadfonts()
+ggthemr('fresh', spacing=3)
+cmuserif = pdfFonts()$`CMU Serif`$family
+thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
+            panel.spacing=unit(3, "lines"),
+            text=element_text(family=cmuserif,size=20), axis.title.y = element_text(angle = 0, vjust=0.5),
+            axis.title = element_text(margin = margin(t = 20, r = 20, b = 0, l = 0)),
+            axis.text = element_text(size=14),
+            panel.grid = element_line(size=0.5, linetype="dotted"),
+            # The new stuff
+            strip.text = element_text(size = 22),
+            strip.text.y = element_text(size = 22, angle = 0, vjust=0.5))
+data = read.csv('metadata.csv') %>% full_join(read.csv('treatment_set.csv'), by="treatment")
+
+
+data %>% 
+  group_by(treatment) %>%
+  filter(lambda==1.0, sigma==0.5) %>%
+  ggplot(aes(migration_rate, summary_stat, group=factor(replicate) )) + 
+  geom_line(size=1,alpha=0.3) +
+  theme_minimal() +
+  scale_x_continuous(limits = c(0.0,1), expand = c(0, 0), breaks = c(0.0, 0.25 ,0.5, 0.75, 1.0)) +
+  labs(title='', y=TeX("$PCC$"), x=TeX("$m$"), color=TeX("$\\alpha$")) + 
+  thm + 
+  facet_grid(vars(sigma), vars(lambda)) +
+  coord_cartesian(xlim = c(0.0,1.0), ylim=c(0,1)) 
+
+convert_to_tex_label = function(string) {
+  return(TeX(string))
+}
+
+
+plt=data %>% 
+  group_by(treatment) %>%
+  mutate(mean_pcc = mean(summary_stat)) %>%
+  mutate(lowest_pcc = quantile(summary_stat, probs=c(0.025))) %>%
+  mutate(lower_pcc = quantile(summary_stat, probs=c(0.25))) %>%
[…]

### 2020-08-29 figs/figure9_variable_scales/figure9.jl

@@ -1,0 +1,49 @@
+
+# Project Folder
+proj_folder = "/home/michael/phase_transitions_in_metapopulation_synchrony/"
+cd(proj_folder)
+
+include(string(proj_folder, "./src/deps.jl"))
+output_dir_path = string(proj_folder , "figs/figure9_variable_scales/output/")
+mkdir(output_dir_path)
+
+
+# _______________________________________________________________________________________
+#
+#
+#   phase transition diagram for two populations
+#
+#
+#
+#
+#
+# _______________________________________________________________________________________
+
+param_dictionary = Dict(
+                        "migration_rate"    => [0.001, 0.01, 0.1, 0.5, 0.9],
+                        "lambda"            => [1.0],
+                        "sigma"             => [0.5],
+                        "carrying_capacity" => [1.0],
+                        "num_populations" => [5,10,20,40],
+                        "alpha"           => [0, 8, 16, 24],  # alpha is meaningless with two pops
+                        "summary_stat" => [random_PCC_within_radius],
+                        # _________________________________________________
+                        #
+                        # simulation parameters
+                        #
+                        # _________________________________________________
+                        "number_of_timesteps" => [300],
+                        "metapopulation_generator" => [get_random_metapopulation],
+                        "fixed_metapopulation" => [false],
+                        "log_abundances" => [false]
+                    )
+
+
+
+treatment_set = create_treatments(param_dictionary, replicates_per_treatment = 50)
+
+df = run_treatments(treatment_set)
+
+
+CSV.write(string(output_dir_path, "treatment_set.csv"), df)
+CSV.write(string(output_dir_path, "metadata.csv"), treatment_set.metadata)

### 2020-08-29 figs/figure9_variable_scales/figure9_vis.R

@@ -1,0 +1,61 @@
+setwd("~/phase_transitions_in_metapopulation_synchrony/figs/figure9_variable_scales/output")
+library(tidyverse)
+library(ggthemr)
+library(latex2exp)
+library(extrafont)
+loadfonts()
+ggthemr('fresh', spacing=3)
+cmuserif = pdfFonts()$`CMU Serif`$family
+thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
+            panel.spacing=unit(3, "lines"),
+            text=element_text(family=cmuserif,size=20), axis.title.y = element_text(angle = 0, vjust=0.5),
+            axis.title = element_text(margin = margin(t = 20, r = 20, b = 0, l = 0)),
+            axis.text = element_text(size=14),
+            panel.grid = element_line(size=0.5, linetype="dotted"),
+            # The new stuff
+            strip.text = element_text(size = 22),
+            strip.text.y = element_text(size = 22, angle = 0, vjust=0.5))
+data = read.csv('metadata.csv') %>% full_join(read.csv('treatment_set.csv'), by="treatment")
+
+convert_to_tex_label = function(string) {
+  return(TeX(string))
+}
+
+plt1= data %>% 
+  filter(migration_rate == 0.5 | migration_rate==0.01 | migration_rate==0.1, num_populations==20) %>%
+  group_by(radius,treatment) %>%
+  mutate(mean_pcc=mean(pcc)) %>%
+  mutate(m_facet=paste("$m =", migration_rate, "$")) %>%
+  mutate(np_facet=paste("$N_p =", num_populations, "$")) %>%
+  ggplot(aes(radius,mean_pcc, color=factor(alpha), group=alpha)) + 
+  geom_point(size=3,aes(shape=factor(alpha))) +
+  geom_line() +
+  scale_shape_manual(values = c(0, 1,5, 6)) + 
+  labs(x=TeX("$radius$"), y=TeX("$PCC$"), color=TeX("$\\alpha$"),shape=TeX("$\\alpha$"))  + 
+  facet_wrap(. ~(m_facet), ncol=3, labeller = as_labeller(convert_to_tex_label, label_parsed)) + 
+  thm 
+
+output_path = "~/phase_transitions_in_metapopulation_synchrony/writing/figs/figure9.png"
[…]

### 2020-08-29 output/vis.R

@@ -1,33 +1,0 @@
-library(tidyverse)
-library(ggthemr)
-library(latex2exp)
-ggthemr('fresh')
-thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
-            text=element_text(family="Iosevka Semibold",size=16), axis.title.y = element_text(angle = 0, hjust=1))
-data = read.csv('metadata.csv') %>% full_join(read.csv('treatment_set.csv'), by="treatment")
-
-data %>% 
-  group_by(treatment) %>%
-  mutate(mean_pcc = mean(summary_stat)) %>%
-  ggplot(aes(migration_rate_distribution, mean_pcc, color=factor(alpha) )) + 
-  geom_point(size=2,alpha=0.5) +
-  theme_minimal() +
-  scale_x_continuous(limits = c(0.01,1), expand = c(0, 0), breaks = c(0.01, 0.25 ,0.5, 0.75, 1.0)) +
-  labs(title='', y=TeX("$PCC$"), x=TeX("$m$"), fill=TeX("$\alpha$")) + 
-  thm + 
-  coord_cartesian(xlim = c(0.01,1.0), ylim=c(0,1)) 
-
-
-data %>% 
-  group_by(treatment) %>%
-  mutate(mean_pcc = mean(summary_stat)) %>%
-  mutate(lower_pcc = quantile(summary_stat, probs=c(0.025))) %>%
-  mutate(upper_pcc = quantile(summary_stat, probs=c(0.975))) %>%
-  ggplot(aes(migration_rate_distribution, summary_stat, group=factor(alpha))) + 
-  geom_ribbon(aes(ymin=lower_pcc, ymax=upper_pcc, fill=factor(alpha)), size=1,alpha=0.4) +
-  theme_minimal() +
-  scale_x_continuous(limits = c(0.01,1), expand = c(0, 0), breaks = c(0.01, 0.25 ,0.5, 0.75, 1.0)) +
-  labs(title='', y=TeX("$PCC$"), x=TeX("$m$"), fill=TeX("$\alpha$")) + 
-  thm + 
-  coord_cartesian(xlim = c(0.01,1.0)) 
-

### 2020-08-29 src/Logging.jl

@@ -1,0 +1,23 @@
+
+function log_abundances(treatment_num, replicate, treatment_instance::TreatmentInstance; filename="abundances.csv")
+    abundance_matrix::Array{Float64,2} = treatment_instance.abundance_matrix
+    df = DataFrame(treatment=[], replicate=[], population=[], timestep=[], abundance=[] )
+    n_pops = length(abundance_matrix[1,:])
+    n_timesteps = length(abundance_matrix[:,1])
+
+    for p = 1:n_pops
+        for t = 1:n_timesteps
+            push!(df.replicate, replicate)
+            push!(df.treatment, treatment_num)
+            push!(df.abundance, abundance_matrix[t,p])
+            push!(df.timestep, t)
+            push!(df.population, p)
+        end
+    end
+
+    if (!isfile(filename))
+        CSV.write(filename, df, append=false)
+    else
+        CSV.write(filename, df, append=true)
+    end
+end

### 2020-08-29 src/StochasticLogisticWDiffusion.jl

@@ -46,13 +46,15 @@ end
 function get_dispersal_matrix(dispersal_potential::DispersalPotential, migration_rate::Vector{Float64})
     n_pops = length(dispersal_potential.matrix[1,:])
 
+    migration_max = 1.0 - (1.0/n_pops)
+
     dispersal_matrix = zeros(n_pops, n_pops)
     for p1 = 1:n_pops
         for p2 = 1:n_pops
             if (p1 == p2)
-                dispersal_matrix[p1,p2] = 1.0 - migration_rate[p1]
+                dispersal_matrix[p1,p2] = 1.0 -  migration_rate[p1]
             else
-                dispersal_matrix[p1,p2] = migration_rate[p1] * dispersal_potential.matrix[p1,p2]
+                dispersal_matrix[p1,p2] =  migration_rate[p1]* dispersal_potential.matrix[p1,p2]
             end
         end
     end

### 2020-08-29 src/SummaryStats.jl

@@ -1,11 +1,42 @@
 
-function PCC(abundance_matrix::Array{Float64, 2})
+function PCC(treatment_instance::TreatmentInstance, treatment_num::Int64, replicate_num::Int64)
+        abundance_matrix::Array{Float64,2} =  treatment_instance.abundance_matrix
         n_pops = length(abundance_matrix[1,:])
         mean_cc::Float64 = 0.0
         s::Float64 =0.0
         ct::Int64 = 0
         for p1 = 1:n_pops
             for p2 = (p1+1):n_pops
+                if (p1 != p2)
+                    v1 = abundance_matrix[:,p1]
+                    v2 = abundance_matrix[:,p2]
+                    cc = crosscor(v1,v2,[0])
+                    s += cc[1]
+                    ct += 1
+                end
+            end
+        end
+
+        mean_cc = s/ct
+        df = DataFrame(treatment=[treatment_num], replicate=[replicate_num], summary_stat=[mean_cc])
+
+        return df
+end
+
+function PCC(abundance_matrix::Array{Float64}, populations::Vector{Int64})
+    n_pops = length(populations)
+
+    if (n_pops < 2)
+        return 1.0
+    end
+
+
+    mean_cc::Float64 = 0.0
+    s::Float64 =0.0
+    ct::Int64 = 0
+    for p1 = 1:n_pops
+        for p2 = (p1+1):n_pops
+            if (p1 != p2)
                 v1 = abundance_matrix[:,p1]
                 v2 = abundance_matrix[:,p2]
                 cc = crosscor(v1,v2,[0])
@@ -13,8 +44,44 @@ function PCC(abundance_matrix::Array{Float64, 2})
                 ct += 1
             end
         end
-
-        mean_cc = s/ct
-        return mean_cc
+    end
+    mean_cc = s/ct
+    return mean_cc
 end
 
+function get_populations_within_circle(mp, center, radius)
+    num_populations = mp.num_populations
+    pops_within_circle::Vector{Int64} = []
+
+    for p in 1:num_populations
+        dist = evaluate(Euclidean(), mp.populations[p].coordinate, center)
+        if (dist < radius)
+            push!(pops_within_circle, p)
+        end
+    end
+    return pops_within_circle
+end
+
+function random_PCC_within_radius(treatment_instance::TreatmentInstance, treatment_num::Int64, replicate_num::Int64; radii = collect(0.0:0.05:(0.5)), num_samples_each_scale=20)
[…]

### 2020-08-29 src/TreatmentFactory.jl

@@ -74,7 +74,7 @@ function create_treatments(param_dict::Dict; replicates_per_treatment::Int64 = 5
 
         dimensionality::Int64 = num_pops
 
-		alpha_dist = Parameter(Normal(alpha, 0.01), 1)
+		alpha_dist = Parameter(Normal(alpha, 0.0), 1)
 		m_dist = Parameter(migration_rate_distribution, dimensionality)
 		lambda_dist = Parameter(lambda_distribution, dimensionality)
 		sigma_dist = Parameter(sigma_distribution, dimensionality)
@@ -85,7 +85,7 @@ function create_treatments(param_dict::Dict; replicates_per_treatment::Int64 = 5
 		dynamics_model = StochasticLogisticWDiffusion()
 		sim_params = SimulationParameters(metadata.number_of_timesteps[t], 0.1, 10, false)
 
-		tr = Treatment(metadata.metapopulation_generator[t], dynamics_model, sim_params, param_bundle, PCC, [])
+		tr = Treatment(metadata.metapopulation_generator[t], dynamics_model, sim_params, param_bundle, metadata.summary_stat[t], metadata.log_abundances[t], [])
 		push!(treatments, tr)
 	end
 
@@ -98,29 +98,35 @@ end
 #
 # =================================================
 
-function run_treatments(treatment_set::TreatmentSet)
+function run_treatments(treatment_set::TreatmentSet; abundances_path = "./abundances.csv")
 	n_treatments::Int64 = length(treatment_set.treatments)
 	n_replicates::Int64 = treatment_set.replicates_per_treatment
 
 	treatments::Vector{Treatment} = treatment_set.treatments
 	@show treatments
-	df = DataFrame(treatment=[], replicate=[], summary_stat=[])
+	df = DataFrame()
 
 	@showprogress for t in (1:n_treatments)
 		summary_stat::Function = treatments[t].summary_stat
 		for r = 1:n_replicates
 			param_values = draw_from_parameter_bundle(treatments[t].theta)
+
 			mp = treatments[t].metapopulation_generator(num_populations = param_values.num_populations, alpha=param_values.alpha)
+
 			abundance_matrix = zeros(Int64(ceil(treatments[t].simulation_parameters.number_of_timesteps/treatments[t].simulation_parameters.log_frequency)), mp.num_populations)
 			initial_condition = [rand(Uniform(0, param_values.carrying_capacity[p])) for p = 1:mp.num_populations]
 
[…]

### 2020-08-29 src/types.jl

@@ -139,6 +139,7 @@ mutable struct Treatment
     simulation_parameters::SimulationParameters
     theta::ParameterBundle
     summary_stat::Function
+    log_abundances::Bool
     instances::Vector{TreatmentInstance}
 end
 

### 2020-09-07 writing/manuscript.tex
ok

@@
 \floatplacement{figure}{h}
+
+\setcounter{secnumdepth}{3}
+\addtocontents{toc}{\setcounter{tocdepth}{2}}
 
@@
 
-\title{Phase Transitions in Landscape Connectivity}
+\title{Critical Transitions in Metapopulation Connectivity}
 \author[1,2]{M.D. Catchen}
@@
 \clearpage
-
+\tableofcontents
+\clearpage
 \section{Introduction}
@@
 presented here, to aid in decision making regarding corridor placement.
-
-\section{What are phase transitions?}
 
@@
 \begin{equation} \label{diffusion_model}
-  d \boldsymbol{N} = \boldsymbol{I} (D^T \boldsymbol{N})\bigg(\boldsymbol{I}\boldsymbol{K}^{-1}(\boldsymbol{K}-D^T \boldsymbol{N})\bigg) d\boldsymbol{t} + \boldsymbol{I}\boldsymbol{\sigma} \boldsymbol{N} \ d\boldsymbol{W}
+    dN_i = \sum_j \bigg[ D_{ji} \lambda_i N_i \bigg(1-\frac{N_i}{K_i}\bigg)  \bigg] dt + \sigma N_i dW_i
 \end{equation}
@@
 \hypertarget{measuring-synchrony}{%
-\subsubsection{Measuring Synchrony}\label{measuring-synchrony}}
+\subsection{Measuring Synchrony}\label{measuring-synchrony}}
 
@@
 \hypertarget{simulating-a-phase-transition-across-a-migration-gradient}{%
-\subsubsection{Simulating a Phase Transition across a Migration
+\subsection{Simulating a Phase Transition across a Migration
 Gradient}\label{simulating-a-phase-transition-across-a-migration-gradient}}
@@
 
+\hypertarget{phase-transitions-across-a-migration-gradient}{%
+\section{Synchrony Across a Migration Gradient}\label{phase-transitions-across-a-migration-gradient}}
+
+Probability distribution of synchrony in a set of populations $L$ given $\alpha$ and $m$.
+
+There is symmetry in these distributions around $m = 1 - \frac{1}{|L|}$
+
 \pagebreak
-\begin{figure}
+\begin{figure}[h]
     \includegraphics[width=15cm]{figs/figure3}
@@
 
-\begin{figure}
+\begin{figure}[h]
     \includegraphics[width=15cm]{figs/figure4}
@@
 
-\begin{figure}
+\begin{figure}[h]
     \includegraphics[width=15cm]{figs/figure5}
@@
 \end{figure}
+
+\hypertarget{critical-transitions-across-space}{%
+\section{Critical Transitions in Synchrony Across Space}\label{critical-transitions-across-space}}
+
+\hypertarget{simulating-a-phase-transition-across-a-migration-gradient}{%
+\subsection{Landscape Connectivity}\label{simulating-a-phase-transition-across-a-migration-gradient}}
+
+\begin{figure}
+    \includegraphics[width=15cm]{figs/figure9}
+    \caption{}
+    \label{}
+\end{figure}
+
+\hypertarget{simulating-a-phase-transition-across-a-migration-gradient}{%
+\subsection{Landscape Subdivision}\label{simulating-a-phase-transition-across-a-migration-gradient}}
+
+\begin{figure}
+    \includegraphics[width=15cm]{figs/figure10}
+    \caption{}
+    \label{}
+\end{figure}
+
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
 

### 2020-09-07 figs/figure1_metapopulations_as_a_network/figure1.R

@@ -7,11 +7,10 @@ library(extrafont)
 library(latex2exp)
 library(Cairo)
 library(extrafont)
-loadfonts()
 setwd("~/phase_transitions_in_metapopulation_synchrony/figs")
 loadfonts()
-cmuserif = p$`CMU Serif`$family
-## ============================================ 
+cmuserif = pdfFonts()$`CMU Serif`$family
+## ============================================
 ## DISPERSAL KERNEL PLOTTING / FIG 1
 ## ============================================
 exp_kern = function(alpha, d_ij){
@@ -26,45 +25,45 @@ gauss_kern = function(alpha, d_ij){
 get_pop_points = function(x_vals, y_vals, alpha, mig_prop, kernel=exp_kern, kernel_name="Exp.", xaxis=F, yaxis=F){
     # make df with all pairwise and group by indiv pairwise val
     df = data.frame(matrix(ncol=5))
-    
+
     line_ct = 1
-    grp_ct = 0 
-    
+    grp_ct = 0
+
     # make a dis potential first
-    
+
     npops = length(x_vals)
-    
+
     dispersal_potential = matrix(nrow=npops, ncol=npops)
     for (pt_ct in seq(1, n_pops)){
         s = 0
         for (pt_ct2 in seq(1, n_pops)){
-            x1 = x_vals[pt_ct] 
-            y1 = y_vals[pt_ct] 
-            x2 = x_vals[pt_ct2] 
-            y2 = y_vals[pt_ct2] 
-            
+            x1 = x_vals[pt_ct]
+            y1 = y_vals[pt_ct]
+            x2 = x_vals[pt_ct2]
+            y2 = y_vals[pt_ct2]
+
             dist = sqrt((x2-x1)^2 + (y2-y1)^2)
-            
+
             dispersal_potential[pt_ct, pt_ct2] = kernel(alpha, dist)
-            
-            s = s + dispersal_potential[pt_ct, pt_ct2] 
+
+            s = s + dispersal_potential[pt_ct, pt_ct2]
         }
         for (pt_ct2 in seq(1, n_pops)){
             dispersal_potential[pt_ct, pt_ct2]  = dispersal_potential[pt_ct, pt_ct2]  / s
         }
     }
-    
+
     for (pt_ct in seq(1, n_pops)){
[…]

### 2020-09-07 figs/figure2_synchrony_example/figure2.jl

@@ -24,7 +24,7 @@ param_dictionary = Dict(
                         "lambda"            => [1.0],
                         "sigma"             => [0.5],
                         "carrying_capacity" => [1.0],
-                        "num_populations" => [2, 8],
+                        "num_populations" => [5],
                         "alpha"           => [0],  # alpha is meaningless with two pops
 
                         # _________________________________________________

### 2020-09-07 figs/figure3_2population_phase_transition/figure3.jl

@@ -7,6 +7,7 @@ include(string(proj_folder, "./src/deps.jl"))
 output_dir_path = string(proj_folder , "figs/figure3_2population_phase_transition/output/")
 
 
+mkdir(output_dir_path)
 
 # _______________________________________________________________________________________
 #
@@ -35,6 +36,7 @@ param_dictionary = Dict(
                         "number_of_timesteps" => [300],
                         "metapopulation_generator" => [get_random_metapopulation],
                         "fixed_metapopulation" => [false],
+                        "summary_stat" => [PCC],
                         "log_abundances" => [false]
                     )
 
@@ -44,7 +46,6 @@ treatment_set = create_treatments(param_dictionary, replicates_per_treatment = 5
 
 df = run_treatments(treatment_set)
 
-mkdir(output_dir_path)
 
 CSV.write(string(output_dir_path, "treatment_set.csv"), df)
 CSV.write(string(output_dir_path, "metadata.csv"), treatment_set.metadata)

### 2020-09-07 figs/figure3_2population_phase_transition/vis.R

@@ -6,7 +6,7 @@ library(extrafont)
 loadfonts()
 ggthemr('fresh', spacing=2)
 cmuserif = pdfFonts()$`CMU Serif`$family
-thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
+thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75),
             panel.spacing=unit(3, "lines"),
             text=element_text(family=cmuserif,size=20), axis.title.y = element_text(angle = 0, vjust=0.5),
             axis.title = element_text(margin = margin(t = 20, r = 20, b = 0, l = 0)),
@@ -17,18 +17,16 @@ thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75
             strip.text.y = element_text(size = 22, angle = 0, vjust=0.5))
 data = read.csv('metadata.csv') %>% full_join(read.csv('treatment_set.csv'), by="treatment")
 
-
-data %>% 
-  group_by(treatment) %>%
-  filter(lambda==1.0, sigma==0.5) %>%
-  ggplot(aes(migration_rate, summary_stat, group=factor(replicate) )) + 
-  geom_line(size=1,alpha=0.3) +
+data %>%
+  group_by(replicate) %>%
+  ggplot(aes(migration_rate, mean_cc, group=(replicate) )) +
+  geom_point(alpha=0.2, size=1.5) +
   theme_minimal() +
   scale_x_continuous(limits = c(0.0,1), expand = c(0, 0), breaks = c(0.0, 0.25 ,0.5, 0.75, 1.0)) +
-  labs(title='', y=TeX("$PCC$"), x=TeX("$m$"), color=TeX("$\\alpha$")) + 
-  thm + 
+  labs(title='', y=TeX("$PCC$"), x=TeX("$m$"), color=TeX("$\\alpha$")) +
+  thm +
   facet_grid(vars(sigma), vars(lambda)) +
-  coord_cartesian(xlim = c(0.0,1.0), ylim=c(0,1)) 
+  coord_cartesian(xlim = c(0.0,1.0), ylim=c(0,1))
 
 
 convert_to_tex_label = function(string) {
@@ -37,29 +35,29 @@ convert_to_tex_label = function(string) {
 
 
 
-plt = data %>% 
-  group_by(treatment) %>%
[…]

### 2020-09-07 figs/figure4_20population_phase_transition_varying_alpha/figure4_vis.R

@@ -42,7 +42,8 @@ plt = data %>%
   mutate(lower_pcc = quantile(summary_stat, probs=c(0.25))) %>%
   mutate(higher_pcc = quantile(summary_stat, probs=c(0.75))) %>%
   mutate(highest_pcc = quantile(summary_stat, probs=c(0.975))) %>%
-  mutate(alpha_facet = paste("$\\alpha =$", alpha)) %>%
+  mutate(alpha_facet = paste("$\\alpha = ", alpha, "$", sep="")) %>%
+  mutate(alpha_facet = factor(alpha_facet, levels=c("$\\alpha = 0$", "$\\alpha = 10$", "$\\alpha = 20$","$\\alpha = 30$"))) %>%
   arrange(alpha) %>%
   ggplot(aes(migration_rate, summary_stat, group=factor(alpha), fill=factor(alpha))) + 
   geom_ribbon(aes(ymin=lower_pcc, ymax=higher_pcc), size=1,alpha=0.4) +
@@ -53,9 +54,17 @@ plt = data %>%
   geom_hline(aes(yintercept=1), linetype='dashed',color='black') +
   scale_x_continuous(limits = c(0.0,1), expand = c(0, 0), breaks = c(0, 0.25,0.5, 0.75, 1)) +
   labs(title='', y=TeX("$PCC$"), x=TeX("$m$"), color=TeX("$\\alpha$"), fill=TeX("$\\alpha$")) + 
-  facet_wrap(. ~ alpha_facet, labeller=as_labeller(convert_to_tex_label, label_parsed), ncol=2) +
+  facet_wrap(. ~ alpha_facet, ncol=2, labeller= as_labeller(convert_to_tex_label, label_parsed)) +
   thm + 
   coord_cartesian(xlim = c(0.0,1.0), ylim=c(0.0,1.0)) 
+
+
+d = seq(0, 1, by=0.01)
+alpha = rep(c(0, 10, 20, 30))
+
+
+exp_kern = exp(-1*d*alpha)
+
 
 output_path = "~/phase_transitions_in_metapopulation_synchrony/writing/figs/figure4.png"
 ggsave(output_path, plot=plt, dpi=320, width = 12, height = 8, units = "in", device=png())

### 2020-09-07 figs/figure6_even_mixing/figure6_vis.R

@@ -17,25 +17,27 @@ data = read.csv('metadata.csv') %>% full_join(read.csv('treatment_set.csv'), by=
 
 
 
-data %>% 
+plt = data %>% 
+  group_by(num_populations, alpha) %>% 
+  mutate(max_pcc =(max(summary_stat))) %>%
+  ungroup() %>%
   group_by(num_populations, replicate, alpha) %>% 
-  filter(alpha==0) %>%
-  filter(abs(max(summary_stat)-summary_stat) < 0.01) %>%
   mutate(low = quantile(migration_rate,0.25), high=quantile(migration_rate, 0.75)) %>%
   mutate(lowest = quantile(migration_rate,0.025), highest=quantile(migration_rate, 0.975)) %>%
   ggplot(aes(num_populations,migration_rate, fill=factor(alpha)))  + 
   geom_ribbon(aes(ymin=lowest, ymax=highest), alpha=0.3) +
   geom_ribbon(aes(ymin=low, ymax=high), alpha=0.4) +
+  geom_point(aes(y=max_pcc, color=factor(alpha))) + 
   stat_function(fun = function(x) 1.0 - 1.0/x) + 
   labs(x=TeX("$N_p$"), y=TeX("$m$"))  + 
-  scale_x_continuous(breaks=seq(2,26,by=2), limits=c(2,25)) +
-  scale_y_continuous(breaks=c(0.5, 0.6,0.7,0.8,0.9,1.0), limits=c(0.5,1)) +
-  thm +
-  coord_cartesian(xlim = c(4,25), ylim=c(0.4,1)) 
-  
+  facet_wrap(. ~ alpha) + 
+  scale_x_continuous(breaks=seq(2,26,by=2), limits=c(2,26)) +
+  scale_y_continuous(breaks=c(0.5,0.6,0.7,0.8,0.9,1.0), limits=c(0.5,1)) +
+  thm
 
-output_path = "~/phase_transitions_in_metapopulation_synchrony/writing/figs/figure5.pdf"
-ggsave(output_path, plot=plt, dpi=320, width = 18, height = 8, units = "in", device=cairo_pdf)
+
+output_path = "~/phase_transitions_in_metapopulation_synchrony/writing/figs/figure6.png"
+ggsave(output_path, plot=plt, dpi=320, width = 18, height = 8, units = "in", device=png())
 
 #+ 
  # coord_cartesian(xlim = c(0.0,1.0)) 

### 2020-09-07 figs/figure9_variable_scales/figure9.jl

@@ -20,11 +20,11 @@ mkdir(output_dir_path)
 # _______________________________________________________________________________________
 
 param_dictionary = Dict(
-                        "migration_rate"    => [0.001, 0.01, 0.1, 0.5, 0.9],
+                        "migration_rate"    => [0.001, 0.01, 0.1, 0.25, 0.5, 0.95],
                         "lambda"            => [1.0],
                         "sigma"             => [0.5],
                         "carrying_capacity" => [1.0],
-                        "num_populations" => [5,10,20,40],
+                        "num_populations" => [5,10,20,40,80],
                         "alpha"           => [0, 8, 16, 24],  # alpha is meaningless with two pops
                         "summary_stat" => [random_PCC_within_radius],
                         # _________________________________________________

### 2020-09-07 figs/figure9_variable_scales/figure9_vis.R

@@ -22,7 +22,7 @@ convert_to_tex_label = function(string) {
 }
 
 plt1= data %>% 
-  filter(migration_rate == 0.5 | migration_rate==0.01 | migration_rate==0.1, num_populations==20) %>%
+  filter(num_populations==20) %>%
   group_by(radius,treatment) %>%
   mutate(mean_pcc=mean(pcc)) %>%
   mutate(m_facet=paste("$m =", migration_rate, "$")) %>%
@@ -31,6 +31,8 @@ plt1= data %>%
   geom_point(size=3,aes(shape=factor(alpha))) +
   geom_line() +
   scale_shape_manual(values = c(0, 1,5, 6)) + 
+  theme(aspect.ratio=1) +
+  ylim(0,1) +
   labs(x=TeX("$radius$"), y=TeX("$PCC$"), color=TeX("$\\alpha$"),shape=TeX("$\\alpha$"))  + 
   facet_wrap(. ~(m_facet), ncol=3, labeller = as_labeller(convert_to_tex_label, label_parsed)) + 
   thm 
@@ -39,20 +41,22 @@ output_path = "~/phase_transitions_in_metapopulation_synchrony/writing/figs/figu
 ggsave(output_path, plot=plt1, dpi=320, width = 18, height = 10, units = "in", device=png())
 
 plt2 =  data %>% 
+  filter((num_populations) < 50) %>%
   group_by(radius,treatment) %>%
   mutate(mean_pcc=mean(pcc)) %>%
   mutate(m_facet=as.factor(paste("$m =", migration_rate, "$"))) %>%
   arrange(num_populations) %>%
   mutate(np_facet=paste("$N_p = ", num_populations, "$", sep="")) %>%
-  mutate(np_facet=factor(np_facet, levels=c("$N_p = 5$", "$N_p = 10$", "$N_p = 20$","$N_p = 40$")) ) %>%
+  mutate(np_facet=factor(np_facet, levels=c("$N_p = 5$", "$N_p = 10$", "$N_p = 20$","$N_p = 40$", "$N_p=80$"))) %>% 
   ggplot(aes(radius,mean_pcc, color=factor(alpha), group=alpha)) + 
   geom_point(size=1.5,aes(shape=factor(alpha))) +
   geom_line() +
   scale_shape_manual(values = c(0, 1,5, 6)) + 
   theme(aspect.ratio=1) + 
[…]

### 2020-09-07 figs/network_vis.R

@@ -1,4 +1,4 @@
-elibrary(ggplot2)
+library(ggplot2)
 library(ggnet)
 library(gridExtra)
 library(tikzDevice)
@@ -10,12 +10,12 @@ setwd("~/phase_transitions_in_metapopulation_synchrony/figs")
 loadfonts()
 
 data(mtcars)
-mtcars %>% ggplot(aes(x=mpg, y=disp)) + geom_col(aes(col=cyl)) + 
+mtcars %>% ggplot(aes(x=mpg, y=disp)) + geom_col(aes(col=cyl)) +
     theme(text  = element_text(size=10, family=p$`CMU Serif`$family)) + ggtitle("for the boys")
 
 
 cmuserif = p$`CMU Serif`$family
-## ============================================ 
+## ============================================
 ## DISPERSAL KERNEL PLOTTING / FIG 1
 ## ============================================
 exp_kern = function(alpha, d_ij){
@@ -30,45 +30,45 @@ gauss_kern = function(alpha, d_ij){
 get_pop_points = function(x_vals, y_vals, alpha, mig_prop, kernel=exp_kern, kernel_name="Exp.", xaxis=F, yaxis=F){
     # make df with all pairwise and group by indiv pairwise val
     df = data.frame(matrix(ncol=5))
-    
+
     line_ct = 1
-    grp_ct = 0 
-    
+    grp_ct = 0
+
     # make a dis potential first
-    
+
     npops = length(x_vals)
-    
+
     dispersal_potential = matrix(nrow=npops, ncol=npops)
     for (pt_ct in seq(1, n_pops)){
         s = 0
         for (pt_ct2 in seq(1, n_pops)){
-            x1 = x_vals[pt_ct] 
-            y1 = y_vals[pt_ct] 
-            x2 = x_vals[pt_ct2] 
-            y2 = y_vals[pt_ct2] 
-            
+            x1 = x_vals[pt_ct]
+            y1 = y_vals[pt_ct]
+            x2 = x_vals[pt_ct2]
+            y2 = y_vals[pt_ct2]
+
             dist = sqrt((x2-x1)^2 + (y2-y1)^2)
-            
+
             dispersal_potential[pt_ct, pt_ct2] = kernel(alpha, dist)
-            
-            s = s + dispersal_potential[pt_ct, pt_ct2] 
+
+            s = s + dispersal_potential[pt_ct, pt_ct2]
         }
         for (pt_ct2 in seq(1, n_pops)){
[…]

### 2020-09-07 src/SummaryStats.jl

@@ -18,7 +18,7 @@ function PCC(treatment_instance::TreatmentInstance, treatment_num::Int64, replic
         end
 
         mean_cc = s/ct
-        df = DataFrame(treatment=[treatment_num], replicate=[replicate_num], summary_stat=[mean_cc])
+        df = DataFrame(treatment=[treatment_num], replicate=[replicate_num], mean_cc=[mean_cc])
 
         return df
 end

### 2020-09-07 src/TreatmentFactory.jl

@@ -80,12 +80,18 @@ function create_treatments(param_dict::Dict; replicates_per_treatment::Int64 = 5
 		sigma_dist = Parameter(sigma_distribution, dimensionality)
 		carrying_cap_dist = Parameter(carrying_capacity_distribution, dimensionality)
 
-		param_bundle = StochasticLogisticParameterBundle(num_pops, alpha, m_dist, lambda_dist,sigma_dist, carrying_cap_dist)
+		param_bundle = StochasticLogisticParameterBundle(num_pops, alpha, m_dist, lambda_dist, sigma_dist, carrying_cap_dist)
 
 		dynamics_model = StochasticLogisticWDiffusion()
 		sim_params = SimulationParameters(metadata.number_of_timesteps[t], 0.1, 10, false)
 
-		tr = Treatment(metadata.metapopulation_generator[t], dynamics_model, sim_params, param_bundle, metadata.summary_stat[t], metadata.log_abundances[t], [])
+		if (metadata.fixed_metapopulation[t])
+
+		end
+
+		mp = metadata.metapopulation_generator[t](num_populations=num_pops, alpha=alpha)
+
+		tr = Treatment(mp, dynamics_model, sim_params, param_bundle, metadata.summary_stat[t], metadata.log_abundances[t], [])
 		push!(treatments, tr)
 	end
 
@@ -111,8 +117,7 @@ function run_treatments(treatment_set::TreatmentSet; abundances_path = "./abunda
 		for r = 1:n_replicates
 			param_values = draw_from_parameter_bundle(treatments[t].theta)
 
-			mp = treatments[t].metapopulation_generator(num_populations = param_values.num_populations, alpha=param_values.alpha)
-
+			mp = treatments[t].metapopulation_generator
 			abundance_matrix = zeros(Int64(ceil(treatments[t].simulation_parameters.number_of_timesteps/treatments[t].simulation_parameters.log_frequency)), mp.num_populations)
 			initial_condition = [rand(Uniform(0, param_values.carrying_capacity[p])) for p = 1:mp.num_populations]
 

### 2020-09-07 src/Vis.jl

@@ -1,3 +1,18 @@
 function plot_timeseries_of_abundances(abundance_matrix)
     plot(collect(1:length(abundance_matrix[:,1])), abundance_matrix, linewidth=2,      size=(700, 400))
 end
+
+
+#=
+using StatsPlots
+data = innerjoin(df, treatment_set.metadata, on=:treatment)
+
+@df data scatter(:migration_rate,:mean_cc,
+                 group=(:lambda,:sigma),
+                 layout=(3,3),
+                 markeralpha=0.1,
+                 markersize=2,
+                 markerstrokewidth=0.5,
+                 markerstrokealpha=0)
+                 =#
+

### 2020-09-13 figs/figure2_synchrony_example/figure2.jl
updating

@@ -1,6 +1,6 @@
 
 # Project Folder
-proj_folder = "/home/michael/phase_transitions_in_metapopulation_synchrony/"
+proj_folder = "/home/michael/phase-transitions-in-metapopulation-synchrony/"
 cd(proj_folder)
 
 include(string(proj_folder, "./src/deps.jl"))
@@ -26,7 +26,9 @@ param_dictionary = Dict(
                         "carrying_capacity" => [1.0],
                         "num_populations" => [5],
                         "alpha"           => [0],  # alpha is meaningless with two pops
+                        "summary_stat"    => [PCC],
 
+                        
                         # _________________________________________________
                         #
                         # simulation parameters

### 2020-09-13 figs/figure2_synchrony_example/figure2_vis.R

@@ -1,4 +1,4 @@
-setwd("~/phase_transitions_in_metapopulation_synchrony/figs/figure2_synchrony_example/output")
+setwd("~/phase-transitions-in-metapopulation-synchrony/figs/figure2_synchrony_example/output")
 library(tidyverse)
 library(ggthemr)
 library(latex2exp)
@@ -33,6 +33,8 @@ plt = data %>%
     geom_hline(aes(yintercept=1.0) ,linetype="dashed", color='black') + 
     thm
 
+
+plt
 
 
 output_path = "~/phase_transitions_in_metapopulation_synchrony/writing/figs/figure2.png"

### 2020-09-13 figs/figure3_2population_phase_transition/figure3.jl

@@ -1,6 +1,5 @@
 
-# Project Folder
-proj_folder = "/home/michael/phase_transitions_in_metapopulation_synchrony/"
+proj_folder = "/home/michael/phase-transitions-in-metapopulation-synchrony/"
 cd(proj_folder)
 
 include(string(proj_folder, "./src/deps.jl"))
@@ -42,7 +41,7 @@ param_dictionary = Dict(
 
 
 
-treatment_set = create_treatments(param_dictionary, replicates_per_treatment = 50)
+treatment_set = create_treatments(param_dictionary, replicates_per_treatment = 100)
 
 df = run_treatments(treatment_set)
 

### 2020-09-13 figs/figure3_2population_phase_transition/vis.R

@@ -1,11 +1,11 @@
-setwd("~/phase_transitions_in_metapopulation_synchrony/figs/figure3_2population_phase_transition/output")
+setwd("~/phase-transitions-in-metapopulation-synchrony/figs/figure3_2population_phase_transition/output")
 library(tidyverse)
 library(ggthemr)
 library(latex2exp)
 library(extrafont)
 loadfonts()
 ggthemr('fresh', spacing=2)
-cmuserif = pdfFonts()$`CMU Serif`$family
+cmuserif = pdfFonts()$`CM Roman`$family
 thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75),
             panel.spacing=unit(3, "lines"),
             text=element_text(family=cmuserif,size=20), axis.title.y = element_text(angle = 0, vjust=0.5),
@@ -20,8 +20,7 @@ data = read.csv('metadata.csv') %>% full_join(read.csv('treatment_set.csv'), by=
 data %>%
   group_by(replicate) %>%
   ggplot(aes(migration_rate, mean_cc, group=(replicate) )) +
-  geom_point(alpha=0.2, size=1.5) +
-  theme_minimal() +
+  geom_point(alpha=0.1, size=1.5) +
   scale_x_continuous(limits = c(0.0,1), expand = c(0, 0), breaks = c(0.0, 0.25 ,0.5, 0.75, 1.0)) +
   labs(title='', y=TeX("$PCC$"), x=TeX("$m$"), color=TeX("$\\alpha$")) +
   thm +
@@ -54,6 +53,7 @@ plt = data %>%
   labs(title='', y=TeX("$PCC$"), x=TeX("$m$"), color=TeX("$\\sigma$"), fill=TeX("$\\sigma$")) +
   facet_grid(vars(sigma_facet), vars(lambda_facet), labeller =as_labeller(convert_to_tex_label, label_parsed)) + thm
 
+plt
 
 output_path = "~/phase_transitions_in_metapopulation_synchrony/writing/figs/figure3.png"
 ggsave(output_path, plot=plt, dpi=320, width = 12, height = 12, units = "in", device=png())

### 2020-09-13 figs/figure4_20population_phase_transition_varying_alpha/figure4.jl

@@ -1,6 +1,6 @@
 
 # Project Folder
-proj_folder = "/home/michael/phase_transitions_in_metapopulation_synchrony/"
+proj_folder = "/home/michael/phase-transitions-in-metapopulation-synchrony/"
 cd(proj_folder)
 
 include(string(proj_folder, "src/deps.jl"))
@@ -25,16 +25,17 @@ param_dictionary = Dict(
                         "carrying_capacity" => [1.0],
                         "num_populations" => [20],
                         "alpha"           => [0, 10, 20, 30],
+                        "summary_stat"    => [PCC],
                         # _________________________________________________
                         #
                         # simulation parameters
                         #
                         # _________________________________________________
                         "number_of_timesteps" => [300],
-                        "metapopulation_generator" => [get_random_metapopulation]
+                        "metapopulation_generator" => [get_random_metapopulation],
+                        "fixed_metapopulation" => [false],
+                        "log_abundances"       => [false]
                     )
-
-
 
 treatment_set = create_treatments(param_dictionary, replicates_per_treatment = 50)
 

### 2020-09-13 figs/figure4_20population_phase_transition_varying_alpha/figure4_vis.R

@@ -1,11 +1,11 @@
-setwd("~/phase_transitions_in_metapopulation_synchrony/figs/figure4_20population_phase_transition_varying_alpha/output")
+setwd("~/phase-transitions-in-metapopulation-synchrony/figs/figure4_20population_phase_transition_varying_alpha/output")
 library(tidyverse)
 library(ggthemr)
 library(latex2exp)
 library(extrafont)
 loadfonts()
 ggthemr('fresh', spacing=3)
-cmuserif = pdfFonts()$`CMU Serif`$family
+cmuserif = pdfFonts()$`CM Roman`$family
 thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
             panel.spacing=unit(3, "lines"),
             text=element_text(family=cmuserif,size=20), axis.title.y = element_text(angle = 0, vjust=0.5),
@@ -37,18 +37,18 @@ convert_to_tex_label = function(string) {
 
 plt = data %>% 
   group_by(treatment) %>%
-  mutate(mean_pcc = mean(summary_stat)) %>%
-  mutate(lowest_pcc = quantile(summary_stat, probs=c(0.025))) %>%
-  mutate(lower_pcc = quantile(summary_stat, probs=c(0.25))) %>%
-  mutate(higher_pcc = quantile(summary_stat, probs=c(0.75))) %>%
-  mutate(highest_pcc = quantile(summary_stat, probs=c(0.975))) %>%
+  mutate(mean_pcc = mean(mean_cc)) %>%
+  mutate(lowest_pcc = quantile(mean_cc, probs=c(0.025))) %>%
+  mutate(lower_pcc = quantile(mean_cc, probs=c(0.25))) %>%
+  mutate(higher_pcc = quantile(mean_cc, probs=c(0.75))) %>%
+  mutate(highest_pcc = quantile(mean_cc, probs=c(0.975))) %>%
   mutate(alpha_facet = paste("$\\alpha = ", alpha, "$", sep="")) %>%
   mutate(alpha_facet = factor(alpha_facet, levels=c("$\\alpha = 0$", "$\\alpha = 10$", "$\\alpha = 20$","$\\alpha = 30$"))) %>%
   arrange(alpha) %>%
-  ggplot(aes(migration_rate, summary_stat, group=factor(alpha), fill=factor(alpha))) + 
[…]

### 2020-09-13 figs/figure5_20population_phase_transition_varying_sigma_and_lambda/figure5.jl

@@ -1,6 +1,6 @@
 
 # Project Folder
-proj_folder = "/home/michael/phase_transitions_in_metapopulation_synchrony/"
+proj_folder = "/home/michael/phase-transitions-in-metapopulation-synchrony/"
 cd(proj_folder)
 
 include(string(proj_folder, "./src/deps.jl"))
@@ -25,13 +25,16 @@ param_dictionary = Dict(
                         "carrying_capacity" => [1.0],
                         "num_populations" => [20],
                         "alpha"           => [0],
+                        "summary_stat"     => [PCC],
                         # _________________________________________________
                         #
                         # simulation parameters
                         #
                         # _________________________________________________
                         "number_of_timesteps" => [300],
-                        "metapopulation_generator" => [get_random_metapopulation]
+                        "metapopulation_generator" => [get_random_metapopulation],
+                        "fixed_metapopulation" => [false],
+                        "log_abundances" => [false]
                     )
 
 

### 2020-09-13 figs/figure5_20population_phase_transition_varying_sigma_and_lambda/figure5_vis.R

@@ -1,4 +1,4 @@
-setwd("~/phase_transitions_in_metapopulation_synchrony/figs/figure5_20population_phase_transition_varying_sigma_and_lambda/output")
+setwd("~/phase-transitions-in-metapopulation-synchrony/figs/figure5_20population_phase_transition_varying_sigma_and_lambda/output")
 library(tidyverse)
 library(ggthemr)
 library(latex2exp)
@@ -37,14 +37,14 @@ convert_to_tex_label = function(string) {
 
 plt=data %>% 
   group_by(treatment) %>%
-  mutate(mean_pcc = mean(summary_stat)) %>%
-  mutate(lowest_pcc = quantile(summary_stat, probs=c(0.025))) %>%
-  mutate(lower_pcc = quantile(summary_stat, probs=c(0.25))) %>%
-  mutate(higher_pcc = quantile(summary_stat, probs=c(0.75))) %>%
-  mutate(highest_pcc = quantile(summary_stat, probs=c(0.975))) %>%
+  mutate(mean_pcc = mean(mean_cc)) %>%
+  mutate(lowest_pcc = quantile(mean_cc, probs=c(0.025))) %>%
+  mutate(lower_pcc = quantile(mean_cc, probs=c(0.25))) %>%
+  mutate(higher_pcc = quantile(mean_cc, probs=c(0.75))) %>%
+  mutate(highest_pcc = quantile(mean_cc, probs=c(0.975))) %>%
   mutate(sigma_facet = paste("$\\sigma =", as.numeric(sigma), "$")) %>%
   mutate(lambda_facet = paste("$\\lambda =", as.numeric(lambda), "$")) %>%
-  ggplot(aes(migration_rate, summary_stat, group=factor(sigma))) + 
+  ggplot(aes(migration_rate, mean_cc, group=factor(sigma))) + 
   geom_ribbon(aes(ymin=lower_pcc, ymax=higher_pcc, fill=factor(sigma)), size=1,alpha=0.4) +
   geom_line(aes(y=mean_pcc, color=factor(sigma)), size=0.5, linetype='dashed',) +
   geom_ribbon(aes(ymin=lowest_pcc, ymax=highest_pcc, fill=factor(sigma)), size=1,alpha=0.4) +
@@ -55,6 +55,8 @@ plt=data %>%
   facet_grid(vars(sigma_facet), vars(lambda_facet), labeller = as_labeller(convert_to_tex_label, label_parsed)) +
   thm
 
+
+plt
[…]

### 2020-09-13 figs/figure6_even_mixing/figure6.jl

@@ -1,6 +1,6 @@
 
 # Project Folder
-proj_folder = "/home/michael/phase_transitions_in_metapopulation_synchrony/"
+proj_folder = "/home/michael/phase-transitions-in-metapopulation-synchrony/"
 cd(proj_folder)
 
 include(string(proj_folder, "src/deps.jl"))
@@ -25,13 +25,17 @@ param_dictionary = Dict(
                         "carrying_capacity" => [1.0],
                         "num_populations" => collect(2:25),
                         "alpha"           => [0.0, 5.0, 10.0, 20.0],
+                        "summary_stat" => [PCC],
+
                         # _________________________________________________
                         #
                         # simulation parameters
                         #
                         # _________________________________________________
                         "number_of_timesteps" => [300],
-                        "metapopulation_generator" => [get_random_metapopulation]
+                        "metapopulation_generator" => [get_random_metapopulation],
+                        "fixed_metapopulation" => [false],
+                        "log_abundances" => [false],
                     )
 
 

### 2020-09-13 figs/figure6_even_mixing/figure6_vis.R

@@ -1,4 +1,4 @@
-setwd("~/phase_transitions_in_metapopulation_synchrony/figs/figure6_even_mixing/output")
+setwd("~/phase-transitions-in-metapopulation-synchrony/figs/figure6_even_mixing/output")
 library(tidyverse)
 library(ggthemr)
 library(latex2exp)
@@ -19,21 +19,24 @@ data = read.csv('metadata.csv') %>% full_join(read.csv('treatment_set.csv'), by=
 
 plt = data %>% 
   group_by(num_populations, alpha) %>% 
-  mutate(max_pcc =(max(summary_stat))) %>%
+  mutate(max_pcc = (max(mean_cc))) %>%
+  filter(abs(mean_cc - max_pcc) < 0.01) %>%
   ungroup() %>%
-  group_by(num_populations, replicate, alpha) %>% 
+  group_by(num_populations, alpha) %>% 
   mutate(low = quantile(migration_rate,0.25), high=quantile(migration_rate, 0.75)) %>%
   mutate(lowest = quantile(migration_rate,0.025), highest=quantile(migration_rate, 0.975)) %>%
   ggplot(aes(num_populations,migration_rate, fill=factor(alpha)))  + 
   geom_ribbon(aes(ymin=lowest, ymax=highest), alpha=0.3) +
   geom_ribbon(aes(ymin=low, ymax=high), alpha=0.4) +
-  geom_point(aes(y=max_pcc, color=factor(alpha))) + 
+ # geom_point(aes(y=migration_rate, color=factor(alpha))) + 
   stat_function(fun = function(x) 1.0 - 1.0/x) + 
   labs(x=TeX("$N_p$"), y=TeX("$m$"))  + 
   facet_wrap(. ~ alpha) + 
   scale_x_continuous(breaks=seq(2,26,by=2), limits=c(2,26)) +
   scale_y_continuous(breaks=c(0.5,0.6,0.7,0.8,0.9,1.0), limits=c(0.5,1)) +
-  thm
+  thm +
+  coord_cartesian(ylim = c(0.5,1.0)) 
+  
 
 
 output_path = "~/phase_transitions_in_metapopulation_synchrony/writing/figs/figure6.png"

### 2020-09-13 figs/figure8_sigma_gradient_single_panel/figure8.jl

@@ -1,6 +1,6 @@
 
 # Project Folder
-proj_folder = "/home/michael/phase_transitions_in_metapopulation_synchrony/"
+proj_folder = "/home/michael/phase-transitions-in-metapopulation-synchrony/"
 cd(proj_folder)
 
 include(string(proj_folder, "./src/deps.jl"))
@@ -25,6 +25,7 @@ param_dictionary = Dict(
                         "carrying_capacity" => [2.0],
                         "num_populations" => [20],
                         "alpha"           => [0],
+                        "summary_stat" => [PCC],
                         # _________________________________________________
                         #
                         # simulation parameters

### 2020-09-13 figs/figure8_sigma_gradient_single_panel/figure8_vis.R

@@ -1,4 +1,4 @@
-setwd("~/phase_transitions_in_metapopulation_synchrony/figs/figure8_sigma_gradient_single_panel/output")
+setwd("~/phase-transitions-in-metapopulation-synchrony/figs/figure8_sigma_gradient_single_panel/output")
 library(tidyverse)
 library(ggthemr)
 library(latex2exp)
@@ -37,14 +37,14 @@ convert_to_tex_label = function(string) {
 
 plt=data %>% 
   group_by(treatment) %>%
-  mutate(mean_pcc = mean(summary_stat)) %>%
-  mutate(lowest_pcc = quantile(summary_stat, probs=c(0.025))) %>%
-  mutate(lower_pcc = quantile(summary_stat, probs=c(0.25))) %>%
-  mutate(higher_pcc = quantile(summary_stat, probs=c(0.75))) %>%
-  mutate(highest_pcc = quantile(summary_stat, probs=c(0.975))) %>%
+  mutate(mean_pcc = mean(mean_cc)) %>%
+  mutate(lowest_pcc = quantile(mean_cc, probs=c(0.025))) %>%
+  mutate(lower_pcc = quantile(mean_cc, probs=c(0.25))) %>%
+  mutate(higher_pcc = quantile(mean_cc, probs=c(0.75))) %>%
+  mutate(highest_pcc = quantile(mean_cc, probs=c(0.975))) %>%
   mutate(sigma_facet = paste("$\\sigma =", as.numeric(sigma), "$")) %>%
   mutate(lambda_facet = paste("$\\lambda =", as.numeric(lambda), "$")) %>%
-  ggplot(aes(migration_rate, summary_stat, group=factor(sigma))) + 
+  ggplot(aes(migration_rate, mean_cc, group=factor(sigma))) + 
   geom_ribbon(aes(ymin=lower_pcc, ymax=higher_pcc, fill=factor(sigma)), size=1,alpha=0.4) +
   geom_line(aes(y=mean_pcc, color=factor(sigma)), size=0.5, linetype='dashed',) +
   geom_ribbon(aes(ymin=lowest_pcc, ymax=highest_pcc, fill=factor(sigma)), size=1,alpha=0.4) +

### 2020-09-13 figs/figure9_variable_scales/figure9.jl

@@ -1,6 +1,6 @@
 
 # Project Folder
-proj_folder = "/home/michael/phase_transitions_in_metapopulation_synchrony/"
+proj_folder = "/home/michael/phase-transitions-in-metapopulation-synchrony/"
 cd(proj_folder)
 
 include(string(proj_folder, "./src/deps.jl"))
@@ -24,8 +24,8 @@ param_dictionary = Dict(
                         "lambda"            => [1.0],
                         "sigma"             => [0.5],
                         "carrying_capacity" => [1.0],
-                        "num_populations" => [5,10,20,40,80],
-                        "alpha"           => [0, 8, 16, 24],  # alpha is meaningless with two pops
+                        "num_populations" => [5,10,20,40],
+                        "alpha"           => [0, 5, 10, 15],  # alpha is meaningless with two pops
                         "summary_stat" => [random_PCC_within_radius],
                         # _________________________________________________
                         #

### 2020-09-13 figs/figure9_variable_scales/figure9_vis.R

@@ -1,10 +1,10 @@
-setwd("~/phase_transitions_in_metapopulation_synchrony/figs/figure9_variable_scales/output")
+setwd("~/phase-transitions-in-metapopulation-synchrony/figs/figure9_variable_scales/output")
 library(tidyverse)
 library(ggthemr)
 library(latex2exp)
 library(extrafont)
 loadfonts()
-ggthemr('fresh', spacing=3)
+ggthemr('fresh', spacing=3, layout='scientific')
 cmuserif = pdfFonts()$`CMU Serif`$family
 thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
             panel.spacing=unit(3, "lines"),
@@ -37,7 +37,7 @@ plt1= data %>%
   facet_wrap(. ~(m_facet), ncol=3, labeller = as_labeller(convert_to_tex_label, label_parsed)) + 
   thm 
 
-output_path = "~/phase_transitions_in_metapopulation_synchrony/writing/figs/figure9.png"
+output_path = "~/phase-transitions-in-metapopulation-synchrony/writing/figs/figure9.png"
 ggsave(output_path, plot=plt1, dpi=320, width = 18, height = 10, units = "in", device=png())
 
 plt2 =  data %>% 
@@ -58,7 +58,7 @@ plt2 =  data %>%
   labs(x=TeX("$radius$"), y=TeX("$PCC$"), color=TeX("$\\alpha$"),shape=TeX("$\\alpha$"))  
   
 
-output_path = "~/phase_transitions_in_metapopulation_synchrony/writing/figs/figure10.png"
+output_path = "~/phase-transitions-in-metapopulation-synchrony/writing/figs/figure10.png"
 ggsave(output_path, plot=plt2, dpi=320, width = 12, height = 14, units = "in", device=png())
 
 

### 2020-09-13 src/DispersalPotential.jl

@@ -1,22 +1,22 @@
 
-function get_dispersal_potential(coordinates, kernel, alpha::Number)
+function get_dispersal_matrix(coordinates, kernel, alpha)
     n_pops::Int64 = length(coordinates[:,1])
     matrix::Array{Float64,2} = zeros(n_pops, n_pops)
     for p1 = 1:n_pops
-        row_sum::Float64 = 0.0
         for p2 = 1:n_pops
             if (p1 != p2)
                 d_ij::Float64 = get_distance_between_pops(coordinates[p1], coordinates[p2])
                 matrix[p1,p2] = kernel(alpha, d_ij)
-                row_sum += matrix[p1,p2]
-            end
-        end
-
-        for p2 = 1:n_pops
-            if (p1 != p2)
-                matrix[p1,p2] /= row_sum
             end
         end
     end
-    return DispersalPotential(matrix, kernel, alpha)
+    return matrix 
+end
+
+function get_dispersal_potential(coordinates, kernel, alpha::Number)
+  dispersal_matrix = get_dispersal_matrix(coordinates, kernel, alpha)
+  for p = 1:n_pops
+    dispersal_matrix[p,:] = dispersal_matrix[p,:] / sum(dispersal_matrix[p,:])    
+  end
+  return DispersalPotential(matrix, kernel, alpha)
 end

### 2020-09-13 src/Metapopulation.jl

@@ -21,3 +21,8 @@ function get_random_metapopulation(;num_populations::Int64=10, dimensions::Int64
 
     return(Metapopulation(num_populations, populations, get_dispersal_potential(coordinates, kernel, alpha)))
 end
+
+
+function get_coordinates(mp::Metapopulation)
+    return [metapopulation.populations[p].coordinate for p = 1:metapopulation.num_populations]
+end

### 2020-09-13 src/SummaryStats.jl

@@ -62,7 +62,7 @@ function get_populations_within_circle(mp, center, radius)
     return pops_within_circle
 end
 
-function random_PCC_within_radius(treatment_instance::TreatmentInstance, treatment_num::Int64, replicate_num::Int64; radii = collect(0.0:0.05:(0.5)), num_samples_each_scale=20)
+function random_PCC_within_radius(treatment_instance::TreatmentInstance, treatment_num::Int64, replicate_num::Int64; radii = collect(0.0:0.05:(0.5)), num_samples_each_scale=50)
     abundance_matrix = treatment_instance.abundance_matrix
     mp = treatment_instance.metapopulation
 
@@ -85,3 +85,44 @@ function random_PCC_within_radius(treatment_instance::TreatmentInstance, treatme
 
     return df
 end
+
+function PCC_and_Eigencentrality(treatment_instance::TreatmentInstnace, treatment_num::Int64, replicate_num::Int64)
+    # for each population, compute the eigenvector centrality of each location
+    # based on its dispersal matrix
+    #           (as defined in hansi & ovaskainen (2000), A_ij = A_iA_j e^{-1*d_{ij}*alpha})
+    # and then compute the mean of each pairwise comparison for each population
+    
+    num_populations::Int64 = treatment_instance.metapopulation.num_populations
+    mp::Metapopulation = treatment_instance.metapopulation
+  
+    df = DataFrame(treatment=[], replicate=[], pcc=[], eigencentrality=[])
+
+    dispersal_matrix = get_dispersal_matrix(get_coordinates(mp), ExpKernel, 1.0)
+    eig_centrality = eigvals(dispersal_matrix) 
+
+    for p in 1:num_populations 
+        s = 0.0
+        for p2 in 1:num_populations
+                 if (p != p2)
+                    v1 = abundance_matrix[:,p1]
+                    v2 = abundance_matrix[:,p2]
+                    cc = crosscor(v1,v2,[0])
+                    s += cc[1]
+                    ct += 1
+                end
+        end
+        
+        mean_pcc = s/(num_populations-1)
+        ec_i = eig_centrality[p] 
+
+        push!(df.treatment, treatment_num)
+        push!(df.replicate, replicate_num)
+        push!(df.pcc, mean_pcc)
+        push!(df.eigencentrality, ec_i)
+    end
+    return df
+end
+
+
+
+

### 2020-09-13 src/deps.jl

@@ -1,7 +1,7 @@
 using StatsBase, Distributions, DataFrames, CSV, ProgressMeter, LinearAlgebra, Distances, Plots
 
 wd = pwd()
-cd("/home/michael/phase_transitions_in_metapopulation_synchrony/src")
+cd("/home/michael/phase-transitions-in-metapopulation-synchrony/src")
 
 include("./types.jl")
 include("./Parameters.jl")

### 2020-09-13 test.jl

@@ -1,50 +1,0 @@
-include("./src/deps.jl")
-
-
-
-#model = get_treatment()
-#@show model
-#run_dynamics(model)
-#plot_timeseries_of_abundances(model)
-
-
-treatment_dictionary =
-
-                    Dict(
-                        # =======================================
-                        # dynamics model parameters
-                        # =======================================
-                        "migration_rate_distribution" => [i for i = 0.01:0.01:1.0],
-                        "lambda_distribution" => [Uniform(0.5,1.5)],
-                        "sigma_distribution" => [1.0],
-                        "carrying_capacity_distribution" => [Uniform(0.5,1.5)],
-
-                        "parameters_fixed_across_populations" => [false],
-
-                        # ========================================
-                        # parameters for metapopulation generation
-                        # ========================================
-                        "metapopulation_generator" => [get_random_metapopulation],
-                        "num_populations" => [20],
-                        "alpha" => [0, 10, 20],
-                        "fixed_metapopulation_per_treatment" => [false],
-
-                        # ===============================================================
-                        # parameters that should not change unless you have a good reason
-                        # ===============================================================
-                        "number_of_timesteps" => [300]
-                    );
-
-
-
-treatment_set = create_treatments(treatment_dictionary, replicates_per_treatment=1)
-df = run_treatments(treatment_set)
-
-inst = treatment_set.treatments[200].instances[10]
-inst
-plot(collect(1:length(inst.abundance_matrix[:,1])), inst.abundance_matrix)
-
-
-cd("/home/michael/phase_transitions_in_metapopulation_synchrony/")
-CSV.write("output/metadata.csv",treatment_set.metadata)
-CSV.write("output/treatment_set.csv", df)

### 2020-09-17 writing/manuscript.tex
updating
updates sept17

@@
 
-\title{Critical Transitions in Metapopulation Connectivity}
+\title{Critical Transitions in the Spatial Synchrony of Metapopulation Dynamics}
 \author[1,2]{M.D. Catchen}
@@
 
-Landscape connectivity intro
+Anthropogenic activity has rapidly reshaped Earth's habitats.
+
+To mitigate the known contribution of habitat loss to declining biodiversity \cite{}, much attention has been paid to promoting \textit{landscape connectivity}.
+
+It is still largely unknown how ecological systems respond to habitat loss and enduing subdivision.
+
+The metapopulation framework has sought to answer this question.
+
+Spatial synchrony
+
+What is the spatial scale that constitutes a `single' population?
+How does landscape connectivity effect the scale of spatial synchrony?
+
+Review of synchrony and literature.
+  Spatial Synchrony
+  Moran's Rule
+  Dispersal Induced Stability
+
+Better understand landscape structure and synchrony, as a step toward understanding
+planning landscapes to minimize the probability of extinction.
 
@@
 From collections of cells acting together: the heart beating in rhythm \cite{womelsdorf_modulation_2007}, neurons firing in unison \cite{strogatz_sync_2003}---to behavior among organisms: the flash of fireflies \cite{otte_theories_1980} or the migration of birds \cite{spottiswoode_extrapair_2004}---to interactions between organisms: synchrony between abundances of predators and prey \cite{lv}, and of phenology \cite{van_asch_phenology_2007,burkle_future_2011}.
+
 Synchrony, by definition, involves different entities changing over time in the same way.
@@
 
+Does eigenvector centrality predict syncrhony?
+
 \hypertarget{conclusion}{%

### 2020-09-17 figs/figure10_lattice/figure10.jl

@@ -4,8 +4,7 @@ proj_folder = "/home/michael/phase-transitions-in-metapopulation-synchrony/"
 cd(proj_folder)
 
 include(string(proj_folder, "./src/deps.jl"))
-output_dir_path = string(proj_folder , "figs/figure2_synchrony_example/output/")
-abd_path = string(proj_folder, "figs/figure2_synchrony_example/output/abundances.csv")
+output_dir_path = string(proj_folder , "figs/figure10_lattice/output/")
 mkdir(output_dir_path)
 
 
@@ -21,31 +20,29 @@ mkdir(output_dir_path)
 # _______________________________________________________________________________________
 
 param_dictionary = Dict(
-                        "migration_rate"    => [0.1, 0.3, 0.7],
+                        "migration_rate"    => [x for x in 0:0.01:1 ],
                         "lambda"            => [1.0],
                         "sigma"             => [0.5],
                         "carrying_capacity" => [1.0],
-                        "num_populations" => [2, 8],
-                        "alpha"           => [0],  # alpha is meaningless with two pops
-                        "summary_stat"    => [PCC],
-
-                        
+                        "num_populations" => [64],
+                        "alpha"           => [0, 5, 10, 15],  # alpha is meaningless with two pops
+                        "summary_stat" => [PCC],
                         # _________________________________________________
                         #
                         # simulation parameters
                         #
                         # _________________________________________________
                         "number_of_timesteps" => [300],
-                        "metapopulation_generator" => [get_random_metapopulation],
+                        "metapopulation_generator" => [get_lattice_metapop],
                         "fixed_metapopulation" => [false],
-                        "log_abundances" => [true]
+                        "log_abundances" => [false] 
                     )
 
 
 
-treatment_set = create_treatments(param_dictionary, replicates_per_treatment = 1)
+treatment_set = create_treatments(param_dictionary, replicates_per_treatment = 50)
 
-df = run_treatments(treatment_set, abundances_path=abd_path)
+df = run_treatments(treatment_set)
 
 
 CSV.write(string(output_dir_path, "treatment_set.csv"), df)

### 2020-09-17 figs/figure10_lattice/figure10_vis.R

@@ -1,0 +1,28 @@
+setwd("~/phase-transitions-in-metapopulation-synchrony/figs/figure10_lattice/output")
+library(tidyverse)
+library(ggthemr)
+library(latex2exp)
+library(extrafont)
+loadfonts()
+ggthemr('fresh', spacing=3, layout='scientific')
+cmuserif = pdfFonts()$`CMU Serif`$family
+thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
+            panel.spacing=unit(3, "lines"),
+            text=element_text(family=cmuserif,size=20), axis.title.y = element_text(angle = 0, vjust=0.5),
+            axis.title = element_text(margin = margin(t = 20, r = 20, b = 0, l = 0)),
+            axis.text = element_text(size=14),
+            panel.grid = element_line(size=0.5, linetype="dotted"),
+            # The new stuff
+            strip.text = element_text(size = 22),
+            strip.text.y = element_text(size = 22, angle = 0, vjust=0.5))
+data = read.csv('metadata.csv') %>% full_join(read.csv('treatment_set.csv'), by="treatment")
+
+convert_to_tex_label = function(string) {
+  return(TeX(string))
+}
+
+
+data %>%
+  group_by(treatment) %>%
+  mutate(var_meancc=var(mean_cc)) %>% 
+    ggplot(aes(migration_rate, var_meancc)) + geom_point() + facet_wrap(. ~ alpha)

### 2020-09-17 figs/figure1_metapopulations_as_a_network/figure1.R

@@ -7,7 +7,7 @@ library(extrafont)
 library(latex2exp)
 library(Cairo)
 library(extrafont)
-setwd("~/phase_transitions_in_metapopulation_synchrony/figs")
+setwd("~/phase-transitions-in-metapopulation-synchrony/figs/figure1_metapopulations_as_a_network/output")
 loadfonts()
 cmuserif = pdfFonts()$`CMU Serif`$family
 ## ============================================
@@ -60,7 +60,7 @@ get_pop_points = function(x_vals, y_vals, alpha, mig_prop, kernel=exp_kern, kern
             x2 = x_vals[pt_ct2]
             y2 = y_vals[pt_ct2]
             #opac =1
-              opac = 0.1 + dispersal_potential[pt_ct, pt_ct2] * 8
+              opac = 0.1 + dispersal_potential[pt_ct, pt_ct2] * 10
               width = dispersal_potential[pt_ct, pt_ct2] * 8
 
 
@@ -108,6 +108,14 @@ get_pop_points = function(x_vals, y_vals, alpha, mig_prop, kernel=exp_kern, kern
 }
 #tikz(file = "~/phase_transitions_in_metapopulation_synchrony/writing/figure0.tex", width = 12, height = 8, standAlone = T)
 
+read.csv('metadata.csv') %>% 
+    full_join(read.csv('metapopulations.csv'), by='treatment') %>%
+    ggplot(aes(x,y)) + geom_point() + facet_wrap(. ~ treatment)
+
+
+
+
+
 n_pops = 20
 
 set.seed(5)
@@ -127,7 +135,7 @@ g = grid.arrange(plt1,plt2,plt3,plt4,plt5,plt6, nrow=2, padding=unit(3.5, 'line'
 plot(g)
 
 
-output_path = "~/phase_transitions_in_metapopulation_synchrony/writing/figs/figure1.png"
+output_path = "~/phase-transitions-in-metapopulation-synchrony/writing/figs/figure1.png"
 ggsave(output_path, plot=g, dpi=320, width = 12, height = 8, units = "in", device=png())
 
 

### 2020-09-17 figs/figure1_metapopulations_as_a_network/figure1.jl

@@ -4,8 +4,9 @@ proj_folder = "/home/michael/phase-transitions-in-metapopulation-synchrony/"
 cd(proj_folder)
 
 include(string(proj_folder, "./src/deps.jl"))
-output_dir_path = string(proj_folder , "figs/figure2_synchrony_example/output/")
-abd_path = string(proj_folder, "figs/figure2_synchrony_example/output/abundances.csv")
+output_dir_path = string(proj_folder , "figs/figure1_metapopulations_as_a_network/output/")
+abd_path = string(proj_folder, "figs/figure1_metapopulations_as_a_network/output/abundances.csv")
+mp_path = string(proj_folder, "figs/figure1_metapopulations_as_a_network/output/metapopulations.csv")
 mkdir(output_dir_path)
 
 
@@ -25,7 +26,7 @@ param_dictionary = Dict(
                         "lambda"            => [1.0],
                         "sigma"             => [0.5],
                         "carrying_capacity" => [1.0],
-                        "num_populations" => [2, 8],
+                        "num_populations" => [8],
                         "alpha"           => [0],  # alpha is meaningless with two pops
                         "summary_stat"    => [PCC],
 
@@ -36,16 +37,17 @@ param_dictionary = Dict(
                         #
                         # _________________________________________________
                         "number_of_timesteps" => [300],
-                        "metapopulation_generator" => [get_random_metapopulation],
-                        "fixed_metapopulation" => [false],
-                        "log_abundances" => [true]
+                        "metapopulation_generator" => [get_metapopulation_from_coordinates],
+                        "fixed_metapopulation" => [true],
+                        "log_abundances" => [true],
+                        "log_metapopulations" => [true]
                     )
 
 
 
 treatment_set = create_treatments(param_dictionary, replicates_per_treatment = 1)
 
-df = run_treatments(treatment_set, abundances_path=abd_path)
+df = run_treatments(treatment_set, abundances_path=abd_path, metapopulations_path=mp_path)
 
 
 CSV.write(string(output_dir_path, "treatment_set.csv"), df)

### 2020-09-17 figs/figure2_synchrony_example/figure2_vis.R

@@ -24,10 +24,11 @@ convert_to_tex_label = function(string) {
 }
 
 plt = data %>% 
+  group_by(treatment) %>%
   mutate(num_pops_facet=paste("$N_p = ", num_populations, "$")) %>%
   mutate(migration_facet=paste("$m = ", migration_rate, "$")) %>%
   ggplot(aes(timestep, abundance, group=interaction(replicate,population), color=factor(population))) + 
-    geom_line(alpha=(0.7), size=1.3) + 
+    geom_line(alpha=(0.8), size=1.3) + 
     facet_grid(vars(num_pops_facet), vars(migration_facet), labeller = as_labeller(convert_to_tex_label, label_parsed)) +
     labs(title='', y=TeX("$Abundance$"), x=TeX("$m$"), color="Population") +
     geom_hline(aes(yintercept=1.0) ,linetype="dashed", color='black') + 

### 2020-09-17 figs/figure5_20population_phase_transition_varying_sigma_and_lambda/figure5_vis.R

@@ -72,10 +72,12 @@ get_logit_slope = function(data, pcc, mig ){
 
 data %>%
   group_by(treatment) %>%
-  mutate(mean_pcc = mean(summary_stat)) %>%
+  mutate(var_pcc = var(mean_cc)) %>%
+  mutate(sigma_facet = paste("$\\sigma =", as.numeric(sigma), "$")) %>%
+  mutate(lambda_facet = paste("$\\lambda =", as.numeric(lambda), "$")) %>%
   ungroup() %>% 
   group_by(lambda, sigma) %>%
-  ggplot(aes(migration_rate, mean_pcc)) + facet_grid(vars(lambda), vars(sigma)) + geom_point()
+  ggplot(aes(migration_rate, var_pcc)) + facet_grid(vars(lambda_facet), vars(sigma_facet)) + geom_point()
   
 data %>%
   group_by(treatment) %>%

### 2020-09-17 src/DispersalPotential.jl

@@ -14,9 +14,10 @@ function get_dispersal_matrix(coordinates, kernel, alpha)
 end
 
 function get_dispersal_potential(coordinates, kernel, alpha::Number)
+  n_pops::Int64 = length(coordinates[:,1])
   dispersal_matrix = get_dispersal_matrix(coordinates, kernel, alpha)
   for p = 1:n_pops
     dispersal_matrix[p,:] = dispersal_matrix[p,:] / sum(dispersal_matrix[p,:])    
   end
-  return DispersalPotential(matrix, kernel, alpha)
+  return DispersalPotential(dispersal_matrix, kernel, alpha)
 end

### 2020-09-17 src/Dynamics.jl

@@ -19,6 +19,8 @@ function run_dynamics(treatment_instance::TreatmentInstance)
                 abundance_matrix[abundance_matrix_time_index,p] = treatment_instance.state[p]
             end
             abundance_matrix_time_index += 1
+
+
         end
     end
     return abundance_matrix

### 2020-09-17 src/Logging.jl

@@ -21,3 +21,23 @@ function log_abundances(treatment_num, replicate, treatment_instance::TreatmentI
         CSV.write(filename, df, append=true)
     end
 end
+
+function log_metapopulation(treatment_num, replicate, treatment_instance::TreatmentInstance; filename="./metapopulations.csv")
+    df = DataFrame(treatment=[], replicate=[], population=[], x=[],y=[])     
+    n_pops = treatment_instance.metapopulation.num_populations
+
+    for p = 1:n_pops
+        pop = treatment_instance.metapopulation.populations[p]
+        push!(df.replicate, replicate)
+        push!(df.treatment, treatment_num)
+        push!(df.population, p)
+        push!(df.x, pop.coordinate[1])
+        push!(df.y, pop.coordinate[2])
+    end
+        
+    if (!isfile(filename))
+        CSV.write(filename, df, append=false)
+    else
+        CSV.write(filename, df, append=true)
+    end
+end

### 2020-09-17 src/Metapopulation.jl

@@ -2,13 +2,13 @@ function get_distance_between_pops(coord1::Array{Float64,1}, coord2::Array{Float
     distance::Float64 = evaluate(norm, coord1, coord2)
     return distance
 end
+
 function get_distance_between_pops(pop1::Population, pop2::Population; norm=Euclidean())
     distance::Float64 = evaluate(norm, pop1.coordinate, pop2.coordinate)
     return(distance)
 end
 
 function get_random_metapopulation(;num_populations::Int64=10, dimensions::Int64 = 2, alpha::Number = 3.0, kernel=ExpKernel)
-
     populations::Vector{Population} = []
     coordinates::Vector{Array{Float64}} = []
 
@@ -22,7 +22,43 @@ function get_random_metapopulation(;num_populations::Int64=10, dimensions::Int64
     return(Metapopulation(num_populations, populations, get_dispersal_potential(coordinates, kernel, alpha)))
 end
 
+function get_lattice_metapop(; num_populations=25, alpha::Number=5.0, kernel=ExpKernel)
+    side_length::Int64 = floor(sqrt(num_populations))
+    unit_distance::Float64 = (1.0/side_length)
+
+    populations::Vector{Population} = []
+    coordinates::Vector{Array{Float64}} = []
+   
+    for x = 1:side_length
+        for y = 1:side_length
+            coordinate = [x*unit_distance, y*unit_distance]
+            push!(populations, Population(coordinate))
+            push!(coordinates, coordinate)
+        end
+    end
+    return(Metapopulation(num_populations, populations, get_dispersal_potential(coordinates, kernel, alpha)))
+end
 
 function get_coordinates(mp::Metapopulation)
-    return [metapopulation.populations[p].coordinate for p = 1:metapopulation.num_populations]
+    return [mp.populations[p].coordinate for p = 1:mp.num_populations]
 end
+
+function construct_metapopulation_for_coordinates(coordinates::Array{Float64}; kernel=ExpKernel(), alpha=3.0)
+    num_pops = length(coordinates[:, 1])
+    dims = length(coordinates[1,:])
+    
+    if (dims > num_pops)
+        print("Dimensionality of given coordinates is greater than the number of populations....\n
+                Did you mean this?\n") 
+    end
+
+    populations::Vector{Population} = []
+
+    for p = 1:num_pops
+        push!(populations, Population(coorindates[p,:]))
+    end
+
+    return (Metapopulation(num_pops, populations, get_dispersal_potential(coordinates, kernel, alpha)))
+end
+
+

### 2020-09-17 src/SummaryStats.jl

@@ -86,7 +86,7 @@ function random_PCC_within_radius(treatment_instance::TreatmentInstance, treatme
     return df
 end
 
-function PCC_and_Eigencentrality(treatment_instance::TreatmentInstnace, treatment_num::Int64, replicate_num::Int64)
+function PCC_and_Eigencentrality(treatment_instance::TreatmentInstance, treatment_num::Int64, replicate_num::Int64)
     # for each population, compute the eigenvector centrality of each location
     # based on its dispersal matrix
     #           (as defined in hansi & ovaskainen (2000), A_ij = A_iA_j e^{-1*d_{ij}*alpha})
@@ -95,30 +95,34 @@ function PCC_and_Eigencentrality(treatment_instance::TreatmentInstnace, treatmen
     num_populations::Int64 = treatment_instance.metapopulation.num_populations
     mp::Metapopulation = treatment_instance.metapopulation
   
-    df = DataFrame(treatment=[], replicate=[], pcc=[], eigencentrality=[])
+    df = DataFrame(treatment=[], replicate=[], pcc=[], closeness=[])
 
     dispersal_matrix = get_dispersal_matrix(get_coordinates(mp), ExpKernel, 1.0)
-    eig_centrality = eigvals(dispersal_matrix) 
+    
+    g = SimpleWeightedGraph(dispersal_matrix)
+
+    closeness = closeness_centrality(g)
+
+    abundance_matrix = treatment_instance.abundance_matrix
 
     for p in 1:num_populations 
         s = 0.0
         for p2 in 1:num_populations
                  if (p != p2)
-                    v1 = abundance_matrix[:,p1]
+                    v1 = abundance_matrix[:,p]
                     v2 = abundance_matrix[:,p2]
                     cc = crosscor(v1,v2,[0])
                     s += cc[1]
-                    ct += 1
                 end
         end
         
         mean_pcc = s/(num_populations-1)
-        ec_i = eig_centrality[p] 
+        centr_i = closeness[p]
 
         push!(df.treatment, treatment_num)
         push!(df.replicate, replicate_num)
         push!(df.pcc, mean_pcc)
-        push!(df.eigencentrality, ec_i)
+        push!(df.closeness, centr_i)
     end
     return df
 end

### 2020-09-17 src/TreatmentFactory.jl

@@ -62,7 +62,6 @@ function create_treatments(param_dict::Dict; replicates_per_treatment::Int64 = 5
 
 	treatments::Vector{Treatment} = []
 
-	# Treatment(mp, dynamics_model, sim_params, param_bundle, param_values, PCC(), log, abundance_matrix, initial_condition)
 	for t = 1:n_treatments
 		num_pops = metadata.num_populations[t]
 		alpha = metadata.alpha[t]
@@ -85,13 +84,9 @@ function create_treatments(param_dict::Dict; replicates_per_treatment::Int64 = 5
 		dynamics_model = StochasticLogisticWDiffusion()
 		sim_params = SimulationParameters(metadata.number_of_timesteps[t], 0.1, 10, false)
 
-		if (metadata.fixed_metapopulation[t])
-
-		end
-
 		mp = metadata.metapopulation_generator[t](num_populations=num_pops, alpha=alpha)
 
-		tr = Treatment(mp, dynamics_model, sim_params, param_bundle, metadata.summary_stat[t], metadata.log_abundances[t], [])
+        tr = Treatment(mp, dynamics_model, sim_params, param_bundle, metadata.summary_stat[t], metadata.log_abundances[t],metadata.log_metapopulations[t] , [])
 		push!(treatments, tr)
 	end
 
@@ -104,7 +99,7 @@ end
 #
 # =================================================
 
-function run_treatments(treatment_set::TreatmentSet; abundances_path = "./abundances.csv")
+function run_treatments(treatment_set::TreatmentSet; abundances_path = "./abundances.csv", metapopulations_path = "./metapopulations.csv")
 	n_treatments::Int64 = length(treatment_set.treatments)
 	n_replicates::Int64 = treatment_set.replicates_per_treatment
 
@@ -132,6 +127,11 @@ function run_treatments(treatment_set::TreatmentSet; abundances_path = "./abunda
 			if (treatments[t].log_abundances)
 				log_abundances(t, r, treatment_instance, filename=abundances_path)
 			end
+
+            if (treatments[t].log_metapopulation)
+                log_metapopulation(t,r,treatment_instance, filename=metapopulations_path)
+            end
+
 		end
 	end
 

### 2020-09-17 src/deps.jl

@@ -1,4 +1,4 @@
-using StatsBase, Distributions, DataFrames, CSV, ProgressMeter, LinearAlgebra, Distances, Plots
+using StatsBase, Distributions, DataFrames, CSV, ProgressMeter, LinearAlgebra, Distances, Plots, LightGraphs, SimpleWeightedGraphs
 
 wd = pwd()
 cd("/home/michael/phase-transitions-in-metapopulation-synchrony/src")

### 2020-09-17 src/types.jl

@@ -140,6 +140,7 @@ mutable struct Treatment
     theta::ParameterBundle
     summary_stat::Function
     log_abundances::Bool
+    log_metapopulation::Bool
     instances::Vector{TreatmentInstance}
 end
 

### 2020-09-24 figs/figure6_even_mixing/figure6_vis.R
moving to overleaf, starting the software paper

@@ -1,8 +1,8 @@
-setwd("~/phase-transitions-in-metapopulation-synchrony/figs/figure6_even_mixing/output")
+setwd("~/papers/phase-transitions-in-metapopulation-synchrony/figs/figure6_even_mixing/output")
 library(tidyverse)
 library(ggthemr)
 library(latex2exp)
-ggthemr('fresh', spacing=3)
+ggthemr('fresh', spacing=3, layout='scientific')
 cmuserif = pdfFonts()$`CMU Serif`$family
 thm = theme(panel.border = element_rect(colour = "#222222", fill = NA, size=0.75), 
             panel.spacing=unit(3, "lines"),
@@ -28,16 +28,16 @@ plt = data %>%
   ggplot(aes(num_populations,migration_rate, fill=factor(alpha)))  + 
   geom_ribbon(aes(ymin=lowest, ymax=highest), alpha=0.3) +
   geom_ribbon(aes(ymin=low, ymax=high), alpha=0.4) +
- # geom_point(aes(y=migration_rate, color=factor(alpha))) + 
+  geom_point(aes(y=migration_rate, color=factor(alpha))) + 
   stat_function(fun = function(x) 1.0 - 1.0/x) + 
   labs(x=TeX("$N_p$"), y=TeX("$m$"))  + 
   facet_wrap(. ~ alpha) + 
-  scale_x_continuous(breaks=seq(2,26,by=2), limits=c(2,26)) +
+  scale_x_continuous(breaks=seq(2,25,by=2), limits=c(2,26)) +
   scale_y_continuous(breaks=c(0.5,0.6,0.7,0.8,0.9,1.0), limits=c(0.5,1)) +
   thm +
   coord_cartesian(ylim = c(0.5,1.0)) 
   
-
+plt
 
 output_path = "~/phase_transitions_in_metapopulation_synchrony/writing/figs/figure6.png"
 ggsave(output_path, plot=plt, dpi=320, width = 18, height = 8, units = "in", device=png())

### 2020-09-24 figs/figure9_variable_scales/figure9_vis.R

@@ -1,4 +1,4 @@
-setwd("~/phase-transitions-in-metapopulation-synchrony/figs/figure9_variable_scales/output")
+setwd("~/papers/phase-transitions-in-metapopulation-synchrony/figs/figure9_variable_scales/output")
 library(tidyverse)
 library(ggthemr)
 library(latex2exp)
